### 2021-11-09 Manuscript/manuscript.tex
PUB

% !TEX TS-program = pdflatex
% !TEX root = Tes.tex
% !TEX spellcheck = en-EN

\documentclass[12pt,%                      % corpo del font principale
               a4paper,%                   % A4 papers
               %twoside,openright,%         % twoside with free right side
               oneside,openany,%           % one side
               titlepage,%                 % use a titlepage
               headinclude,footinclude,%  % header and foot header
               BCOR5mm,%                   % rilegatura di 5 mm
               cleardoublepage=empty,%     % empty pages with no header and foot
               tablecaptionabove,%         % table caption above tables
               floatperchapter,
               ]{scrreprt}                 % KOMA-Script report class;

\usepackage{braket}
\usepackage{changepage}
\usepackage[english]{babel}	% latest language is predefined
\usepackage[T1]{fontenc}		% font coding

\usepackage{indentfirst}		% indent first paragraph of each section
\usepackage{mparhack,fixltx2e,relsize}	% fancy typographies stuff

\usepackage[eulerchapternumbers,%	% chapter font Euler
            subfig,%			% in subfig objects
            beramono,%			% Bera Mono as fixed spacing font
            eulermath,%			% AMS Euler as math font
            pdfspacing,%		% improves line filling
            listings,%			% code output
 %          parts,%			% uncomment for a document divided in parts
            listsseparated,
            ]{classicthesis}		% style ClassicThesis
%\setlength{\cftbeforeloftitleskip}{100pt}
%\renewcommand{\cftbeforeloftitleskip}{1000pt}

\usepackage{arsclassica}		% modifies some aspects of ClassicThesis package
\let\marginpar\oldmarginpar		% for margin notes with \todonotes (overwise conflict with the new definition of \marginpar in classic thesis)
\usepackage[shadow]{todonotes}			% for margin notes and comments

\usepackage{bookmark}			% bookmarks

%*********************************************************************************
% Bibliography
%*********************************************************************************
%%\usepackage[style=authoryear,hyperref,backref,natbib, ,maxcitenames=1, mincitenames = 1, citestyle=authoryear-comp, backend=biber,sortcites,sorting=ynt]{biblatex}
%\usepackage[style=numeric,hyperref,backref,natbib]{biblatex}

%*********************************************************************************
% Graphics
%*********************************************************************************
\usepackage{graphicx}			% images
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage{tikz}
\usetikzlibrary{mindmap,trees}
\usetikzlibrary{backgrounds}
\usepackage{verbatim}
\usepackage[dvipsnames]{xcolor}
% \usepackage{morefloats}
%\usepackage{chngcntr}
\usepackage{pdfpages}

%*********************************************************************************
% Tables
%*********************************************************************************
\usepackage{tabularx}			% table of predefined length
\usepackage{siunitx}
\usepackage{pbox}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{hhline}
\setlength\tabcolsep{3pt}
\usepackage{changepage}
% \usepackage[showframe=true]{geometry}

%*********************************************************************************
% Mathematics and symbols
%*********************************************************************************
\usepackage{amsmath, amssymb, amsthm}	% mathematics stuff
\usepackage{mathrsfs}
\usepackage{calc}
\usepackage{algorithmic}
\usepackage[ruled]{algorithm}
\usepackage{latexsym}
\usepackage[geometry]{ifsym}
\usepackage{mathabx}
\usepackage{pifont}

%*********************************************************************************
% Personal
%*********************************************************************************
\usepackage[font=itshape]{quoting}			% fancy quotation packages. [font=small] old option
\usepackage[english]{varioref}		% complete reference package
\usepackage{hyperref}
\usepackage{url}
\usepackage[intoc, english, noprefix]{nomencl}	%for list of symbols
\usepackage[normalem]{ulem}
\usepackage{chemfig} %for chemical formulas
\usepackage{eurosym}			% euro symbol
\usepackage{epigraph}
\usepackage{calligra}
\usepackage{soul}
\usepackage[utf8]{inputenc}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\setlength{\epigraphwidth}{\textwidth}

%*********************************************************************************
% Calling personal settings and making nomenclature
%*********************************************************************************
\input{custom-commands}
\input{general-settings}  % general custom settings (margins etc)

% \makenomenclature
% \renewcommand{\nomname}{List of Symbols and Abbreviations}

\bibliographystyle{plain}

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{\normalfont\spacedallcaps{Hazardous asteroids forecast via Markov random fields}} % The article title

\subtitle{Project for the course Probabilistic modelling (DSE)} % Uncomment to display a subtitle

\author{
  Marzio De Corato
}

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

\maketitle % Print the title/author/date block

%\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

%\listoffigures % Print the list of figures%

%\listoftables % Print the list of tables

\newpage

\epigraph{
\textit{This day may possibly be my last: but the laws of probability, so true in general, so fallacious in particular, still allow about fifteen years. }\\Edward Gibbon (1737-1794)
}

\newpage

\section*{\begin{center}
Abstract
\end{center}}

\newpage

\tableofcontents % Print the table of contents

\newpage % Start the article content on the second page, remove this if you have a longer abstract that goes onto the second page

\newpage
%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\chapter{Introduction}

\chapter{Theoretical Framework}

\chapter{Dataset description}

\chapter{Results}

\chapter{Conclusion}

\bibliographystyle{unsrt}

\bibliography{sample.bib} % The file containing the bibliography

\newpage

%----------------------------------------------------------------------------------------

\end{document}

### 2021-11-09 Manuscript/structure.tex
PUB

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% Structure Specification File
%
% This file has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	REQUIRED PACKAGES
%----------------------------------------------------------------------------------------

\usepackage[ % Turn off chapters since this is an article        
beramono, % Use the Bera Mono font for monospaced text (\texttt)
eulermath,% Use the Euler font for mathematics
pdfspacing, % Makes use of pdftex’ letter spacing capabilities via the microtype package
dottedtoc % Dotted lines leading to the page numbers in the table of contents
]{classicthesis} % The layout is based on the Classic Thesis style

\titleformat{\chapter}[hang]% "hang" instead of "block"
    {\normalfont\Large\sffamily}%
    {{\color{halfgray}\chapterNumber\thechapter%
    \hspace{10pt}\vline}  }{10pt}%
    {\spacedallcaps}

\newenvironment{unnumbered}%
{\setcounter{secnumdepth}{-1}}
{\setcounter{secnumdepth}{2}}

\usepackage{arsclassica} % Modifies the Classic Thesis package

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs

\usepackage[utf8]{inputenc} % Required for including letters with accents

\usepackage{graphicx} % Required for including images

\usepackage{lscape}

\usepackage[table,xcdraw]{xcolor}

\graphicspath{{Figures/}} % Set the default folder for images

\usepackage{enumitem} % Required for manipulating the whitespace between and within lists

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{subfig} % Required for creating figures with multiple parts (subfigures)

\usepackage{amsmath,amssymb,amsthm} % For including math equations, theorems, symbols, etc

\usepackage{varioref} % More descriptive referencing

\usepackage{epigraph}

%----------------------------------------------------------------------------------------
%	THEOREM STYLES
%---------------------------------------------------------------------------------------

\theoremstyle{definition} % Define theorem styles here based on the definition style (used for definitions and examples)
\newtheorem{definition}{Definition}

\theoremstyle{plain} % Define theorem styles here based on the plain style (used for theorems, lemmas, propositions)
\newtheorem{theorem}{Theorem}

\theoremstyle{remark} % Define theorem styles here based on the remark style (used for remarks and notes)

%----------------------------------------------------------------------------------------
%	HYPERLINKS
%---------------------------------------------------------------------------------------

\hypersetup{
%draft, % Uncomment to remove all links (useful for printing in black and white)
colorlinks=true, breaklinks=true, bookmarks=true,bookmarksnumbered,
urlcolor=webbrown, linkcolor=RoyalBlue, citecolor=webgreen, % Link colors
pdftitle={Association analysis and tweets}, % PDF title
pdfauthor={\textcopyright Martina Corsini, Marzio De Corato}, % PDF Author
pdfsubject={Project for the course Algorithms for massive datasets (DSE) - DSE}, % PDF Subject
pdfkeywords={}, % PDF Keywords
pdfcreator={pdfLaTeX}, % PDF Creator
pdfproducer={LaTeX with hyperref and ClassicThesis} % PDF producer
}

### 2021-11-09 Script.R

@@ -28,11 +28,12 @@ library(randomForestExplainer)
 library(e1071)
 library(MASS)
 library(infotheo)
+library(varrank)
 
 #Data Loading
 
-Asteroids2 <- read_excel("Dataset/Asteroids2.xlsx")
-Asteroids2<-na.omit(Asteroids2)
+#Asteroids2 <- read_excel("Dataset/Asteroids2.xlsx")
+#Asteroids2<-na.omit(Asteroids2)
 
 
 
@@ -89,7 +90,7 @@ fviz_famd_ind(res.famd, col.ind = "cos2",
               gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
               repel = TRUE)
 
-heatmap(x=M_cor)
+heatmap(x=M_cor,margins = c(10,10))
 
 
 Hazardous_subset<-subset(Asteroids_FINAL,Hazardous==TRUE)
@@ -98,19 +99,21 @@ Not_Hazardous_subset<-subset(Asteroids_FINAL,Hazardous==FALSE)
 ggdensity(Asteroids_FINAL,x="Eccentricity",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
 ggdensity(Asteroids_FINAL,x="Absolute_Magnitude",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
 ggdensity(Asteroids_FINAL,x="Min_Orbit_Intersection",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
+ggdensity(Asteroids_FINAL,x="Miss_Dist",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
+ggdensity(Asteroids_FINAL,x="Close Approach Date",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
+ggdensity(Asteroids_FINAL,x="Eccentricity",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
 
 
-library(varrank)
-pippo<- varrank(data= Asteroids_FINAL_double, variable.important = "Hazardous", discretization.method = "sturges")
-plot(pippo,margins = c(15, 15, 4, 2))
+MI_var<- varrank(data= Asteroids_FINAL_double, variable.important = "Hazardous", discretization.method = "sturges")
[…]

### 2021-11-11 Manuscript/manuscript.tex
PUB

@@
 \usepackage{pdfpages}
+\usepackage{braket}
 
@@
 
+In this section we are going to review the theoretical concepts that underlies to the probabilistic methods here used as well to the physical laws that describe the dynamics of the asteroids.
+Concerning the methods,  we will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller  \cite{koller2009probabilistic} and Russel \cite{russell2010artificial}.
+Furthermore we will provide also a rapid overview of the main concepts of information theory,  since we used some of its concepts in the preliminary analysis of the dataset.
+On the other side for the concepts related to the celestial mechanics we will follow the Murray textbook \cite{murray1999solar}.
+
+\section{Markov random fields}
+
+Lets start by supposing that we would represent compactly a joint distribution such as \cite{murphy2012machine}:
+
+\begin{equation}
+p(x_{1},x_{2},...,x_{n})
+\end{equation}
+
+that can represent for instance words in a documents or pixels of an image.  Firstly we know that using the chain rule,  we can decompose it, into the following form \cite{murphy2012machine}:
+
+\begin{equation}
+p(x_{1:V})=p(x_{1})p(x_{2}|x_{1})p(x_{3}|x_{2},x_{1})...p(x_{V}|x_{1:V-1})
+\end{equation}
+
+where V is the number of variables and 1:V stands for ${1,2,...,V}$.
+This decomposition makes explicit the conditional probability tables, or in other terms the transition probability tensors \cite{wu2017markov}.
+As one can point out the number of parameter is cumbersome as the number of variables grows: indeed the number of parameter required scales as $\mathcal{O}(K^{V})$.
+Such formidable problem can be attacked by considering the concept of conditional independence.  This is defined as  \cite{murphy2012machine}:
+
+\begin{equation}
+X  \perp Y| Z \iff  p(X,Y|Z) = p(X|Z)p(Y|Z)
+\end{equation}
+
+A particular case of this definition is the Markov assumption,  by which \textit{the future is independent from the past given the present } or in symbols \cite{murphy2012machine}:
+
+\begin{equation}
+p(\textbf{x}_{1:V})=p(x_{1})\prod^{V}_{t=1}p(x_{t}|x_{t-1})
+\end{equation}
+
+In this case a first order Markov chain is obtained,  where the transtion tensor is of second order \cite{wu2017markov}.
+Given this formalism we are interested in finding a smart way to plot such joint distribution into an intuitive way: the graph theory provide the answer to this quest.
+In particular the random variables can be represented by nodes and presence of conditional indipendence for two random variables by the lack of an edge that interconnects them.
+Bayesian networks consider directed edges,  while Markov random fields (MRF) only undirected.
+As consequence,  while the the concept of topological ordering,  by which the parents n nodes are labelled with a lower  with respect to their children,  is well defined for Bayesian network,  for MRF is not.
+In order to solve this issue it is useful to consider the Hammersley-Clifford theorem as stated in \cite{murphy2012machine}:
+
+\begin{theorem}[Hammersley-Clifford]
+A positive distribution p(\textbf{y})>0 satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
+\begin{equation}
+p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
+\end{equation}
+where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
+\begin{equation}
+Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
+\end{equation}
+Note that this partition function is what ensures the overall distribution sums to 1
+\end{theorem}
+
+Such theorem allows to represent a probability distribution with potential functions for each maximal clique in the graph.
+A particular case of these is the Gibbs distribution \cite{murphy2012machine}:
+
+\begin{equation}
+p(y|\theta)=\dfrac{1}{Z(\theta)} exp\left(-\sum_{c}E(y_{c}|\theta_{c})\right)
+\end{equation}
+
+where $E(y_{c})>0$ represent the energy associated with the variables in the clique c.  This form can be adapted to a UGM with the following expression \cite{murphy2012machine}:
+
+\begin{equation}
+\psi_{c}(y_{c}|\theta_{c})=exp\left(-E(y_{c}|\theta_{c})\right)
+\end{equation}
+
+Finally in order to reduce the computational cost,  one can consider only the pairwise interaction instead of the maximum clique.
+This is the analogue of what is usually performed in solid state physics (but surely not always)  when only the interaction between the first neighbour is considered.
+Another example is the  Ising model: here we have a lattice of spins that can be or in $\ket{+}$ or in $\ket{-}$ and their interaction is modelled by\cite{murphy2012machine}:
+
+\begin{equation}
+\psi_{st}\left(y_{s},y_{t}\right) =
+\begin{pmatrix}
+e^{w_{st}} & e^{-w_{st}} \\
+e^{-w_{st}} & e^{w_{st}} \\
+\end{pmatrix}
+\end{equation}
+
+where $w_{st}=J$ represent the coupling strength between two neighbour site.  The collective state is described by
+
+\begin{equation}
+\ket{i_{1},i_{2},...,i_{n}}=\ket{i_{1}}\otimes\ket{i_{2}}\otimes...\otimes\ket{i_{n}}
+\end{equation}
+
+where $\otimes$ is the tensor product.
+If this parameter is associated with a positive finite value we have an associative Markov network: basically collective states in which all sites have the same configuration is favoured.
+Thus we will have two collective states: one for which we have all $\ket{+}$ and another in which we have all $\ket{-}$  Such situation would model,  in principle,  the ferromagnet materials where the external magnetic field induce into the material a magnetic filed with the same direction.
+On the other side if the magnetization of the material is opposite with respect to the external field,  and thus $J<0$,  we have an anti-ferromagnetic system in which frustrated states are present.
+Furthermore lets consider the unnormalized log probability of a collective state $\textbf{y}=\ket{i_{1},i_{2},...,i_{n}}$
+
+\begin{equation}
+\log\tilde{\textbf{p}}(y)= -\sum_{s\sim t}y_{s}w_{st}y_{t}
+\end{equation}
+
+If we also consider an external field
+
+\begin{equation}
+\log\tilde{\textbf{p}}(y)= -\sum_{s\sim t}y_{s}w_{st}y_{t}+\sum_{s}b_{s}y_{s}
+\end{equation}
+
+But this is nothing more that the well know \footnote{In physics} Hamiltonian of an Ising system.
+This is not a simple coincidence:  indeed the Hamiltonian of a system represent,  rudely speaking,  its total energy.
+Thus according to the Boltzmann or Gibbs distribution we have
+
+\begin{equation}
+P_{\beta}(\textbf{y})=\dfrac{e^{-\beta H(\textbf{y}})}{Z_{\beta}}
+\end{equation}
+
+where $\beta$ is proportional to the inverse of the system temperature.
+Coming back the unnormalized probability of a collective state $\textbf{y}$,  if we set $\Sigma^{-1}=\textbf{W}$, $\boldsymbol{\mu}=boldsymbol{\Sigma} b$ and $c=\dfrac{1}{2}\mu^{T}\boldsymbol{\Sigma}^{-1}\mu$ we obtain a Gaussian \cite{murray1999solar}:
+
+\begin{equation}
+\tilde{\textbf{p}}(y)\sim exp\left( -\frac{1}{2} (\textbf{y}-\boldsymbol{\mu})^{T} \boldsymbol{\Sigma}^{-1} (\textbf{y}-\boldsymbol{\mu}) + c \right)
+\end{equation}
+
 \chapter{Dataset description}

### 2021-11-11 Manuscript/manuscript.tex
PUB

@@
 Concerning the methods,  we will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller  \cite{koller2009probabilistic} and Russel \cite{russell2010artificial}.
-Furthermore we will provide also a rapid overview of the main concepts of information theory,  since we used some of its concepts in the preliminary analysis of the dataset.
+Furthermore we will provide also a rapid overview of the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches ,  since we used some of its concepts in the preliminary analysis of the dataset.
 On the other side for the concepts related to the celestial mechanics we will follow the Murray textbook \cite{murray1999solar}.
@@
 where $\beta$ is proportional to the inverse of the system temperature.
-Coming back the unnormalized probability of a collective state $\textbf{y}$,  if we set $\Sigma^{-1}=\textbf{W}$, $\boldsymbol{\mu}=boldsymbol{\Sigma} b$ and $c=\dfrac{1}{2}\mu^{T}\boldsymbol{\Sigma}^{-1}\mu$ we obtain a Gaussian \cite{murray1999solar}:
+Coming back the unnormalized probability of a collective state $\textbf{y}$,  if we set $\Sigma^{-1}=\textbf{W}$,  $\boldsymbol{\mu}=\boldsymbol{\Sigma} \textbf{b}$ and $c=\dfrac{1}{2}\mu^{T}\boldsymbol{\Sigma}^{-1}\mu$ we obtain a Gaussian \cite{murray1999solar}:
 

### 2021-11-11 Manuscript/structure.tex
PUB

@@
 urlcolor=webbrown, linkcolor=RoyalBlue, citecolor=webgreen, % Link colors
-pdftitle={Association analysis and tweets}, % PDF title
-pdfauthor={\textcopyright Martina Corsini, Marzio De Corato}, % PDF Author
-pdfsubject={Project for the course Algorithms for massive datasets (DSE) - DSE}, % PDF Subject
+pdftitle={Hazardous asteroids forecast via Markov random fields}, % PDF title
+pdfauthor={\textcopyright Marzio De Corato}, % PDF Author
+pdfsubject={Project for the course Probabilistic modelling (DSE)}, % PDF Subject
 pdfkeywords={}, % PDF Keywords

### 2021-11-12 Manuscript/manuscript.tex
PUB

@@
 \usepackage{braket}
+\usepackage{amssymb}
+\def\delequal{\mathrel{\ensurestackMath{\stackon[1pt]{=}{\scriptstyle\Delta}}}}
 
@@
 In this section we are going to review the theoretical concepts that underlies to the probabilistic methods here used as well to the physical laws that describe the dynamics of the asteroids.
-Concerning the methods,  we will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller  \cite{koller2009probabilistic} and Russel \cite{russell2010artificial}.
+Concerning the methods,  we will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller et al.
+\cite{koller2009probabilistic}, H{\o}jsgaardand \cite{hojsgaard2012graphical} et al.
+and Russel et al.
+\cite{russell2010artificial}.
 Furthermore we will provide also a rapid overview of the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches ,  since we used some of its concepts in the preliminary analysis of the dataset.
@@
 On the other side if the magnetization of the material is opposite with respect to the external field,  and thus $J<0$,  we have an anti-ferromagnetic system in which frustrated states are present.
-Furthermore lets consider the unnormalized log probability of a collective state $\textbf{y}=\ket{i_{1},i_{2},...,i_{n}}$
+Furthermore lets consider the unnormalized log probability of a collective state $\textbf{y}=\ket{i_{1},i_{2},...,i_{n}}$ \cite{murphy2012machine}:
 
@@
 
-If we also consider an external field
+If we also consider an external field \cite{murphy2012machine}:
 
@@
 This is not a simple coincidence:  indeed the Hamiltonian of a system represent,  rudely speaking,  its total energy.
-Thus according to the Boltzmann or Gibbs distribution we have
+Thus according to the Boltzmann or Gibbs distribution we have \cite{murphy2012machine}:
 
@@
 where $\beta$ is proportional to the inverse of the system temperature.
-Coming back the unnormalized probability of a collective state $\textbf{y}$,  if we set $\Sigma^{-1}=\textbf{W}$,  $\boldsymbol{\mu}=\boldsymbol{\Sigma} \textbf{b}$ and $c=\dfrac{1}{2}\mu^{T}\boldsymbol{\Sigma}^{-1}\mu$ we obtain a Gaussian \cite{murray1999solar}:
+Coming back the unnormalized probability of a collective state $\textbf{y}$,  if we set $\Sigma^{-1}=\textbf{W}$,  $\boldsymbol{\mu}=\boldsymbol{\Sigma} \textbf{b}$ and $c=\dfrac{1}{2}\mu^{T}\boldsymbol{\Sigma}^{-1}\mu$ we obtain a Gaussian \cite{murphy2012machine}:
 
@@
 \end{equation}
+
+In general we refer to Gaussian Markov random fields for a joint distribution that can be decomposed in the following way \cite{murphy2012machine}:
+
+\begin{equation}
+p\left(\textbf{y}|\boldsymbol{\theta}\right) \propto \prod_{s\sim t} \psi_{st}\left(y_{s},y_{t}\right)\prod_{t}\psi_{t}\left(y_{t}\right)
+\end{equation}
+
+\begin{equation}
+\psi_{st}\left( y_{s},y_{t} \right)=exp\left( -\dfrac{1}{2} y_{s}\Delta_{st}y_{t} \right)
+\end{equation}
+
+\begin{equation}
+\psi_{t}\left(y_{t}\right)= \exp \left( -\dfrac{1}{2}\Delta_{tt}y^{2}_{t}+\eta_{t}y_{t}\right)
+\end{equation}
+
+\begin{equation}
+p\left(\textbf{y}|\boldsymbol{\theta}\right) \propto \exp \left( \boldsymbol{\eta}^{T} \textbf{y}-\dfrac{1}{2}y^{T}\Delta \textbf{y} \right)
+\end{equation}
+
+(this last expression can be reconducted to the multivariate gassian if one consider $\boldsymbol{\Delta}=\boldsymbol{\Sigma}^{-1}$ and $\boldsymbol{\eta}=\boldsymbol{\Delta}\boldsymbol{\mu}$.
+Given the network, we would now move on how the parameters can be achieved.
+Lets start from a Markov random field in log-linear form \cite{murphy2012machine}:
+
+\begin{equation}
+p\left(\textbf{y}|\boldsymbol{\theta}\right) = \dfrac{1}{Z(\theta)}\exp \left( \sum_{c}\boldsymbol{\theta}^{T}_{c}\phi_{c}\left(\textbf{y}\right)\right)
+\end{equation}
+
+thus we can define the log-likehood as \cite{murphy2012machine}:
+
+\begin{equation}
+\mathcal{L}\left(\boldsymbol{\theta}\right):= \frac{1}{N}\sum_{i}\log p\left(\textbf{y}_{i}|\boldsymbol{\theta}\right)=\frac{1}{N}\sum_{i}\left[\sum_{c} \boldsymbol{\theta}^{T}_{c}\phi_{c}(y_{i})-\log Z\left(\boldsymbol{\theta}\right)\right]
+\end{equation}
+
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\frac{1}{N}\sum_{i}\left[\phi_{c}(y_{i})-\frac{\partial}{\partial\boldsymbol{\theta}_{c}}\log Z(\boldsymbol{\theta})\right]
+\end{equation}
+
+\begin{equation}
+\frac{\partial \log Z(\boldsymbol{\theta})}{\partial\boldsymbol{\theta}}=\mathbb{E}\left[\phi_{c}(\textbf{y})\right|\theta]=\sum_{\textbf{y}}\phi_{c}(\textbf{y})p(\textbf{y}|\boldsymbol{\theta})}
+\end{equation}
+
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\left[\frac{1}{N}\sum_{i}\phi_{c}(y_{i})\right]-\mathbb{E}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+
+In the first term $\textbf{y}$ is fixed to its observed values while in the second it is free. Such expression can be recasted in to a more explicative form \cite{murphy2012machine}:
+
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]-\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+
+Therfore at the optimum we will have \cite{murphy2012machine}:
+
+\begin{equation}
+mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]=\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+
+From this expression it is clear why this method is called moment matching; it is worth nothing tha such computation is largely expensive from a computational point of view: thus scholar usually consider other techinques or at least stochastic gradient descent method.
+A full review can be found in \cite{murphy2012machine} and \citep{koller2009probabilistic}.
 

### 2021-11-12 Manuscript/manuscript.tex
PUB

@@
 
-In this section we are going to review the theoretical concepts that underlies to the probabilistic methods here used as well to the physical laws that describe the dynamics of the asteroids.
-Concerning the methods,  we will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller et al.
+In this section we are going to review the theoretical concepts that underlies to the probabilistic methods here used:  we will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller et al.
 \cite{koller2009probabilistic}, H{\o}jsgaardand \cite{hojsgaard2012graphical} et al.
@@
 Furthermore we will provide also a rapid overview of the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches ,  since we used some of its concepts in the preliminary analysis of the dataset.
-On the other side for the concepts related to the celestial mechanics we will follow the Murray textbook \cite{murray1999solar}.
+On the other side the concepts related to the celestial mechanics here used will be described, following the Murray approach \cite{murray1999solar} into the Appendix A
 
@@
 \begin{equation}
-mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]=\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
+\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]=\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
 \end{equation}
@@
 From this expression it is clear why this method is called moment matching; it is worth nothing tha such computation is largely expensive from a computational point of view: thus scholar usually consider other techinques or at least stochastic gradient descent method.
-A full review can be found in \cite{murphy2012machine} and \citep{koller2009probabilistic}.
+A full review can be found in \cite{murphy2012machine} and \cite{koller2009probabilistic}.
+Finally we consider, as for the dataset in analysed in this work, the case where we have both discrete and continuous variables i.e.
+$x=\left(i_{1},...,i_{d},y_{1},...,y_{q} \right)$ with d discrete variable and q continuos variables.
+The are called in the literature Mixed Interaction Models.
+In this case the following density has to be considered \cite{hojsgaard2012graphical}:
+
+\begin{equation}
+\begin{split}
+f(i,y)=& p(i)(2\pi)^{-q/2}det(\Sigma)^{-1/2} \\
+& exp\left[-\dfrac{1}{2}\left(y-\mu(i)\right)^{T}\Sigma^{-1}\left(y-\mu(i)\right)\right]
+\end{split}
+\label{gaussMix}
+\end{equation}
+
+Which can be rewritten in the exponential family form \cite{hojsgaard2012graphical}:
+
+\begin{equation}
+\begin{split}
+f(i,y) & = \exp\left\lbrace g(i)+\sum_{u}h^{u}(i)y_{u}-\dfrac{1}{2}\sum_{uv} y_{u}y_{v}k_{uv}\right\rbrace \\
+&= \exp\left\lbrace g(i)+h(i)^{T}y-\dfrac{1}{2}y^{T}Ky \right\rbrace
+\end{split}
+\end{equation}
+
+where $g(i)$, $h(i)$ and $K$ are the canonical parameters. These are connected with the parameters of expression \ref{gaussMix} by the following identities \cite{hojsgaard2012graphical}:
+
+\begin{equation}
+\begin{split}
+K=&\Sigma^{-1} \\
+h(i)=&\Sigma^{-1}\mu(i) \\
+g(i)=&\log p(i) -\frac{1}{2}\log det (\Sigma) \\
+&-\dfrac{1}{2}\mu(i)^{T}\Sigma^{-1}\mu(i)-\dfrac{q}{2}\log 2\pi
+\end{split}
+\end{equation}
+
+Moreover one can further modify the previous form in order to obtain a particular factorial expansion: such models are referred as homogeneous mixed interaction models \cite{hojsgaard2012graphical}.
+
+\section{Information theory}
+Given an ensemble of random variable, we can find the amount of information that one variable contains of another one: such quantity is called mutual information and it is a key concept within the information theory.
+This approach, that was implemented by Claude Shannon decades before the probabilistic modelling, represent a complementary way by which one can attack the problem of conditional dependence between random variables.
+Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allows to properly define the concept of mutual information.
+The starting concept of information theory is the entropy which express the uncertainty of a random variable.
+Given a random variable $X$ with alphabet (the accessible states) $\cite{cover2006elements}:$ and probability mass function $p(x)=Pr\left\lbrace X=x \rbrace\; x \in \chi$ we define the entropy of $X$ as $H(X)=-\sum_{x \in X} p(x)log p(x)$ where the logarithm has to be considered with basis 2.
+In analogous way the joint entropy of two random variables $(X,Y)$ with a joint distribution p(x,y) is defined as \cite{cover2006elements}:
+\begin{equation}
+H(X,Y)=-\sum_{x\in \chi}\sum_{y\in \mathcal{Y}}p(x,y)\log p(x,y)
+\end{equation}
+Furthermore, we can define also the conditional entropy as \cite{cover2006elements}:
+\begin{equation}
+\begin{split}
+H(X|Y)=& \sum_{x\in \mathcal{X} }p(x)H(Y|X=x)\\
+=& -\sum_{x\in \mathcal{X}}p(x)\sum_{y\in \mathcal{Y}}p(x,y)\log p(y|x) \\
+=& -E\log p(Y|X)
+\end{split}
+\end{equation}
+The joint entropy and the conditional entropy are related by the chain rule \cite{cover2006elements}:
+\begin{equation}
+H(X,Y)=H(X)+H(Y|X)
+\end{equation}
+Such rule can be extended to to the following from \cite{cover2006elements}:
+\begin{equation}
+H(X,Y|Z)=H(X|Z)+H(Y|X,Z)
+\end{equation}
+Given a distribution q and another distribution p, one can quantify how inefficently the second one describe the first one using the concept of relative entropy or Kullback-Leibler distance \cite{cover2006elements,mackay2003information}:
+\begin{equation}
+D(p||q)=\sum p(x)\log\frac{p(x)}{q(x)}
+\end{equation}
+As stated by the Gibbs inequality \cite{cover2006elements,mackay2003information}:
+\begin{equation}
+D(p||q)\geq 0
+\end{equation}
+this quantity can not be negative: the entropy of a random variable associated to another cannot have a degree of uncertainty lower with respect to the quantity that its aimed to describe.
+On these basis we are now ready to introduce the concept of mutual information.
+This is defined as  \cite{cover2006elements}:
+\begin{equation}
+\begin{split}
+I(X;Y)&=\sum\sump(x,y)\log\dfrac{p(x,y)}{p(x)p(y)}=\\
+&=D(p(x,y)||p(x)p(y)) \\
+&=H(X)-H(X|Y)=H(Y)-H(Y|X)
+\end{split}
+\end{equation}
+As for the joint distribution also in this case we have a chain rule  \cite{cover2006elements}:
+\begin{equation}
+I(X_{1},X_{2},...,X_{n};Y)=\sum^{n}_{i=1}I(X_{i};Y|X_{i-1},X_{i-2},...,X_{1})
+\end{equation}
+Finally we would report the data process inequality theorem that connects the information theory with the Markov chain: if we have a Markov chains, $X\rightarrow Y \rightarrow Z$  then $I(X;Y)\geq I(X;Z) $.
+As for the Gibbs inequality, the underlying idea is that no clever manipulation of the data can improve the inference that can be made on them \cite{cover2006elements,mackay2003information}.
+Otherwise we would have a clear violation of the second principle of thermodynamics (see for instance the Maxwell's demon \cite{feynman2018feynman})
 
@@
 \chapter{Conclusion}
+
+\chapter{Appendix A: Concepts of celestial mechanics}
 

### 2021-11-15 Manuscript/manuscript.tex
PUB

@@
 \def\delequal{\mathrel{\ensurestackMath{\stackon[1pt]{=}{\scriptstyle\Delta}}}}
+\usepackage{lscape}
 
@@
 
+This chapter is organized in the following way: first we are going to report the results concerning the preliminary analysis performed on the dataset.
+This include the factor analysis of mixed data and the mutual information analysis of the continuous variables of asteroids vs their hazard.
+Then it will follow the analysis of the dataset performed with the probabilistic methods: after a preliminary analysis on the continuous variables, the mixed interaction model as well the minforest model obtained for the whole dataset will be presented and discussed.
+Finally the probabilistic models previously obtained will be compared with the outputs and the performances of four machine learning algorithm (Random Forest, Support vector machines Quadratic Discriminant Analysis and Logistic regression).
+
+\section{Preliminary analysis}
+
+In Fig.
+\ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig.
+\ref{FAMD_Quantitative variables} we can recognize different correlations that are given by the celestial mechanics law: for instance see that there is strong correlation of the mean motion with the semi-major axis.
+This is due to the Kepler law discussed in Appendix A.
+Furthermore we see that that the mean motion is correctly almost independent with respect to the the diameter (max or min) of  the asteroid.
+As explained in Appendix A this is an another result of Newton gravitation theory for a two body interaction: the mean motion of an asteroid, as long as the the mass of it is much lower with respect to the sun, is independent from its mass.
+On the other side the fact that relative velocity is correlated with the diameter is not surprising since here there is a three body dynamics (the Sun, the Earth and the Asteroid).
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/FAMD.pdf}
+\caption{The FAMD main plot in which the correlation between the continuous and discrete variables is reported. Plot obtained from FactoMineR package \cite{le2008factominer}}
+\label{FADM}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/FAMD_Quantitative variables.pdf}
+\caption{The FAMD correlation circle for continuos variables as obtained from FactoMineR package \cite{le2008factominer}}
+\label{FAMD_Quantitative variables}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/FAMD_Individuals_(c).pdf}
+\caption{Graph of individuals, for the qualititative variables, as obtained from FactoMineR package \cite{le2008factominer}}
+\label{FAMD_Individuals_(c)}
+\end{center}
+\end{figure}
+
+\begin{landscape}
+\begin{figure}
+\begin{center}
+\includegraphics[width=1.75\textheight]{Figures/Mutual_information.pdf}
+\caption{Mutual information as obtained with the varrank package \cite{kratzer2018varrank}}
+\label{Mutual_information}
+\end{center}
+\end{figure}
+\end{landscape}
+
+\pagebreak
+
+\section{Probabilistic models}
+
+\section{Machine learning algorithms}
+
 \chapter{Conclusion}

### 2021-11-15 Script.R

@@ -105,8 +105,8 @@ ggdensity(Asteroids_FINAL,x="Eccentricity",rug = TRUE, color = "Hazardous",fill 
 
 
 MI_var<- varrank(data= Asteroids_FINAL_double, variable.important = "Hazardous", discretization.method = "sturges")
-plot(MI_var,margins = c(13, 13, 4, 2),notecex=1,digitcell=2,labelscex=1)
-
+plot(MI_var,margins = c(8, 8, 4, 0.1),notecex=0.5,digitcell=2,labelscex=0.6,textlines = 0.,maincex =0.5)
+summary(MI_var)
 
 #Continuous variables analysis
 

### 2021-11-15 Manuscript/manuscript.tex
PUB

@@
 \ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig.
-\ref{FAMD_Quantitative variables} we can recognize different correlations that are given by the celestial mechanics law: for instance see that there is strong correlation of the mean motion with the semi-major axis.
+\ref{FAMD_Quantitative variables} we can recognize the different correlations that are given by the celestial mechanics laws: for instance see that there is strong correlation of the mean motion with the semi-major axis.
 This is due to the Kepler law discussed in Appendix A.
@@
 On the other side the fact that relative velocity is correlated with the diameter is not surprising since here there is a three body dynamics (the Sun, the Earth and the Asteroid).
+Beside this inspection a further analysis based on the concept of mutual information (summarized in the previous chapter) was performed: its result is reported in Fig.
+\ref{Mutual_information}.
+This figure summarize the ranking of the features considered in the dataset for the dangerousness classification of the asteroids according to the following expression \cite{kratzer2018varrank}
+
+\begin{equation}
+g(\alpha,\textbf{C},\textbf{S},f_{i})=MI(f_{i};\textbf{C})-\sum_{f_{s}\in S}\alpha(f_{i},f_{s},\textbf{C},\textbf{S})MI(f_{i};f_{s})
+\end{equation}
+where the first term $MI(f_{i};\textbf{C})$ is called relevance and measures the Mutual Information between the interesting feature set $\textbf{C}$ (only Hazardous in our case) and the analysed one $f_{i}$; the third term $MI(f_{i};f_{s})$ is called redundancy and measures the MI between the analysed feature and a chosen set of them.
+Finally the $\alpha(f_{i},f_{s},\textbf{C},\textbf{S})$ is a normalization function and in our case was set to \cite{kratzer2018varrank}
+
+\begin{equation}
+\alpha(f_{i},f_{s},\textbf{C},\textbf{S})=\dfrac{1}{|\textbf{S}|}
+\end{equation}
+
+following the Peng.
+et al approach \cite{peng2005feature}.
+We see that the first placein the mutual information ranking, looking to the diagonal element,  is taken by minimum orbit intersection: this is correct since this is one parameter used by NASA for deciding if an asteroid is hazardous or not (see Appendix A).
+The second place is occupied by Epoch date close approach.
+The third place is the eccentricity value: this parameter is entangled with the minimum orbit intersection and thus it is reasonable that it is important.
+Then we have two parameter that are related to the dimension of the asteroid: this is meaningful since if the asteroid has a too reduced volume it will be destroyed by the Earth atmosphere.
+On the other side the other parameters seems to have a too low MI for being interesting in this preliminary analysis.
+The results obtained for FAMD and MI will provide a useful path-guide, together with the celestial mechanics theory, for the interpretation of the results obtained in the following section
 

### 2021-11-16 Manuscript/manuscript.tex
PUB

@@
 
-In this case a first order Markov chain is obtained,  where the transtion tensor is of second order \cite{wu2017markov}.
+In this case a first order Markov chain is obtained,  where the transition tensor is of second order \cite{wu2017markov}.
 Given this formalism we are interested in finding a smart way to plot such joint distribution into an intuitive way: the graph theory provide the answer to this quest.
-In particular the random variables can be represented by nodes and presence of conditional indipendence for two random variables by the lack of an edge that interconnects them.
+In particular the random variables can be represented by nodes and presence of conditional independence for two random variables by the lack of an edge that interconnects them.
 Bayesian networks consider directed edges,  while Markov random fields (MRF) only undirected.
@@
 
+We started the analysis with the graphical models by inspecting the relations between continuous variables in the dataset, thus in this first step the binary variable \textit{Hazardous} was excluded.
+For this purpose, among the different methods available, we considered the \textit{Graphical least absolute shrinkage and selection operator} GLASSO as implemented in the \textit{glasso} R package \cite{friedman2008sparse,glasso}.
+After different tests reported in Fig.
+\ref{GLASSO_convergence}, we considered as final result the graph obtained with the value of $\rho$ (the one that penalize further connections) equal to $0.2$.
+In this plot we see that the conditional dependences/independences state by the Celestial Mechanics are reproduced: for instance we see that the mean motion is conditionally independent with respect to the the estimated diameters, but the relative velocity not.
+Furthermore the eccentricity is conditionally depended, as one can expect to quantity that describe the elliptical orbits: the perihelion distance, the semi-major axis and the Jupiter Tisserand's parameter.
+Finally we would compare the results of GLASSO with the graph intersection obtained from the stepwise model (forward direction, BIC) (as implemented in the grim package \cite{hojsgaard2012graphical}), SIN (simultaneous p-values method, as implemented in the SIN package \cite{SIN_package}) and thresholding method (as implemented in the grim package \cite{hojsgaard2012graphical}) shown in Fig.
+\ref{ST+TH+SIN_PLOT}: in this case the volume/mass of the asteroid is not connect with the Relative velocity.
+
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/GLASSO_0.1.pdf}
+    \subcaption{$\rho$=0.1}
+    \vspace{4ex}
+  \end{minipage} 
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/GLASSO_0.2.pdf}
+    \subcaption{$\rho$=0.2}
+    \vspace{4ex}
+  \end{minipage} \\
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/GLASSO_0.3.pdf}
+    \subcaption{$\rho$=0.3}
+    \vspace{4ex}
+  \end{minipage}%%
+    \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/GLASSO_0.4.pdf}
+    \subcaption{$\rho$=0.4}
+    \vspace{4ex}
+  \end{minipage}%%
+\caption{GLASSO analysis, performed with the glasso package \cite{friedman2008sparse,glasso}, with different $\rho$ parameter (the one that penalize further connections) for the Asteroid dataset without the discrete variable Hazardous.
+The plots were obtained with the igraph package for R \cite{igraph}}
+\label{GLASSO_convergence}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[\linewidth]{Figures/ST+TH+SIN_PLOT.pdf}
+\caption{Intersection of the graphs obtained from the stepwise method (forward, BIC),from the  threshold method (as implemented in the grim package \cite{hojsgaard2012graphical}) and the SIN one (as implemented in the SIN package \cite{SIN_package}).
+The plots were obtained with the igraph package for R \cite{igraph} }
+\label{ST+TH+SIN_PLOT}
+\end{center}
+\end{figure}
+
 \section{Machine learning algorithms}
 
-\chapter{Conclusion}
+\chapter{Conclusions}
 

### 2021-11-16 Script.R

@@ -111,8 +111,8 @@ summary(MI_var)
 #Continuous variables analysis
 
 
-#msx<-cmod(~.^1.,data=Asteroids_FINAL_double[,-21])
-#msx2<-stepwise(msx,direction="forward",k=log(nrow(Asteroids_FINAL_double[,-21])),details=1)
+msx<-cmod(~.^1.,data=Asteroids_FINAL_double[,-21])
+msx2<-stepwise(msx,direction="forward",k=log(nrow(Asteroids_FINAL_double[,-21])),details=1)
 #plot.igraph(as(msx2,'igraph'),layout=layout.fruchterman.reingold(as(msx2,'igraph'), niter=300000), vertex.color="red")
 
 S.asteroids <- cov.wt(Asteroids_FINAL_double[,-21], method="ML")$cov
@@ -147,6 +147,15 @@ gsin.asteroids <-as(getgraph(psin.asteroids,0.1),"graphNEL")
 plot(as(gsin.asteroids,"igraph"))
 
 
+C.asteroids  <- cov2cor(S.asteroids )
+res.lasso <- glasso(C.asteroids , rho=0.4)
+AM <- res.lasso$wi!=0
+diag(AM)<-FALSE
+g.lasso <- as(AM,"graphNEL")
+nodes(g.lasso)<-names(Asteroids_FINAL_double[,-21])
+glasso.asteroids <- cmod(edgeList(g.lasso),data=Asteroids_FINAL_double[,-21])
+plot(as(glasso.asteroids ,"igraph"),vertex.color="red",vertex.label.dist=2,alpha=TRUE,vertex.label.cex=0.8,vertex.size=10,edge.color="black",edge.size=2,layout=layout_with_fr(as(glasso.asteroids,'igraph')), vertex.label.family='Helvetica')
+
 
 
 
@@ -160,7 +169,26 @@ nodes(g.lasso)<-names(Asteroids_FINAL_double[,-21])
 glasso.asteroids <- cmod(edgeList(g.lasso),data=Asteroids_FINAL_double[,-21])
 plot(as(glasso.asteroids ,"igraph"),vertex.color="red",vertex.label.dist=2,alpha=TRUE,vertex.label.cex=0.8,vertex.size=10,edge.color="black",edge.size=2,layout=layout_with_fr(as(glasso.asteroids,'igraph')), vertex.label.family='Helvetica')
 
+#S.body<- cov.wt(Asteroids_FINAL_double[,-21])$cov
[…]

### 2021-11-16 Script.R
dd

@@ -193,7 +193,7 @@ plot(as(glasso.asteroids ,"igraph"),vertex.color="red",vertex.label.dist=2,alpha
 
 
 commonedges.asteroids <-as(msx2,"graphNEL")
-othermodels<-list(gsin.asteroids ,thresh.asteroids ,msx2)
+othermodels<-list(gsin.asteroids ,thresh.asteroids ,glasso.asteroids)
 
 
 othermodels<-lapply(othermodels,as,"graphNEL")
@@ -208,10 +208,18 @@ plot(as(commonedges.asteroids ,"igraph"),vertex.color="orange",vertex.label.dist
 
 #minForest model
 
-bf<-minForest(Asteroids_FINAL_double,homog=TRUE,forbEdges=NULL,stat="LR")
-plot(bf)
-mbG<-stepw(model=bf,data=Asteroids_FINAL_double,exact=TRUE)
+bf<-minForest(Asteroids_FINAL_double)
+plot(bf,cex.vert.label=0.6,numIter=6000,col.labels=c("red"),vert.hl=c(21),col.hl=c("blue"),energy=TRUE)
+mbG<-stepw(model=bf,data=Asteroids_FINAL_double,threshold=1)
 plot(mbG,cex.vert.label=0.6,numIter=6000,col.labels=c("red"),vert.hl=c(21),col.hl=c("blue"),energy=TRUE)
+
+
+aa<-mmod(~.^1.,data=Asteroids_FINAL[,2:22])
+aa2<-stepwise(aa,k=log(nrow(Asteroids_FINAL[,2:22])),direction="forward",details=1)
+plot(aa2)
+
+
+plot(as(aa2 ,"igraph"),vertex.color="red",vertex.label.dist=2,alpha=TRUE,vertex.label.cex=0.5,vertex.size=10,edge.color="black",edge.size=2,layout=layout_with_fr(as(aa2,'igraph')), vertex.label.family='Helvetica')
 
 
 #Mixed interaction analysis

### 2021-11-17 Script.R
PUB

@@ -101,7 +101,11 @@ ggdensity(Asteroids_FINAL,x="Absolute_Magnitude",rug = TRUE, color = "Hazardous"
 ggdensity(Asteroids_FINAL,x="Min_Orbit_Intersection",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
 ggdensity(Asteroids_FINAL,x="Miss_Dist",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
 ggdensity(Asteroids_FINAL,x="Close Approach Date",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
-ggdensity(Asteroids_FINAL,x="Eccentricity",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
+ggdensity(Asteroids_FINAL,x="Semi Major Axis",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
+ggdensity(Asteroids_FINAL,x="Perihelion Distance",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
+ggdensity(Asteroids_FINAL,x="Perihelion Time",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
+ggdensity(Asteroids_FINAL,x="Epoch_Osculation",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
+ggdensity(Asteroids_FINAL,x="Orbital Period",rug = TRUE, color = "Hazardous",fill = "Hazardous",size=2)+theme_bw()
 
 
 MI_var<- varrank(data= Asteroids_FINAL_double, variable.important = "Hazardous", discretization.method = "sturges")

### 2021-11-17 Manuscript/manuscript.tex
PUB

--- 2021-11-15 Manuscript/manuscript.tex
@@
 
+The asteroid dataset was retrieved from Kaggle \cite{kaggle_dataset}, which reports into an more user-friendly mode the dataset of The Center for Near-Earth Object Studies (CNEOS) \cite{cneos+nasa}, a NASA research center.
+Among the 40 features there were present in the dataset we excluded the ones that were clearly redundant (such as the distance quantity evaluated in miles instead of kilometres), the ones that were connected only to the other name of the asteroid, the one connected to the name of the orbit and the one connected with the orbiting planet, since for all it was the Earth.
+Thus the features obtained were reduced to 22.
+Their enumeration and their description will be postponed to Appendix A, since a description of them cannot be disjoint with celestial mechanics concepts.
+Since only the 16 \% of the asteroids were hazardous, we considered to reduce the number of the non-hazardous in order to improve the portion of the hazardous ones.
+For this purpose we constructed a database in which the proportion hazardous/not hazardous was 1:5: thus all the hazardous one were included, while the not-hazardous were randomly selected \footnote{The author is aware that in principle the proportion should be much less unbalanced, however the tool that has to be paid for this operation is an overall reduction of the cases in the dataset that lowers the performances of the algorithms here used.
+The proportion here used was a trade-off for these two contrasting facts}.
+Furthermore, concerning the continuous measures, the dataset was standardised and demeaned.
+
 \chapter{Results}
@@
 
+The first inspection that was performed on the dataset was related the density distributions of a selection of continuous features that are known, from the celestial mechanics, to be important for the prediction of the asteroids dangerousness.
+These are reported in Fig.
+\ref{Density_relevant}.
+Next we moved on a more systematic analysis with FADM and mutual information.
 In Fig.
 \ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig.
-\ref{FAMD_Quantitative variables} we can recognize the different correlations that are given by the celestial mechanics laws: for instance see that there is strong correlation of the mean motion with the semi-major axis.
+\ref{FAMD_Quantitative variables} the different correlations that are given by the celestial mechanics laws can be recognized: for instance, see that there is strong correlation of the mean motion with the semi-major axis.
 This is due to the Kepler law discussed in Appendix A.
@@
 As explained in Appendix A this is an another result of Newton gravitation theory for a two body interaction: the mean motion of an asteroid, as long as the the mass of it is much lower with respect to the sun, is independent from its mass.
-On the other side the fact that relative velocity is correlated with the diameter is not surprising since here there is a three body dynamics (the Sun, the Earth and the Asteroid).
+On the other side the fact that relative velocity is correlated with the diameter is a spurios correlation.
+At this point one can ask why the mean motion and the relative velocity are orthogonal: this point will be clarified from a theoretical point of view in appendix A and also with graphical models.
+Briefly the motion of an object on an ellipse as seen from a focus is not uniform: this is faster as the two bodies approaches.
 Beside this inspection a further analysis based on the concept of mutual information (summarized in the previous chapter) was performed: its result is reported in Fig.
@@
 et al approach \cite{peng2005feature}.
-We see that the first placein the mutual information ranking, looking to the diagonal element,  is taken by minimum orbit intersection: this is correct since this is one parameter used by NASA for deciding if an asteroid is hazardous or not (see Appendix A).
+We see that the first place in the mutual information ranking, looking to the diagonal element, is taken by minimum orbit intersection: this is correct since this is one parameter used by NASA for deciding if an asteroid is hazardous or not (see Appendix A).
 The second place is occupied by Epoch date close approach.
@@
 The results obtained for FAMD and MI will provide a useful path-guide, together with the celestial mechanics theory, for the interpretation of the results obtained in the following section
+
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Perihelion_Distance.pdf}
+    \subcaption{ \begin{center}
+    a) Perihelion Distance
+    \end{center}}
+    \vspace{4ex}
+  \end{minipage} 
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Eccentricity.pdf}
+    \subcaption{ \begin{center}
+    b) Eccentricity
+    \end{center}}
+    \vspace{4ex}
+  \end{minipage} \\
+    \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Min_orbit_intersection.pdf}
+    \subcaption{ \begin{center}
+    c) Min orbit intersection
+    \end{center}}
+    \vspace{4ex}
+  \end{minipage}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/DENSITY_Absolute_magnitude.pdf}
+    \subcaption{ \begin{center}
+    d) Semi Major Axis
+    \end{center}}
+    \vspace{4ex}
+  \end{minipage}
+\caption{Comparison between the density distributions of hazardous (red) and non-hazardous (light blue) asteroids for a selected set of features that, according to the theory, are interesting.}
+\label{Density_relevant}
+\end{figure}
 
@@
 
+We started the analysis with the graphical models by inspecting the relations between continuous variables in the dataset: thus in this first step the binary variable \textit{Hazardous} was excluded.
+For this purpose, among the different methods available, we considered the \textit{Graphical least absolute shrinkage and selection operator} GLASSO as implemented in the \textit{glasso} R package \cite{friedman2008sparse,glasso}.
+After different tests reported in Fig.
+\ref{GLASSO_convergence}, we considered as final result the graph obtained with the value of $\rho$ (the one that penalize further connections) equal to $0.3$.
+This is because, in this plot, we see that the conditional dependences/independences stated by the Celestial Mechanics are correctly reproduced: for instance we see that the diameter of the asteroids is independent from all features related to its motion.
+This is definitely meaningful since the mass of the asteroids is largely lower with respect to the mass of earth: thus there is no way by which the asteroid orbit can be modified by its mass as explained in Appendix A.
+Furthermore we see that the Close approach Date and the Epoch date are dependant each other but in no way from the other features: this is right since the date and epoch are set with an arbitrary scale.
+Also the Perihelion Epoch and Osculation Time are dependant, as expected, but there is no dependence with the orbit parameters.
+We see that the mean motion is conditionally independent with respect to relative velocity, but correctly this is dependent from perihelion distance.
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/GLASSO_0.1.pdf}
+    \subcaption{$\rho$=0.1}
+    \vspace{4ex}
+  \end{minipage} 
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/GLASSO_0.2.pdf}
+    \subcaption{$\rho$=0.2}
+    \vspace{4ex}
+  \end{minipage} \\
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/GLASSO_0.3.pdf}
+    \subcaption{$\rho$=0.3}
+    \vspace{4ex}
+  \end{minipage}%%
+    \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/GLASSO_0.4.pdf}
+    \subcaption{$\rho$=0.4}
+    \vspace{4ex}
+  \end{minipage}%%
+\caption{GLASSO analysis, performed with the glasso package \cite{friedman2008sparse,glasso}, with different $\rho$ parameter (the one that penalize further connections) for the Asteroid dataset without the discrete variable Hazardous.
+The plots were obtained with the \textit{igraph} package for R \cite{igraph}}
+\label{GLASSO_convergence}
+\end{figure}
+Lets now move on the mixed interaction model: first we considered its implementation in the mgm package \cite{mgm,haslbeck2015mgm}.
+We choose as $k$ parameter 2 and a cross validation (CV) with ten folds.
+The result is reported in Fig.
+\ref{mgm}.
+From this figure we see that the features that are conditionally dependant to the Hazardous features are: the absolute magnitude, the min orbit intersection, and the eccentricity.
+All these connections are supported by the theory: the min orbit intersection is the main parameter for the hazardous value, the eccentricity on the other side can be tough as a parameter that describes how near the celestial body moves at the perihelion (indeed this parameter is correctly linked with the perihelion distance) and the Absolute Magnitude is also a meaningful parameter for the hazard evaluation since, if the asteroid is too small it will be destroyed by the earth atmosphere.
+It is worth mentioning that this parameter is correctly not connected with all quantities related to the orbital parameter except for the mean anomaly.
+Furthermore the parameters related to the diameters are also not connected with the orbital parameters.
+Furthermore we see also that the mean velocity is not directly related to the mean motion but there are the orbital parameters within them as expected.
+Thus it can be said that the model produced is almost consistent with the astronomic laws.
+On this basis, we can evaluate the performances of the model in term of confusion matrix, ROC curve and $\phi$ parameters (we used for this purpose the Caret R package \cite{kuhn2008building,caret}).
+The confusion matrix obtained is reported in Fig.
+\ref{mgm_confusion}: the ROC curve is given in Fig \ref{ROC_mgm} and it corresponds to a $\phi=0.6$
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/mgm.pdf}
+\caption{The graphical model obtained with the mixed interaction model as implemented in the mgm package \cite{mgm,haslbeck2015mgm}. The plot was obtained with the \cite{qgraph} package}
+\label{mgm}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/mgm_confusion.pdf}
+\caption{The confusion matrix of the graphical model reported in Fig. \ref{mgm} as obtained from the Caret package \cite{kuhn2008building,caret}} 
+\label{mgm_confusion}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/ROC_mgm.pdf}
+\caption{The ROC curve as obtained from the ROCR package \cite{sing2005rocr} and ggplot2 \cite{ggplot2}. The corresponding $\phi$ value is $0.6$ } 
+\label{ROC_mgm}
+\end{center}
+\end{figure}
+
+\pagebreak
+
 \section{Machine learning algorithms}
 
-\chapter{Conclusion}
+\chapter{Conclusions}
 

### 2021-11-17 Script.R

@@ -265,23 +265,26 @@ conf<-data.frame(lapply(Asteroids_FINAL[,22], as.factor),Predicted_vs_real)
 
 Confusion_matrix_ASTEROIDS<-confusionMatrix(conf$Hazardous,conf$PRED)
 
-fourfoldplot(Confusion_matrix_ASTEROIDS$table,color = c("red","darkgreen"), main = "Mixed Interaction",conf.level = 0)
+fourfoldplot(Confusion_matrix_ASTEROIDS$table,color = c("red","darkgreen"),conf.level = 0)
 
 Confusion_matrix_ASTEROIDS
 
 pred_numeric<-prediction(as.numeric(conf$PRED),as.numeric(conf$Hazardous))
 
-roc_svm.perf <- performance(pred_numeric, measure = "tpr", x.measure = "fpr")
+roc_mgm.perf <- performance(pred_numeric,measure = "tpr", x.measure = "fpr")
 
 phi_Asteroids<-performance(pred_numeric, "phi")
 
 phi_Asteroids@y.values[[1]][2]
 
-plot(roc_svm.perf,cex.lab=1.5,yaxis.cex.axis=1.5,xaxis.cex.axis=1.5)
+pippo<-plot(roc_svm.perf,cex.lab=1.5,yaxis.cex.axis=1.5,xaxis.cex.axis=1.5)
 abline(a=0, b= 1)
 
+df <- data.frame (roc_mgm.perf@x.values,roc_mgm.perf@y.values)
 
+colnames(df)<-c("FPR", "TPR")
 
+ggplot(df,aes(FPR,TPR))+geom_point(color="red")+geom_line(color="red")+ggtitle("ROC CURVE")+theme_bw()+theme(axis.text=element_text(size=16),axis.title=element_text(size=14,face="bold"))
 
 
 

### 2021-11-18 Manuscript/manuscript.tex
PUB

@@
 The confusion matrix obtained is reported in Fig.
-\ref{mgm_confusion}: the ROC curve is given in Fig \ref{ROC_mgm} and it corresponds to a $\phi=0.6$
+\ref{mgm_confusion}: the ROC curve is given in Fig.
+\ref{ROC_mgm} and it corresponds to a $\phi=0.6$.
+We will comment these performances in a few when we will present the performances of other algorithms for the same dataset.
+Before this we would show the graphical models obtained with different implementation of mixed interaction model: we considered the function \textit{mmod()} of gRim package \cite{hojsgaard2012graphical} and the \textit{minforest()} of gRapHD (which uses the minForest method) \cite{de2009high}.
+Both were obtained with a stepwise algorithm.
+Their result are reported in Fig.
+\ref{mmod} and \ref{minforest}.
+We see that in these models the minimum diameter is linked with the hazordous feature, and this is correct, but what is without a physical meaning is that this quantity is also linked to the features connected with the orbital parameters.
+Indeed, while for the magnitude (as for mgm model) a weak relationship can be admitted since this depends also on the distance between the asteroids and the observer and from the orbit, a volume and thus a mass dependence is not acceptable.
+In principle one can set these link as forbidden in the algorithm (blacklist), but in the author view it is preferable a model were the result is obtained without boundaries, instead of a one obtained with a large number of boundaries.
+Thus since a correct model was obtained without the imposition a blacklist, the mgm, we simply reject the mmod and minforest ones in favour of the first one.
+As consequence this is the reason why we report the performances of mgm model only.
 
@@
 
+\begin{landscape}
+\begin{figure}
+\begin{center}
+\includegraphics[width=1.5\textheight]{Figures/mmod.pdf}
+\caption{The graphical mixed interaction model obtained with the grim package \cite{hojsgaard2012graphical}. The plot was obtained with the igraph package for R \cite{igraph} }
+\label{mmod}
+\end{center}
+\end{figure}
+\end{landscape}
+\\
+\begin{landscape}
+\begin{figure}
+\begin{center}
+\includegraphics[width=1.5\textheight]{Figures/minforest.pdf}
+\caption{The graphical mixed interaction model obtained with the gRapHD package \cite{de2009high} with a stepwise algorithm. The plot was obtained with the \cite{qgraph} package }
+\label{minforest}
+\end{center}
+\end{figure}
+\end{landscape}
+
 \pagebreak
@@
 \section{Machine learning algorithms}
+
+Given the mgm model as the one obtained with the probabilistic graphical methods, we would compare its performances with the ones obtained from the following machine learning algorithms: the random-forest (as implemented in the RandomForest package \cite{rfor}), the support vector machines (as implemented in the e1071 package \cite{dimitriadou2008misc}) , the quadratic discriminant analysis (as implemented in the MASS package \cite{MASS} and the logistic regression (as implented in stats package \cite{stats}).
+Their performances are reported in Fig.
+\ref{CF_ML} and \ref{ROC_mgm} as well in the Tab.
+\ref{phi_values}.
+From these comparison we can see that most of these algorithms, expect for the QDA outperform the graphical method mgm or have similar performances as for the logistic algorithm.
+Thus at this point one can ask what is the advantage of using a graphical method instead of a random forest or a SVM since its performances seems lower.
+The answer is the interpretability of the model provided: the random forest, at lest, can provide a variable rank importance as reported in Fig.
+\ref{RF_Importance} (note that apart from the Min Orbit intersection, there is a slight reshuffling in the feature importance with respect to the one established by the MI in fig \ref{Mutual_information}; however apart from this is a black-box as the SVM, the logistic regression and the QDA.
+Differently the probabilistic graph, providing the conditional independences, give to the user a interpretable model whose properties can be also compared, discussed and validated with the theory.
+Thus the model developed in this way allows a more scientific evaluation with respect to the black-box one.
+In the author view this characteristic definitely compensates the lack of predictive power with respect to the Random Forest or the SVM methods.
+
+\begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/RF_confusion.pdf}
+    \subcaption{Random Forest}
+    \vspace{4ex}
+  \end{minipage} 
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/SVM_confusion.pdf}
+    \subcaption{SVM}
+    \vspace{4ex}
+  \end{minipage} \\
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/QDA_confusion.pdf}
+    \subcaption{QDA}
+    \vspace{4ex}
+  \end{minipage}%%
+    \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/Logisic_confusion.pdf}
+    \subcaption{Logistic}
+    \vspace{4ex}
+  \end{minipage}%%
+\caption{Confusion matrices for a selected set of ML algorithms as obtained from the Caret package \cite{kuhn2008building,caret}}
+\label{CF_ML}
+  \end{figure}
+
+  \begin{figure}[ht]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/ROC_RF.pdf}
+    \subcaption{Random Forest}
+    \vspace{4ex}
+  \end{minipage} 
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/ROC_SVM.pdf}
+    \subcaption{SVM}
+    \vspace{4ex}
+  \end{minipage} \\
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/ROC_QDA.pdf}
+    \subcaption{QDA}
+    \vspace{4ex}
+  \end{minipage}%%
+    \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/ROC_logistic.pdf}
+    \subcaption{Logistic}
+    \vspace{4ex}
+  \end{minipage}%%
+\caption{ROC curves for a selected set of ML algorithms. Plot obtained with caret and ggplot2 package \cite{kuhn2008building},ggplot2}}
+\label{ROC_ML}
+  \end{figure}
+
+\begin{table}[]
+\label{phi_values}
+\caption{$\phi$ factor for a selected set of ML algorithms as compared with the mgm}
+\begin{center}
+\begin{tabular}{c|c}
+Algorithm & $\phi$ \\ \hline
+RF        & 0.9876 \\ \hline
+SVM       & 0.7111 \\ \hline
+logistic  & 0.6173 \\ \hline
+mgm       & 0.5997 \\ \hline
+QDA       & 0.5562
+
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/RF_Importance.pdf}
+\caption{Variable importance according to the random forest algorithm as implemented in \cite{rfor} package. Plot obtained with ggplot2 package \cite{ggplot2}}
+\label{RF_Importance}
+\end{center}
+\end{figure}
 

### 2021-11-18 Script.R

@@ -29,6 +29,7 @@ library(e1071)
 library(MASS)
 library(infotheo)
 library(varrank)
+library(minet)
 
 #Data Loading
 
@@ -214,7 +215,7 @@ plot(as(commonedges.asteroids ,"igraph"),vertex.color="orange",vertex.label.dist
 
 bf<-minForest(Asteroids_FINAL_double)
 plot(bf,cex.vert.label=0.6,numIter=6000,col.labels=c("red"),vert.hl=c(21),col.hl=c("blue"),energy=TRUE)
-mbG<-stepw(model=bf,data=Asteroids_FINAL_double,threshold=1)
+mbG<-stepw(model=bf,data=Asteroids_FINAL_double)
 plot(mbG,cex.vert.label=0.6,numIter=6000,col.labels=c("red"),vert.hl=c(21),col.hl=c("blue"),energy=TRUE)
 
 
@@ -224,6 +225,10 @@ plot(aa2)
 
 
 plot(as(aa2 ,"igraph"),vertex.color="red",vertex.label.dist=2,alpha=TRUE,vertex.label.cex=0.5,vertex.size=10,edge.color="black",edge.size=2,layout=layout_with_fr(as(aa2,'igraph')), vertex.label.family='Helvetica')
+
+
+
+
 
 
 #Mixed interaction analysis
@@ -277,14 +282,18 @@ phi_Asteroids<-performance(pred_numeric, "phi")
 
 phi_Asteroids@y.values[[1]][2]
 
-pippo<-plot(roc_svm.perf,cex.lab=1.5,yaxis.cex.axis=1.5,xaxis.cex.axis=1.5)
-abline(a=0, b= 1)
+
 
 df <- data.frame (roc_mgm.perf@x.values,roc_mgm.perf@y.values)
 
 colnames(df)<-c("FPR", "TPR")
 
 ggplot(df,aes(FPR,TPR))+geom_point(color="red")+geom_line(color="red")+ggtitle("ROC CURVE")+theme_bw()+theme(axis.text=element_text(size=16),axis.title=element_text(size=14,face="bold"))
+
+mim <- build.mim(Asteroids_FINAL[,-1])
+graph_mu<-aracne(mim)
+plot(graph_from_graphnel( as( graph_mu ,"graphNEL")),layout=layout_with_fr(as(graph_mu,'igraph')) )
+
 
 
 
[…]

### 2021-11-19 Manuscript/Eq_varie.tex
PUB

Therefore if one sums the two equation, the following differential equation set will be obtained \cite{murray1999solar}:

\begin{equation}
m_{1}\ddot{\textbf{r}}_{1}+m_{2}\ddot{\textbf{r}}_{2}=0
\end{equation}

Its solution can be find with a direct integration \cite{murray1999solar}:

\begin{equation}
m_{1}\dot{\textbf{r}}_{1}+m_{2}\dot{\textbf{r}}_{2}=\textbf{a}
\end{equation}

\begin{equation}
m_{1}\textbf{r}_{1}+m_{2}\textbf{r}_{2}=\textbf{a}t+\textbf{b}
\end{equation}

Using the definition center of mass as \cite{murray1999solar}:

\begin{equation}
\textbf{R}=\frac{m_{1}\textbf{r}_{1}+m_{2}\textbf{r}_{2}}{m_{1}+m_{2}}
\end{equation}

the previous solution can be recast in the following form \cite{murray1999solar}:

\begin{equation}
\textbf{\dot{R}}=\frac{\textbf{a}}{m_{1}+m_{2}}
\end{equation}

\begin{equation}
\textbf{\dot{R}}=\frac{\textbf{a}t+\textbf{b}}{m_{1}+m_{2}}
\end{equation}

Where whe can see that

### 2021-11-19 Manuscript/manuscript.tex
PUB

@@
 \begin{table}[]
-\label{phi_values}
 \caption{$\phi$ factor for a selected set of ML algorithms as compared with the mgm}
@@
 QDA       & 0.5562 
-
 \end{tabular}
 \end{center}
+\label{phi_values}
 \end{table}
@@
 
+In this Appendix is dedicated to review the basics concepts of celestial dynamics, following the Murray approach \cite{murray1999solar}, that are needed to understand the features of the dataset and their connections.
+Lets start by considering two masses $m_{1}$ and $m_{2}$, which in the present case will be respectively the planet Earth and the asteroid.
+Their position will be given, respectively, by two vectors $\textbf{r}_{1}$ and $\textbf{r}_{2}$ considering the origin $O$ bounded in to an inertial space.
+Furthermore we can define also the relative position with the vector $\textbf{r}=\textbf{r}_{2}-\textbf{r}_{1}$.
+Since we suppose that the masses are not interacting with the electromagnetic force, these will be bounded only by the gravitational interaction which is given by the Newton law \cite{murray1999solar}:
+
+\begin{equation}
+\textbf{F}_{1}=\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{1}
+\end{equation}
+
+\begin{equation}
+\textbf{F}_{2}=-\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{2}
+\end{equation}
+
+where $\mathcal{G}$ is universal gravitational constant, and the second equivalence is given by the second Newton law $\textbf{F}=m\cdot \textbf{a}$ in which $\textbf{a}$ is the acceleration calculated as the second derivative of the position vector \textbf{r}.
+Setting $\ddot{\textbf{r}}=\ddot{\textbf{r}}_{2}-\ddot{\textbf{r}}_{1}$ (thus we consider the motion of the second item with respect to the first one) and $\mu=\mathcal{G}(m_{1}+m_{2})$ the following differential equation will be obtained from the previous two ones \cite{murray1999solar}:
+
+\begin{equation}
+\dfrac{d^{2}\textbf{r}}{dt^{2}}+\mu\dfrac{\textbf{r}}{r^{3}}=0
+\end{equation}
+
+It can be seen that the $\textbf{r}$ and $\cdot\textbf{r}$ lies always in the same plane: this is because the product vector $\textbf{r} \times \ddot{ \textbf{r}}=0$, thus if one integrates she will get that the product vector $\textbf{r} \times \textbf{r}=\textbf{h}$ where \textbf{h} is a constant vector.
+Furthermore the problem can be simplified by using polar coordinates  $\hat{\textbf{r}}$ and $\hat{\boldsymbol{\theta}}$.
+Indeed since the position, speed and acceleration in polar coordinates have the following form \cite{murray1999solar}:
+
+\begin{equation}
+\textbf{r}=r\hat{\textbf{r}}
+\end{equation}
+
+\begin{equation}
+\dot{\textbf{r}}=\dot{r}\hat{\textbf{r}}+r\dot{\theta}\hat{\boldsymbol{\theta}}
+\end{equation}
+
+\begin{equation}
+\ddot{\textbf{r}}=\left(\ddot{r}-r\dot{\theta}^{2}\right)\hat{\textbf{r}}+\left[\dfrac{1}{r}\frac{d}{dt}\left(r^{2}\dot{\theta}\right)\right]\hat{\boldsymbol{\theta}}
+\end{equation}
+
+Using these coordinates in the product vector between the speed and the position the following equation will be obtained \cite{murray1999solar}:
+
+\begin{equation}
+\textbf{h}=r^{2}\dot{\theta}\hat{\textbf{z}}
+\end{equation}
+
+where $\textbf{z}$ is a vector perpendicural to the plane, whose module is equal to
+
+\begin{equation}
+h=r^{2}\dot{\theta}
+\end{equation}
+
+If we consider the motion of the body $m_{2}$ in the time interval $\delta t$ we have that the area $\delta A$ illustrated in Fig will be \cite{murray1999solar}:
+
+\begin{equation}
+\delta A \approx \dfrac{1}{2} r(r+dr)\sin(\delta\theta) \approx  \dfrac{1}{2} r^{2}\delta\theta
+\end{equation}
+
+where the Taylor expansion at first order was used. Therfore
+
+\begin{equation}
+\dfrac{dA}{dt}=\dfrac{1}{2}r^{2}\dfrac{d\theta}{dt}=\dfrac{1}{2}h
+\end{equation}
+
+but we know that $h$ is constant, therefore the first derivative, is constant. This is the $2^{th}$ Kepler law
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Area_dynamics.png}
+\caption{The portion of area $\delta A$ obtained when the position vector moves with an angle $\delta\theta$. Image taken from \cite{murray1999solar}}
+\label{RF_Importance}
+\end{center}
+\end{figure}
+
 \bibliographystyle{unsrt}

### 2021-11-19 Manuscript/manuscript.tex
PUB

@@
 \dot{\textbf{r}}=\dot{r}\hat{\textbf{r}}+r\dot{\theta}\hat{\boldsymbol{\theta}}
+\label{eq_dyn_nop}
 \end{equation}
@@
 
-If we consider the motion of the body $m_{2}$ in the time interval $\delta t$ we have that the area $\delta A$ illustrated in Fig will be \cite{murray1999solar}:
+If we consider the motion of the body $m_{2}$ in the time interval $\delta t$ we have that the area $\delta A$ illustrated in Fig \ref{Area_dynamics} will be \cite{murray1999solar}:
 
@@
 
-where the Taylor expansion at first order was used. Therfore
+where the Taylor expansion at first order was used. Therfore  \cite{murray1999solar}:
 
@@
 
-but we know that $h$ is constant, therefore the first derivative, is constant. This is the $2^{th}$ Kepler law
+but we know that $h$ is constant, therefore the first derivative, is constant. This is the $2^{th}$ Kepler law. The Eq. \ref{eq_dyn_nop} in polar coordinates  \cite{murray1999solar}:
+
+\begin{equation}
+\ddot{r}-r\dot{\theta}^{2}=-\frac{\mu}{r^{2}}
+\end{equation}
+
+This differential equation can be rewritten as an armonic oscillator with the  substitutions $u=\dfrac{1}{r}$ $h=r^{2}\dot{\theta}$ \cite{murray1999solar}:
+
+\begin{equation}
+\dot{r}=-\frac{1}{u}\dfrac{du}{d\theta}\dot{\theta}=-h\frac{du}{d\theta}
+\end{equation}
+
+\begin{equation}
+\ddot{r}=-h\dfrac{d^{2}u}{d\theta^{2}}\dot{\theta}=-h^{2}u^{2}\frac{d^{2}u}{d\theta^{2}}
+\end{equation}
+
+\begin{equation}
+\dfrac{d^{2}u}{d\theta^{2}}+u=\frac{\mu}{h^{2}}
+\end{equation}
+
+\begin{equation}
+u=\frac{\mu}{h^{2}}\left[1+e\cos(\theta-\phi\right]
+\end{equation}
+
+where the integration constants e and $\phi$ are respectively the amplitude and the phase. Therefore we have \cite{murray1999solar}:
+
+\begin{equation}
+r=\dfrac{p}{1+e\cos(\theta-\phi}
+\end{equation}
+
+In this form we can recognize in e the eccentricity and p is the semilatus rectum \cite{murray1999solar}:
+
+\begin{equation}
+p=\frac{h^{2}}{\mu}
+\end{equation}
+
+Depening on the eccentricity we have four possibile conics
+
+\begin{itemize}
+\item circle:  e=0 \quad $p=a$
+\item ellipse: 0<e<1 \quad $p=a$
+\item parabola: e=1 \quad $p=2q$
+\item hyperbola: e>1 \quad $p=a(e^{2}-1)$
+\end{itemize}
+
+in which a is the semi-major axis of the conic. The shape of these orbits is reported in Fig. \ref{Conics}.
 
@@
 \caption{The portion of area $\delta A$ obtained when the position vector moves with an angle $\delta\theta$. Image taken from \cite{murray1999solar}}
-\label{RF_Importance}
+\label{Area_dynamics}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Conics.png}
+\caption{The four possible orbits as obtained from a sectionof a cone. Image taken from \cite{murray1999solar}}
+\label{Conics}
 \end{center}

### 2021-11-22 Manuscript/manuscript.tex
PUb

@@
 
-In this Appendix is dedicated to review the basics concepts of celestial dynamics, following the Murray approach \cite{murray1999solar}, that are needed to understand the features of the dataset and their connections.
+This Appendix is dedicated to review the basics concepts of celestial dynamics, following the Murray approach \cite{murray1999solar}, that are needed to understand the features of the dataset and their connections.
 Lets start by considering two masses $m_{1}$ and $m_{2}$, which in the present case will be respectively the planet Earth and the asteroid.
@@
 
-Depening on the eccentricity we have four possibile conics
+Depening on the eccentricity we have four possible conics \cite{murray1999solar}:
 
 \begin{itemize}
-\item circle:  e=0 \quad $p=a$
-\item ellipse: 0<e<1 \quad $p=a$
-\item parabola: e=1 \quad $p=2q$
-\item hyperbola: e>1 \quad $p=a(e^{2}-1)$
+\item circle:  $e=0$ \quad $p=a$
+\item ellipse: $0<e<1$ \quad $p=a$
+\item parabola: $e=1$ \quad $p=2q$
+\item hyperbola: $e>1$ \quad $p=a(e^{2}-1)$
 \end{itemize}
 
-in which a is the semi-major axis of the conic. The shape of these orbits is reported in Fig. \ref{Conics}.
+in which $a$ is the semi-major axis of the conic.
+The shape of these orbits is reported in Fig.
+\ref{Elliptical_orbit}.
+All the asteroids considered here an eccentricity $0<e<1$, therefore they have an elliptical orbits in which the earth lies in to one of the two focal point.
+It is worth nothing that this is the first Kepler law.
+Looking to Fig.
+we can define the point of minimum distance between $m_{1}$ and the orbiting body as the pericentre or perihelion, and the maximum distance as the apocentre or the aphelion.
+The semi-major axis, here denoted as $b$ on the other side is defined as the distance between the pericentre and the apocentre.
+Using the following identity \cite{murray1999solar}:
+
+\begin{equation}
+b^{2}=a^{2}(1-e^{2})
+\end{equation}
+
+we get \cite{murray1999solar}:
+
+\begin{equation}
+r=\frac{a(1-e^{2}}{1+e\cdot cos(\theta-\phi}
+\label{eq-mot}
+\end{equation}
+
+Furthermore the third Kepler law, can be quickly obtained considering the area swept in one orbital period T (the time needed to complete a full round of the orbit) $A=\pi ab$.
+Since we know that this area is equal to $hT/2$ and $h^{2}=\mu a(1-e^{2})$ \cite{murray1999solar}:
+
+\begin{equation}
+T^{2}=\dfrac{4\pi^{2}}{\mu}a^{3}
+\end{equation}
+
+If we have two bodies, of masses $m$ and $m'$, that orbit around the Earth $m_{c}$, we can use the previous equation to obtain \cite{murray1999solar}:
+
+\begin{equation}
+\frac{m_{c}+m}{m_{c}+m'}=\left(\frac{a}{a'}\right)\left(\frac{T'}{T}\right)^{2}
+\end{equation}
+
+But since $m$,$m'<<m_{c}$ \cite{murray1999solar}:
+
+\begin{equation}
+(a/a')^{3}\approx(T/T')^{2}
+\end{equation}
+
+Therefore \cite{murray1999solar}:
+
+\cite{murray1999solar}:
+
+We see that the two orbital parameters $T'$ and $a'$ are independent with respect to orbiting mass.
+This statement can be extended to the other orbital parameters by considering the previous approximation $m$,$m'<<m_{c}$.
+This is why we expect that the mass/volume of the asteoroid can not be conditionally dependant to the orbital parameters.
+It is useful to define also the mean motion (feature that is also present in the asteroids dataset) as \cite{murray1999solar}:
+
+\begin{equation}
+n=\frac{2\pi}{T}
+\end{equation}
+
+Therefore \cite{murray1999solar}:
+
+\begin{equation}
+\mu=n^{2}a^{3}
+\end{equation}
+
+\begin{equation}
+h=na^{2}\sqrt{1-e^{2}}=\sqrt{\mu a(1-e^{2})}
+\label{eq_h}
+\end{equation}
+
+From which we can see that the angular velocity $\ddot{f}$ is function of the longitude.
+We are now going more in deep with this statement.
+Lets come back to the Eq.
+\ref{eq-mot}, this can be rewritten as \cite{murray1999solar}:
+
+\begin{equation}
+\dot{\textbf{r}}\cdot\ddot{\textbf{r}}+\mu\dfrac{\dot{r}}{r^{2}}=0
+\end{equation}
+
+whose integration gives \cite{murray1999solar}:
+
+\begin{equation}
+\frac{1}{2}v^{2}-\frac{\mu}{r}=C
+\label{energy_cons}
+\end{equation}
+
+In which $v^{2}=\dot{\textbf{r}}\cdot\dot{\textbf{r}}$, and C the integration costant.
+This expression express the energy conservation: on the left side we have the vis-via term (basically the kinetic energy without the mass), on the right side the potential energy (rescaled with the reduced mass).
+Lets come back to Eq.
+\ref{eq-mot}, and make the following substitution $f=\theta-\phi$, which is called the true anomaly.
+If we differentiate it we will obtain \cite{murray1999solar}:
+
+\begin{equation}
+\dot{r}=\frac{r\dot{f}e\sin f}{1+e\cos f}
+\end{equation}
+
+Remembering the the definition of $h=r^{2}\ddot{f}$, from Eq. \ref{eq_h} we have  \cite{murray1999solar}:
+
+\begin{equation}
+\dot{r}=\frac{na}{\sqrt{1-e^{2}}}e\sin f
+\end{equation}
+
+\begin{equation}
+r\dot{f}=\frac{na}{\sqrt{1-e^{2}}}\left(1+e\cos f\right)
+\end{equation}
+
+Therefore \cite{murray1999solar}:
+
+\begin{equation}
+\begin{split}
+v^{2}&=\dfrac{n^{2}a^{2}}{1-e^{2}}\left(1+2e\cos f +e^{2}\right)= \\
+&=\dfrac{n^{2}a^{2}}{1-e^{2}}\left(\dfrac{2a(1-e^{2})}{r}-(1-e^{2})\right)
+\end{split}
+\end{equation}
+
+\begin{equation}
+v^{2}=\mu\left(\frac{2}{r}-\dfrac{1}{a}\right)
+\end{equation}
+
+From which we get that the velocity of the asteroid is maximum at the perihelion, and minimum at the aphelion. Their values that are equal respectively to \cite{murray1999solar}:
+
+\begin{equation}
+v_{perihelion}=na\sqrt{\dfrac{1+e}{1-e}}
+\end{equation}
+
+\begin{equation}
+v_{aphelion}=na\sqrt{\dfrac{1-e}{1+e}}
+\end{equation}
+
+Another quantity that is contained in the asteroids dataset, and is useful to describe their orbits is the mean anomaly. This is defined as \cite{murray1999solar}:
+
+\begin{equation}
+M=n(t-\tau)
+\end{equation}
+
+where $\tau$, the time of pericentre passage, increases linearly with time at a costant rate equal to the mean motion.
+Furthermore is bounded by the following relations for the perihelion and aphelion \cite{murray1999solar}:
+
+\begin{itemize}
+\item $M=f=0$\quad$t=\tau$\quad Perihelion
+\item $M=f=\pi$\quad$t=\tau+T/2$ \quad Aphelion
+\end{itemize}
+
+Such boundaries should be intended as periodic for multiple of the orbital period T.
+The geometrical interpretation of the angle associated to the mean anomaly is given in Fig.
+\ref{Mean_anomaly}.
+It can be proven (see \cite{murray1999solar}) that the value of this angle, which describe the position of the orbiting item, is given by the following expression, know as Kepler equation \cite{murray1999solar}:
+
+\begin{equation}
+M=E-e\sin E
+\end{equation}
+
+Finally as move on to space, other two angles are necessary for the description of an orbit: these are shown in Fig.
+\ref{Inclination} and are the inclination of the orbit (I) and the longitude of the ascending node $\Omega$.
+Given this quantities the Tisserard invariant can be calculated as \cite{murray1999solar}:
+
+\begin{equation}
+T_{P}=\frac{a_{p}}{a}+2\cos I\sqrt{\dfrac{a}{a_{p}}(1-e^{2})}
+\end{equation}
+
+If Jupiter is considered as perturbing body, we have the Juptier Tisserard Invariant.
+The underlying reason for this choice is to distinguish the Jupiter family comets ($2<T_{j}<3$) and the asteroids $T_{j}<2$.
 
@@
 
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Elliptical_orbit.png}
+\caption{Main features of an elliptical orbit. Image taken from \cite{murray1999solar}}
+\label{Elliptical_orbit}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Mean_anomaly.png}
+\caption{The geometrical interpretation of mean anomaly: on the left panel a) is reported how the circumscribed circle should be draw, while on the right panel b) it is shown how the angle associated with the mean anomaly should be interpreted and its relation with the true anomaly angle f.
+Image taken from \cite{murray1999solar}}
+\label{Mean_anomaly}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Inclination.png}
+\caption{The parameters that are necessary for the description of an orbit in three dimension. Image taken from \cite{murray1999solar}}
+\label{Inclination}
+\end{center}
+\end{figure}
+
 \bibliographystyle{unsrt}

### 2021-11-22 Manuscript/manuscript.tex
PUB

@@
 \begin{equation}
-u=\frac{\mu}{h^{2}}\left[1+e\cos(\theta-\phi\right]
+u=\frac{\mu}{h^{2}}\left[1+e\cos(\theta-\phi)\right]
 \end{equation}
@@
 \begin{equation}
-r=\dfrac{p}{1+e\cos(\theta-\phi}
-\end{equation}
-
-In this form we can recognize in e the eccentricity and p is the semilatus rectum \cite{murray1999solar}:
+r=\dfrac{p}{1+e\cos(\theta-\phi)}
+\end{equation}
+
+In this form we can recognize in $e$ the \textcolor{red}{eccentricity} and $p$ is the semilatus rectum \cite{murray1999solar}:
 
@@
 Looking to Fig.
-we can define the point of minimum distance between $m_{1}$ and the orbiting body as the pericentre or perihelion, and the maximum distance as the apocentre or the aphelion.
-The semi-major axis, here denoted as $b$ on the other side is defined as the distance between the pericentre and the apocentre.
+we can define the point of minimum distance between $m_{1}$ and the orbiting body as the pericentre or \textcolor{red}{perihelion}, and the maximum distance as the apocentre or the aphelion.
+The \textcolor{red}{semi-major axis}, here denoted as $b$ on the other side is defined as the distance between the pericentre and the apocentre.
 Using the following identity \cite{murray1999solar}:
@@
 
-Furthermore the third Kepler law, can be quickly obtained considering the area swept in one orbital period T (the time needed to complete a full round of the orbit) $A=\pi ab$.
+Furthermore the third Kepler law, can be quickly obtained considering the area swept in one \textcolor{red}{orbital period} T (the time needed to complete a full round of the orbit) $A=\pi ab$.
 Since we know that this area is equal to $hT/2$ and $h^{2}=\mu a(1-e^{2})$ \cite{murray1999solar}:
@@
 This is why we expect that the mass/volume of the asteoroid can not be conditionally dependant to the orbital parameters.
-It is useful to define also the mean motion (feature that is also present in the asteroids dataset) as \cite{murray1999solar}:
+It is useful to define also the \textcolor{red}{mean motion} (feature that is also present in the asteroids dataset) as \cite{murray1999solar}:
 
@@
 In which $v^{2}=\dot{\textbf{r}}\cdot\dot{\textbf{r}}$, and C the integration costant.
-This expression express the energy conservation: on the left side we have the vis-via term (basically the kinetic energy without the mass), on the right side the potential energy (rescaled with the reduced mass).
+This expression express the energy conservation: on the left side we have the \textit{vis-viva} term (basically the kinetic energy without the mass), on the right side the potential energy (rescaled with $\mu$).
 Lets come back to Eq.
@@
 
-Another quantity that is contained in the asteroids dataset, and is useful to describe their orbits is the mean anomaly. This is defined as \cite{murray1999solar}:
+Another quantity that is contained in the asteroids dataset, and is useful to describe their orbits is the \textcolor{red}{mean anomaly}. This is defined as \cite{murray1999solar}:
 
@@
 Finally as move on to space, other two angles are necessary for the description of an orbit: these are shown in Fig.
-\ref{Inclination} and are the inclination of the orbit (I) and the longitude of the ascending node $\Omega$.
+\ref{Inclination} and are the \textcolor{red}{inclination} of the orbit (I) and the \textcolor{red}{longitude of the ascending node} $\Omega$.
 Given this quantities the Tisserard invariant can be calculated as \cite{murray1999solar}:
@@
 
-If Jupiter is considered as perturbing body, we have the Juptier Tisserard Invariant.
+If Jupiter is considered as perturbing body, we have the \textcolor{red}{Juptier Tisserard Invariant}.
 The underlying reason for this choice is to distinguish the Jupiter family comets ($2<T_{j}<3$) and the asteroids $T_{j}<2$.

### 2021-11-23 Manuscript/manuscript.tex
PUB

@@
 \cite{russell2010artificial}.
-Furthermore we will provide also a rapid overview of the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches ,  since we used some of its concepts in the preliminary analysis of the dataset.
+Furthermore we will provide also a rapid overview of the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches,  since we used some of its concepts in the preliminary analysis of the dataset.
+Finally we concluded this chapter with a rapid overview about the concept of algorithm intepretability, which together with the accuracy and the confusion matrix, is necessary for the assessment of a ML performances.
 On the other side the concepts related to the celestial mechanics here used will be described, following the Murray approach \cite{murray1999solar} into the Appendix A
@@
 
+\section{Algorithm intepretability}
+
+Beside the accuracy and other features that are connected with the confusion matrix, another important feature is the algorithm intepretability.
+This feature, at a first instance, is connected on how much the algorithm provide explanation to a human on how it choose a forecast instead of another one, or in other word how much the algorithm is not a blackbox.
+Apart from all legal problems connected with an algorithm that is a blackbox\footnote{Think for instance the conditions imposed by the GDPR (e.g the right to explanation)}, in the author opinion results provided by such algorithm cannot be considered scientific, since the scientific method also require to provide an explanation and not simply almost correct forecast.
+Following the \cite{tarski1953undecidable} argument, which is root on mathematical logic, it can be said that the a formal theory T into S if and only if S is able to proof the theorem of T in its own language.
+On the other side we would also that an algorithm, as an explanation, is complete: we expect that it will be also able to make correct forecasts.
+However as shown by \cite{doshi2017towards} and by \cite{gilpin2018explaining}  the intepretability of an algorithm is linked with its incompleteness.
+This is something that is nested both in formal system as algorithms or scientific theories.
+For instance lets consider the fall of object: at first glance one can consider the only the items on Earth, in this case the acceleration is constant and it is a very simple theory; but as one moves to consider also the interaction within planets a much more complicate law should be considered (for which the previous case is a particular one).
+The first one is largely explainable but not poorly complete, the second one more difficult to explain but more complete.
+Therefore, in principle, the evaluation of an algorithm performances should be considered not only in to one dimension, the accuracy, but also by considering its intepretability as shown in Fig.
+\ref{ML_intepretability}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/ML_intepretability.png}
+\caption{Different ML algorithms classified by their interpretability and by their accuracy. Image taken from \cite{ml_interpretability}}
+\label{ML_intepretability}
+\end{center}
+\end{figure}
+
 \chapter{Dataset description}
@@
 Thus the features obtained were reduced to 22.
-Their enumeration and their description will be postponed to Appendix A, since a description of them cannot be disjoint with celestial mechanics concepts.
+Their description will be postponed to Appendix A, since a description of them cannot be disjoint with celestial mechanics concepts.
+Their enumeration is provided in Tab.
+\ref{tab_features}.
 Since only the 16 \% of the asteroids were hazardous, we considered to reduce the number of the non-hazardous in order to improve the portion of the hazardous ones.
@@
 Furthermore, concerning the continuous measures, the dataset was standardised and demeaned.
+
+\begin{table}[]
+\caption{The features used for the present analysis. The unit of measure are reported in the original dataset \cite{kaggle_dataset}.}
+\begin{center}
+\begin{tabular}{c|c}
+\hline
+\textbf{Features}             & \textbf{Type}        \\ \hline
+Neo Reference ID              & not used             \\ \hline
+Absolute Magnitude            & Continuous           \\ \hline
+Est Dia in KM (min)           & Continuous           \\ \hline
+Est Dia in KM (max)           & Continuous           \\ \hline
+Close Approach Date           & Continuous           \\ \hline
+Epoch Date Close Approach     & Continuous           \\ \hline
+Relative\_Velocity            & Continuous           \\ \hline
+Miss\_Dist                    & Continuous           \\ \hline
+Min\_Orbit\_Intersection      & Continuous           \\ \hline
+Jupiter\_Tisserand\_Invariant & Continuous           \\ \hline
+Epoch\_Osculation             & Continuous           \\ \hline
+Eccentricity                  & Continuous           \\ \hline
+Semi Major Axis               & Continuous           \\ \hline
+Inclination                   & Continuous           \\ \hline
+Asc Node Longitude            & Continuous           \\ \hline
+Orbital Period                & Continuous           \\ \hline
+Perihelion Distance           & Continuous           \\ \hline
+Perihelion Arg                & Continuous           \\ \hline
+Perihelion Time               & Continuous           \\ \hline
+Mean\_Anomaly                 & Continuous           \\ \hline
+Mean\_Motion                  & Continuous           \\ \hline
+Hazardous                     & Categorical (Binary)
+\end{tabular}
+\end{center}
+\label{tab_features}
+\end{table}
 
@@
 
-\chapter{Appendix A: Concepts of celestial mechanics}
+\chapter{Appendix A: Concepts of astronomy}
 
 This Appendix is dedicated to review the basics concepts of celestial dynamics, following the Murray approach \cite{murray1999solar}, that are needed to understand the features of the dataset and their connections.
+The first part will be dedicatated to the features that are connected with the orbital parameters, while the second one to the parameters that are connected with the observation of the asteroids
+
+\section{Celestial mechanics}
+
 Lets start by considering two masses $m_{1}$ and $m_{2}$, which in the present case will be respectively the planet Earth and the asteroid.
@@
 
-It can be seen that the $\textbf{r}$ and $\cdot\textbf{r}$ lies always in the same plane: this is because the product vector $\textbf{r} \times \ddot{ \textbf{r}}=0$, thus if one integrates she will get that the product vector $\textbf{r} \times \textbf{r}=\textbf{h}$ where \textbf{h} is a constant vector.
+It can be seen that the $\textbf{r}$ and $\dot{\textbf{r}}$ lies always in the same plane: this is because the product vector $\textbf{r} \times \ddot{ \textbf{r}}=0$, thus if one integrates she will get that the product vector $\textbf{r} \times \textbf{r}=\textbf{h}$ where \textbf{h} is a constant vector.
 Furthermore the problem can be simplified by using polar coordinates  $\hat{\textbf{r}}$ and $\hat{\boldsymbol{\theta}}$.
-Indeed since the position, speed and acceleration in polar coordinates have the following form \cite{murray1999solar}:
+In this case speed and acceleration have the following form \cite{murray1999solar}:
 
@@
 
-Using these coordinates in the product vector between the speed and the position the following equation will be obtained \cite{murray1999solar}:
+Therefore the product vector between the speed and the position will have the following form \cite{murray1999solar}:
 
@@
 
-where $\textbf{z}$ is a vector perpendicural to the plane, whose module is equal to
+where $\textbf{z}$ is a vector perpendicural to the plane, whose module is equal to \cite{murray1999solar}:
 
@@
 
-but we know that $h$ is constant, therefore the first derivative, is constant. This is the $2^{th}$ Kepler law. The Eq. \ref{eq_dyn_nop} in polar coordinates  \cite{murray1999solar}:
+but we know that $h$ is constant, therefore the first derivative, is constant.
+This is the $2^{th}$ Kepler law.
+Lets come back to the Eq.
+\ref{eq_dyn_nop}, if one recast it in polar coordinates, the following form will be obtained \cite{murray1999solar}:
 
@@
 Looking to Fig.
-we can define the point of minimum distance between $m_{1}$ and the orbiting body as the pericentre or \textcolor{red}{perihelion}, and the maximum distance as the apocentre or the aphelion.
+\ref{Elliptical_orbit} we can define the point of minimum distance between $m_{1}$ and the orbiting body as the pericentre or \textcolor{red}{perihelion}, and the maximum distance as the apocentre or the aphelion.
 The \textcolor{red}{semi-major axis}, here denoted as $b$ on the other side is defined as the distance between the pericentre and the apocentre.
@@
 
+\section{Observation}
+
+The luminosity of an asteroid can be quantified with the concepts of magnitude.
+It is worth pointing out that asteroids have no intrinsic luminosity, but instead reflects the radiation, usually not uniformly since they have no atmosphere, of stars or other celestial bodies with intrinsic luminosity.
+First of all, before defining the concept of magnitude is useful to provide the concept of radiation flux.
+In our case it can be though as the number of photons \footnote{One can think, at first approximation, photons as packet of energy associated to the emitted light.
+A more formal and complete description can be found in \cite{feynman2018feynman} } that moves across a sphere centred on the light source and with radius equal to the observer distance \cite{burbine2016asteroids}:
+
+\begin{equation}
+\Phi=\frac{L}{4\pi r^{2}}
+\end{equation}
+
+As one can note there is a dependence from $1/r^{2}$ which is given by the fact that the electromagnetic radiation, as for gravitational force, is spherically symmetric\footnote{A formal argument of this point can be found in \cite{zee2013einstein}}.
+On this basis, it is possible to define the relative magnitude as \cite{burbine2016asteroids}:
+
+\begin{equation}
+m=-2.5\log_{10}\Phi+C
+\end{equation}
+
+This definition is useful for the evaluation for the comparison between two light sources (e.g. two stars), since in this case we have \cite{burbine2016asteroids}:
+
+\begin{equation}
+m_{1}-m_{2}=-2.5\log_{10}\frac{\Phi_{1}}{\Phi_{2}}
+\end{equation}
+
+Finally the \textcolor{red}{Absolute magnitude} $M$ can be defined as the magnitude that the item will have if it is put at 1 AU\footnote{Astronomic Units, the distance between Earth and Sun: 1.49\cdot 10^{11} m} or 10 parsec \footnote{3,08\cdot 10^{16} m.
+The physical meaning of this measure can be found in \cite{burbine2016asteroids}} from the observer.
+Therefore  \cite{burbine2016asteroids}:
+
+\begin{equation}
+M-m=-2.5\log_{10}\frac{\Phi\cdot d^{2} }{\Phi\cdot 10^{2}}
+\end{equation}
+
+where $d$ is the light source distance in parsec. Rearranging the terms we have \cite{burbine2016asteroids}:
+
+\begin{equation}
+M-m=-5\log_{10}\frac{ d^{2} }{10}+C
+\end{equation}
+
+\begin{equation}
+M=m+5-5log_{10}d
+\end{equation}
+
 \begin{figure}[h]

### 2021-11-24 Manuscript/manuscript.tex
PU

@@
 
-Finally the \textcolor{red}{Absolute magnitude} $M$ can be defined as the magnitude that the item will have if it is put at 1 AU\footnote{Astronomic Units, the distance between Earth and Sun: 1.49\cdot 10^{11} m} or 10 parsec \footnote{3,08\cdot 10^{16} m.
+Finally the \textcolor{red}{Absolute magnitude} $M$ can be defined as the magnitude that the item will have if it is put at 1 AU\footnote{Astronomic Units, the distance between Earth and Sun: 1.49\cdot 10^{11} m} or 10 parsec \footnote{$3,08\cdot 10^{16}$ m.
 The physical meaning of this measure can be found in \cite{burbine2016asteroids}} from the observer.
@@
 
+\section{Classification}
+
+The solar system is composed not only by planets and the Sun, but also by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}.
+In particular if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO).
+Such items are of three type: comets, meteoroids and asteroids.
+The first one are icy bodies that moving near the sun melts and release gases giving coloured tails.
+Meteroids are small \footnote{With respect to asteroids} rocky/metallic items with a diameter less than 1 meter.
+Asteroids, on the other side, are items with a diameter larger than 1 m \cite{burbine2016asteroids}.
+The first two items are excluded from the present analysis.
+The near asteroids are classified, as shown in Fig. \ref{neo_orbit_types}, according to their semi-major axis, perihelion distance, and aphelion distance \cite{nasa_classification}:
+
+\begin{itemize}
+\item \textbf{Atiras}
+\item \textbf{Atens}
+\item \textbf{Apollos}
+\item \textbf{Amors}
+\end{itemize}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/neo_orbit_types.jpg}
+\caption{NASA classification of NEO asteroids accompained by the legend for the parameters a,q and Q. Image taken from \cite{nasa_classification}}
+\label{neo_orbit_types}
+\end{center}
+\end{figure}
+
 \bibliographystyle{unsrt}

### 2021-11-24 Manuscript/manuscript.tex
PUB

@@
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1\textwidth]{Figures/Area_dynamics.png}
-\caption{The portion of area $\delta A$ obtained when the position vector moves with an angle $\delta\theta$. Image taken from \cite{murray1999solar}}
-\label{Area_dynamics}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1\textwidth]{Figures/Conics.png}
-\caption{The four possible orbits as obtained from a sectionof a cone. Image taken from \cite{murray1999solar}}
-\label{Conics}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1\textwidth]{Figures/Elliptical_orbit.png}
-\caption{Main features of an elliptical orbit. Image taken from \cite{murray1999solar}}
-\label{Elliptical_orbit}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1\textwidth]{Figures/Mean_anomaly.png}
-\caption{The geometrical interpretation of mean anomaly: on the left panel a) is reported how the circumscribed circle should be draw, while on the right panel b) it is shown how the angle associated with the mean anomaly should be interpreted and its relation with the true anomaly angle f.
-Image taken from \cite{murray1999solar}}
-\label{Mean_anomaly}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1\textwidth]{Figures/Inclination.png}
-\caption{The parameters that are necessary for the description of an orbit in three dimension. Image taken from \cite{murray1999solar}}
-\label{Inclination}
-\end{center}
-\end{figure}
-
 \section{Classification}
@@
 The first two items are excluded from the present analysis.
-The near asteroids are classified, as shown in Fig. \ref{neo_orbit_types}, according to their semi-major axis, perihelion distance, and aphelion distance \cite{nasa_classification}:
+The near asteroids are classified, as shown in Fig.
+\ref{neo_orbit_types}, according to their semi-major axis (a), perihelion distance (q), and aphelion distance (Q).
+Here we report the Center for Near-Earth Object Studies (CNEOS) definition \cite{nasa_classification}:
 
 \begin{itemize}
-\item \textbf{Atiras}
-\item \textbf{Atens}
-\item \textbf{Apollos}
-\item \textbf{Amors}
+\item \textbf{Atiras} $a < 1.0$ au $Q < 0.983$ au\quad \textit{NEAs whose orbits are contained entirely with the orbit of the Earth (named after asteroid 163693 Atira)}
+\item \textbf{Atens} $a < 1.0$ au $Q > 0.983$ au \quad \textit{Earth-crossing NEAs with semi-major axes smaller than Earth's (named after asteroid 2062 Aten)}
+\item \textbf{Apollos} $a>1.0$ au $q<1.017$ au \quad  \textit{Earth-crossing NEAs with semi-major axes larger than Earth's (named after asteroid 1862 Apollo) }
+\item \textbf{Amors} $a>1.0$ au $1.017<q<1.3$ au \textit{Earth-approaching NEAs with orbits exterior to Earth's but interior to Mars' (named after asteroid 1221 Amor)}
 \end{itemize}
+
+Beside this classification, there is the one which involve the hazardousness of an asteroid \cite{nasa_classification}:
+
+\begin{itemize}
+\item \textbf{Potentially Hazardous Asteroids}: MOID $\leq 0.05$ au $M \leq22.0$ \textit{NEAs whose Minimum Orbit Intersection Distance (MOID) with the Earth is 0.05 au or less and whose absolute magnitude (M) is 22.0 or brighter}
+\end{itemize}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Area_dynamics.png}
+\caption{The portion of area $\delta A$ obtained when the position vector moves with an angle $\delta\theta$. Image taken from \cite{murray1999solar}}
+\label{Area_dynamics}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Conics.png}
+\caption{The four possible orbits as obtained from a sectionof a cone. Image taken from \cite{murray1999solar}}
+\label{Conics}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Elliptical_orbit.png}
+\caption{Main features of an elliptical orbit. Image taken from \cite{murray1999solar}}
+\label{Elliptical_orbit}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Mean_anomaly.png}
+\caption{The geometrical interpretation of mean anomaly: on the left panel a) is reported how the circumscribed circle should be draw, while on the right panel b) it is shown how the angle associated with the mean anomaly should be interpreted and its relation with the true anomaly angle f.
+Image taken from \cite{murray1999solar}}
+\label{Mean_anomaly}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Inclination.png}
+\caption{The parameters that are necessary for the description of an orbit in three dimension. Image taken from \cite{murray1999solar}}
+\label{Inclination}
+\end{center}
+\end{figure}
 

### 2021-11-24 Manuscript/manuscript.tex
PUB

@@
 
+\newpage
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{Figures/vesta_asteroid.jpg}
+\captionsetup{labelformat=empty}
+\caption{A NASA image of asteroid Vesta as taken by the spacecraft Dawn.
+Vesta is one of the biggest asteroid in the asteroid belt of solar system (its volume is equal to $7,5\cdot10^{7}$ km$^{3}$.
+Image taken from \cite{vesta_source}}
+\end{center}
+\end{figure}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{Figures/Tunguska.png}
+\captionsetup{labelformat=empty}
+\caption{A picture provided by the Soviet Academy of Science in 1927 of Podkamennaya Tunguska.
+This place was hit, in 1908, by an asteroid of 50-60 m.
+Such asteroid flattened almost 80 million trees over an area of 2,150 km$^{2}$.
+The explosion intensity was near to 12 megatons: for reference the modern US nuclear bomb are in the range of 0.3 kilotons to 1.2 megatons.
+The Hiroshima bomb was nearly to 15 kilotons.
+Image taken from \cite{Tunguska_source}}
+\end{center}
+\end{figure}
 \newpage
@@
 
-This Appendix is dedicated to review the basics concepts of celestial dynamics, following the Murray approach \cite{murray1999solar}, that are needed to understand the features of the dataset and their connections.
-The first part will be dedicatated to the features that are connected with the orbital parameters, while the second one to the parameters that are connected with the observation of the asteroids
+This Appendix is dedicated to review the basics concepts of astronomy that are needed to understand the features of the dataset and the theoretical framework in which these are involved.
+First the basic concepts of celestial mechanics will be recap following the Murray approach \cite{murray1999solar}: these will provide a description of the orbital parameters of the dataset.
+Then the concepts connected to the observation of the asteroids will be reviewed following the Burbine textboox \cite{burbine2016asteroids}.
+Finally, using the previous definition, the classification of the asteroids and in particular the definition of their hazardousness will be provided following the Center for Near-Earth Object Studies (CNEOS) statements \cite{nasa_classification}.
 
@@
 
-The solar system is composed not only by planets and the Sun, but also by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}.
+The solar system is composed not only by planets and the Sun, but also, as shown in Fig.
+\ref{Orbit_asteorids}, by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}.
 In particular if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO).
@@
 The first two items are excluded from the present analysis.
-The near asteroids are classified, as shown in Fig.
+The near asteroids, which are mostly located on the sites reported in Fig.
+\ref{heic1715c}, are classified, as shown in Fig.
 \ref{neo_orbit_types}, according to their semi-major axis (a), perihelion distance (q), and aphelion distance (Q).
-Here we report the Center for Near-Earth Object Studies (CNEOS) definition \cite{nasa_classification}:
+Here we report the CNEOS definition \cite{nasa_classification}:
 
@@
 
+On the basis of this definition, the asteroids contained in the dataset here analysed are classified as hazardous or not.
+
 \begin{figure}[h]
@@
 \begin{center}
+\includegraphics[width=1\textwidth]{Figures/Orbit_asteorids.png}
+\caption{Orbits of potentially hazardous asteorids. Image taken from \cite{orbits-of-potentially-hazardous}}
+\label{Orbit_asteorids}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/heic1715c.jpg}
+\caption{Location of the solar system asteroids belt. Image taken from \cite{esahubble}}
+\label{heic1715c}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
 \includegraphics[width=1\textwidth]{Figures/neo_orbit_types.jpg}

### 2021-11-24 Manuscript/manuscript.tex
PUB

@@
 Furthermore the parameters related to the diameters are also not connected with the orbital parameters.
-Furthermore we see also that the mean velocity is not directly related to the mean motion but there are the orbital parameters within them as expected.
+In addition we see also that the mean velocity is not directly related to the mean motion but there are the orbital parameters within them as expected.
 Thus it can be said that the model produced is almost consistent with the astronomic laws.
@@
 
-From which we get that the velocity of the asteroid is maximum at the perihelion, and minimum at the aphelion. Their values that are equal respectively to \cite{murray1999solar}:
+From which we get that the \textcolor{red}{velocity} \footnote{In the dataset this quanitity is called relative velocity because the it the reference system is Earth} of the asteroid is maximum at the perihelion, and minimum at the aphelion.
+Their values that are equal respectively to \cite{murray1999solar}:
 
@@
 \ref{Orbit_asteorids}, by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}.
-In particular if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO).
+In particular, if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO).
 Such items are of three type: comets, meteoroids and asteroids.
@@
 Asteroids, on the other side, are items with a diameter larger than 1 m \cite{burbine2016asteroids}.
-The first two items are excluded from the present analysis.
+The first two bodies are excluded from the present analysis.
 The near asteroids, which are mostly located on the sites reported in Fig.
@@
 
-On the basis of this definition, the asteroids contained in the dataset here analysed are classified as hazardous or not.
+This is the formal definition by which the asteroids contained in the dataset here analysed are classified as hazardous or not.
 

### 2021-11-25 Manuscript/manuscript.tex
PUB

@@
 
+The description and the forecast of physical \footnote{Where here physical should be inteded as mesuarable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms one can formulate theoretical models that provide equations that describe such phenomena.
+Such approach, which will be called here \textit{ab-initio}.
+On the other side one can start from data and fit them into some general models (with fixed or not fixed number of parameters).
+This method will be called here as machine learning (ML) method.
+In the first case one have the advantage to have an almost full explanation of the meaning of the laws obtained and an overall idea of why nature works in this way.
+The toll that has to be paid in this case is that, since the phenomena are much more complicated with respect to the starting principles, a consistent part of them may be excluded by the assumption made at the beginning.
+Furthermore the calculation of the solution given the dynamical equation can be computationally expensive\footnote{This is the case for instance of Hartree-Fock method or of Density Functional Theory: such methods rewrite, with some approximations, the Schroedinger equation in to a much more computationally affordable way.
+However also in this case the solution for solids can be very expensive \cite{martin_2004}} .
+On the other side, the ML methods since they make much more general assumptions and work directly on data, are not limited to a particular class of phenomena.
+Therefore they do not require a development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation.
+The toll that has to be paid in this case it that the underlying mechanism by which a forecast is preferred to another one may be not so interpretable.
+Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century astronomers know that mercury follows an elliptical orbit that instead of being fixed rotates around the Sun as shown in Fig.
+\ref{Perihelion_precession2} Such phenomenon is called \textit{Precession of the perihelion of Mercury}.
+The crucial point is that there was no way to explain from the Newton gravity theory such phenomenon.
+At this point scholars had two paths: describe its motion with an empirical law (perhaps with an empirical modification of Newtons gravity law) or instead reformulate from scratch the Newton gravity.
+The second approach was the one followed by A.Einstein with his general relativity theory (further details are given in the A.Zee textbook \cite{zee2013einstein}).
+The graphical methods represent a good compromise  between these two approaches: they can be used for every physical phenomena and, since they provide the conditional dependencies/independences they also give a fine and intuitive representation  of their mechanisms.
+Therefore, if the theory of the process is known, one can use this class of algorithm and compare its finding with the theory.
+In this way one can not only evaluate the quality of the forecasts, but also of the model provided by the graphical methods.
+The aim of the present work is to perform such analysis for a dataset that contain the features and the hazardousness \footnote{For Earth} of asteroids.
+For this dataset, indeed, the relationships between the various features are known.
+The present work is organized as follows: first a brief introduction about the basic theoretical concepts of the statistical methods here used will be provided, then the dataset will be described and finally, the main results here obtained with graphical methods will be presented as compared with other type of machine learning algorithms.
+Furthermore an appendix about the theoretical concepts of the theory that explains and interconnects the dataset features is also provided at the end of the present work.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Perihelion_precession2.png}
+\caption{Precession of the perihelion of Mercury. Image taken from \cite{Perihelion_precession}}
+\label{Perihelion_precession2}
+\end{center}
+\end{figure}
+
 \chapter{Theoretical Framework}
 
-In this section we are going to review the theoretical concepts that underlies to the probabilistic methods here used:  we will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller et al.
+In this section we are going to review the theoretical concepts on which the probabilistic methods here used are based.
+We will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller et al.
 \cite{koller2009probabilistic}, H{\o}jsgaardand \cite{hojsgaard2012graphical} et al.
@@
 
-\chapter{Results}
+\chapter{Results and discussion}
 
@@
 
+The Glasso (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that correctly capture the physical connections of the dataset feature.
+In particular for both models we saw that the no one of the orbital parameters are conditionally dependent with respect to the diameter of the asteroid, as stated by theory.
+Furthermore the mgm model is able to provide forecasts about the hazardousness of the asteroids with performances that are almost equal or better  than the logistic and QDA algorithms.
+On the other side the fact that its performances are lower with respect to the RF and SVM is definitely compensated by its interpretability.
+
 \chapter{Appendix A: Concepts of astronomy}

### 2021-11-26 Manuscript/manuscript.tex
PUB

@@
 \cite{russell2010artificial}.
-Furthermore we will provide also a rapid overview of the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches,  since we used some of its concepts in the preliminary analysis of the dataset.
-Finally we concluded this chapter with a rapid overview about the concept of algorithm intepretability, which together with the accuracy and the confusion matrix, is necessary for the assessment of a ML performances.
+Furthermore we will provide also a rapid overview about the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches,  since we used some of its concepts in the preliminary analysis of the dataset.
+Finally this chapter will be concluded with a rapid overview about the concept of algorithm intepretability.
+Such concept, together with the accuracy and the confusion matrix, is necessary for the assessment of a ML performances.
 On the other side the concepts related to the celestial mechanics here used will be described, following the Murray approach \cite{murray1999solar} into the Appendix A
@@
 Lets start by supposing that we would represent compactly a joint distribution such as \cite{murphy2012machine}:
-
 \begin{equation}
@@
 \end{equation}
-
 that can represent for instance words in a documents or pixels of an image.  Firstly we know that using the chain rule,  we can decompose it, into the following form \cite{murphy2012machine}:
-
 \begin{equation}
@@
 \end{equation}
-
 where V is the number of variables and 1:V stands for ${1,2,...,V}$.
 This decomposition makes explicit the conditional probability tables, or in other terms the transition probability tensors \cite{wu2017markov}.
-As one can point out the number of parameter is cumbersome as the number of variables grows: indeed the number of parameter required scales as $\mathcal{O}(K^{V})$.
-Such formidable problem can be attacked by considering the concept of conditional independence.  This is defined as  \cite{murphy2012machine}:
+As one can point out, the number of parameters is cumbersome as the number of variables grows: indeed the number of parameter required scales as $\mathcal{O}(K^{V})$.
+Such formidable problem can be attacked by considering the concept of conditional independence.
+This is defined as\cite{murphy2012machine}:
 
@@
 \section{Information theory}
-Given an ensemble of random variable, we can find the amount of information that one variable contains of another one: such quantity is called mutual information and it is a key concept within the information theory.
+Given an ensemble of random variable, we can qunaitify the amount of information that one variable contains of another one: such quantity is called mutual information and it is a key concept within the information theory.
 This approach, that was implemented by Claude Shannon decades before the probabilistic modelling, represent a complementary way by which one can attack the problem of conditional dependence between random variables.
-Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allows to properly define the concept of mutual information.
-The starting concept of information theory is the entropy which express the uncertainty of a random variable.
-Given a random variable $X$ with alphabet (the accessible states) $\cite{cover2006elements}:$ and probability mass function $p(x)=Pr\left\lbrace X=x \rbrace\; x \in \chi$ we define the entropy of $X$ as $H(X)=-\sum_{x \in X} p(x)log p(x)$ where the logarithm has to be considered with basis 2.
+Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allow to properly define the concept of mutual information.
+The founding concept of information theory is the entropy.
+This quantity express the uncertainty of a random variable.
+Given a random variable $X$ with alphabet (the accessible states) $\cite{cover2006elements}$ and probability mass function $p(x)=Pr\left\lbrace X=x \rbrace\; x \in \chi$ we define the entropy of $X$ as $H(X)=-\sum_{x \in X} p(x)log p(x)$ where the logarithm has to be considered with basis 2.
 In analogous way the joint entropy of two random variables $(X,Y)$ with a joint distribution p(x,y) is defined as \cite{cover2006elements}:
@@
 This feature, at a first instance, is connected on how much the algorithm provide explanation to a human on how it choose a forecast instead of another one, or in other word how much the algorithm is not a blackbox.
-Apart from all legal problems connected with an algorithm that is a blackbox\footnote{Think for instance the conditions imposed by the GDPR (e.g the right to explanation)}, in the author opinion results provided by such algorithm cannot be considered scientific, since the scientific method also require to provide an explanation and not simply almost correct forecast.
-Following the \cite{tarski1953undecidable} argument, which is root on mathematical logic, it can be said that the a formal theory T into S if and only if S is able to proof the theorem of T in its own language.
-On the other side we would also that an algorithm, as an explanation, is complete: we expect that it will be also able to make correct forecasts.
+Apart from all legal problems connected with an algorithm that is a blackbox\footnote{Think for instance the conditions imposed by the GDPR (e.g the right to explanation)}, in the author opinion results provided by such algorithm cannot be considered scientific, since the scientific method also requires to provide an explanation and not simply almost correct forecast.
+Following the Tarski et al.
+\cite{tarski1953undecidable} argument, which is based on mathematical logic, it can be said that the a formal theory T can be translated into S if and only if S is able to proof the theorem of T in its own language.
+On the other side we would also that an algorithm, as an explanation, is complete: we expect that it will be also able to make correct forecasts for all available data.
 However as shown by \cite{doshi2017towards} and by \cite{gilpin2018explaining} the intepretability of an algorithm is linked with its incompleteness.
 This is something that is nested both in formal system as algorithms or scientific theories.
-For instance lets consider the fall of object: at first glance one can consider the only the items on Earth, in this case the acceleration is constant and it is a very simple theory; but as one moves to consider also the interaction within planets a much more complicate law should be considered (for which the previous case is a particular one).
-The first one is largely explainable but not poorly complete, the second one more difficult to explain but more complete.
+For instance lets consider the fall of object: at first glance one can consider the only the items on Earth, in this case the acceleration is constant and it is a very simple theory; but as one moves to consider also the interaction within planets a much more complicate law should be considered (in which the previous case is a particular one).
+The first one is fully explainable but poorly complete, the second one more difficult to explain but much more complete.
 Therefore, in principle, the evaluation of an algorithm performances should be considered not only in to one dimension, the accuracy, but also by considering its intepretability as shown in Fig.
@@
 The asteroid dataset was retrieved from Kaggle \cite{kaggle_dataset}, which reports into an more user-friendly mode the dataset of The Center for Near-Earth Object Studies (CNEOS) \cite{cneos+nasa}, a NASA research center.
-Among the 40 features there were present in the dataset we excluded the ones that were clearly redundant (such as the distance quantity evaluated in miles instead of kilometres), the ones that were connected only to the other name of the asteroid, the one connected to the name of the orbit and the one connected with the orbiting planet, since for all it was the Earth.
-Thus the features obtained were reduced to 22.
+Among the 40 features that were present in the dataset we excluded the ones that were clearly redundant (such as the distance quantity evaluated in miles instead of kilometres), the ones that were connected only to the other name of the asteroid, the one connected to the name of the orbit and the one connected with the orbiting planet, since for all it was the Earth.
+Thus the features obtained were therefore reduced to 22.
 Their description will be postponed to Appendix A, since a description of them cannot be disjoint with celestial mechanics concepts.

### 2021-11-26 Manuscript/manuscript.tex
pub

@@
 
-The first inspection that was performed on the dataset was related the density distributions of a selection of continuous features that are known, from the celestial mechanics, to be important for the prediction of the asteroids dangerousness.
+The first inspection that was performed on the dataset was related to the density distributions of a selection of continuous features that are known, from the celestial mechanics, to be important for the prediction of the asteroids dangerousness.
 These are reported in Fig.

### 2021-11-27 Script_FINAL.R
PUB

@@ -1,0 +1,410 @@
+library(readxl)
+library(PerformanceAnalytics)
+library(corrplot)
+library(ggcorrplot)
+library(mgm)
+library(gRim)
+library(bnlearn)
+library(gRapHD)
+library(qgraph)
+library(Rgraphviz)
+library(igraph)
+library(gRbase)
+library(RBGL)
+library(caret)
+library(pROC)
+library(ROCR)
+library(ggplot2)
+library(FactoMineR)
+library(factoextra)
+library(ggplot2)
+library(ggpubr)
+library(gRapHD)
+library(glasso)
+library(SIN)
+library(gRain)
+library(randomForest)
+library(randomForestExplainer)
+library(e1071)
+library(MASS)
+library(infotheo)
+library(varrank)
+library(minet)
+
+#Data Loading
+
+#Asteroids2 <- read_excel("Dataset/Asteroids2.xlsx")
+#Asteroids2<-na.omit(Asteroids2)
+
+
+
+
+#Asteroids2_h<-subset(Asteroids2,Hazardous==TRUE)
+#Asteroids2_n<-subset(Asteroids2,Hazardous==FALSE)
+
+#Asteroids_NH<- sample(3878,2800)
+
+#Asteroids2_n<-Asteroids2[Asteroids_NH,]
+
+#Asteroids_NH_frame<-Asteroids2[Asteroids_NH,]
+
+#Asteroids_FINAL<-rbind(Asteroids2_h,Asteroids2_n)
+
+#save(Asteroids_FINAL,file="Asteroids_FINAL.rda")
+
+load("Asteroids_FINAL.rda")
+
+Asteroids_FINAL[,2:21]<-scale(Asteroids_FINAL[,2:21])
+
+Asteroids_FINAL_double<-as.data.frame(lapply(Asteroids_FINAL[,2:22], as.double))
+Asteroids_FINAL_double["Hazardous"] <- as.factor(Asteroids_FINAL_double[,21])
+
+
+SS <- CGstats(Asteroids_FINAL_double)
+SS
+cc<-apply(SS$center,1,sd) /apply(SS$center,1,mean)
+cc
+
+M_cov <- cov(Asteroids_FINAL_double[,-21])
+M_cor <- cor(Asteroids_FINAL_double[,-21])
+
+chart.Correlation(Asteroids_FINAL_double , histogram=TRUE, pch=19)
+testRes = cor.mtest(M_cor, conf.level = 0.95)
[…]

### 2021-11-27 Manuscript/manuscript.tex
PUB

@@
 \setlength{\epigraphwidth}{\textwidth}
+\newcommand{\textgreek}[1]{\begingroup\fontencoding{LGR}\selectfont#1\endgroup}
+\lstset{language=R,
+    basicstyle=\small\ttfamily,
+    stringstyle=\color{DarkGreen},
+    otherkeywords={0,1,2,3,4,5,6,7,8,9},
+    morekeywords={TRUE,FALSE},
+    deletekeywords={data,frame,length,as,character},
+    keywordstyle=\color{blue},
+    commentstyle=\color{DarkGreen},
+}
 
@@
 
+\chapter{Appendix B: R code}
+
+\definecolor{light-gray}{gray}{0.95}
+\lstset{ columns=fullflexible, basicstyle=\ttfamily, backgroundcolor=\color{light-gray},xleftmargin=0.5cm,frame=lr,framesep=8pt,framerule=0pt,frame=single,breaklines=true, postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}}
+\lstinputlisting[language=R]{Script_FINAL.R}
+
 \bibliographystyle{unsrt}

### 2021-11-27 Manuscript/Script_FINAL.R

@@ -1,0 +1,372 @@
+library(readxl)
+library(PerformanceAnalytics)
+library(corrplot)
+library(ggcorrplot)
+library(mgm)
+library(gRim)
+library(bnlearn)
+library(gRapHD)
+library(qgraph)
+library(Rgraphviz)
+library(igraph)
+library(gRbase)
+library(RBGL)
+library(caret)
+library(pROC)
+library(ROCR)
+library(ggplot2)
+library(FactoMineR)
+library(factoextra)
+library(ggplot2)
+library(ggpubr)
+library(gRapHD)
+library(glasso)
+library(SIN)
+library(gRain)
+library(randomForest)
+library(randomForestExplainer)
+library(e1071)
+library(MASS)
+library(infotheo)
+library(varrank)
+library(minet)
+
+#Data Loading: this portion of code loads the original asteroid dataset and randomly extracts a set of not Hazardous asteroids. It is not necessary to 
+#execute since the operation was previously performed and its output was put on to an .rda file. 
+
+#Asteroids2 <- read_excel("Dataset/Asteroids2.xlsx")
+#Asteroids2<-na.omit(Asteroids2)
+
+#Asteroids2_h<-subset(Asteroids2,Hazardous==TRUE)
+#Asteroids2_n<-subset(Asteroids2,Hazardous==FALSE)
+
+#Asteroids_NH<- sample(3878,2800)
+
+#Asteroids2_n<-Asteroids2[Asteroids_NH,]
+
+#Asteroids_NH_frame<-Asteroids2[Asteroids_NH,]
+
+#Asteroids_FINAL<-rbind(Asteroids2_h,Asteroids2_n)
+
+#save(Asteroids_FINAL,file="Asteroids_FINAL.rda")
+
+load("Asteroids_FINAL.rda")
+
+Asteroids_FINAL[,2:21]<-scale(Asteroids_FINAL[,2:21])
+
+Asteroids_FINAL_double<-as.data.frame(lapply(Asteroids_FINAL[,2:22], as.double))
+Asteroids_FINAL_double["Hazardous"] <- as.factor(Asteroids_FINAL_double[,21])
+
+#Preliminary analysis
+
+SS <- CGstats(Asteroids_FINAL_double)
+SS
+cc<-apply(SS$center,1,sd) /apply(SS$center,1,mean)
+cc
+
+M_cov <- cov(Asteroids_FINAL_double[,-21])
[…]

### 2021-11-27 Manuscript/manuscript.tex
PUB

@@
 
-\section*{\begin{center}
-Abstract
+\section*{Abstract} The machine learning algorithms provide a promising approach to classify and predict natural or social phenomena.
+Differently to the oldest theoretical approach in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynman2018feynman}), the machine learning algorithms start from very general assumptions/expressions and then their parameters, are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}.
+The first approach allows obtaining interpretable predictions, the second one good forecast with a reduced effort.
+However, the tool that has to be paid in this case is the low interpretability \cite{russell2010artificial,murphy2012machine}.
+Since they provide the list of conditional independences/dependencies, the graphical methods represent a good compromise between the need for interpretability and the forecasts obtained without developing a general theory.
+In order to prove the validity of this statement, I considered a dataset for which the theory that interconnects its feature was known: in particular, I chose the asteroids hazardousness as provided by CNEOS \cite{cneos+nasa} and published on Kaggle \cite{kaggle_dataset}.
+The outcomes of the probabilistic methods for predicting the asteroid hazardousness were compared with the ones provided by the theory and with the ones of Random Forest, Support Vector Machines, and Logistic Regression and Quadric Discriminant analysis.
+The results show that the forecast performances of probabilistic methods are better than QDA and almost equal to the logistic regression but lower to RF and SVM.
+However, since the list of conditional dependencies correctly reflects the laws of celestial mechanics \cite{murray1999solar}, it can be said that in this case, the probabilistic methods provide an interpretable and correct explanation of their mechanism.
+Therefore, contrary to other machine learning algorithms, such methods can be fully validated scientifically.
+
 \end{center}}

### 2021-11-28 Manuscript/manuscript.tex
PUB

@@
 \section*{Abstract} The machine learning algorithms provide a promising approach to classify and predict natural or social phenomena.
-Differently to the oldest theoretical approach in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynman2018feynman}), the machine learning algorithms start from very general assumptions/expressions and then their parameters, are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}.
+Differently to the oldest theoretical approach in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynmanlectures}), the machine learning algorithms start from very general assumptions/expressions and then their parameters, are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}.
 The first approach allows obtaining interpretable predictions, the second one good forecast with a reduced effort.
-However, the tool that has to be paid in this case is the low interpretability \cite{russell2010artificial,murphy2012machine}.
+However, the toll that has to be paid in this case is the low interpretability \cite{russell2010artificial,murphy2012machine}.
 Since they provide the list of conditional independences/dependencies, the graphical methods represent a good compromise between the need for interpretability and the forecasts obtained without developing a general theory.
 In order to prove the validity of this statement, I considered a dataset for which the theory that interconnects its feature was known: in particular, I chose the asteroids hazardousness as provided by CNEOS \cite{cneos+nasa} and published on Kaggle \cite{kaggle_dataset}.
-The outcomes of the probabilistic methods for predicting the asteroid hazardousness were compared with the ones provided by the theory and with the ones of Random Forest, Support Vector Machines, and Logistic Regression and Quadric Discriminant analysis.
+The outcomes of the probabilistic methods for predicting the asteroid hazardousness were compared with the ones provided by the theory and with the ones of Random Forest (RF), Support Vector Machines (SVM), and Logistic Regression and Quadric Discriminant analysis (QDA).
 The results show that the forecast performances of probabilistic methods are better than QDA and almost equal to the logistic regression but lower to RF and SVM.
@@
 
-\chapter{Introduction}
-
-The description and the forecast of physical \footnote{Where here physical should be inteded as mesuarable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms one can formulate theoretical models that provide equations that describe such phenomena.
+\chapter{Introduction} The description and the forecast of physical \footnote{Where here physical should be intended as mesuarable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms, one can formulate theoretical models that provide equations that describe such phenomena.
 Such approach, which will be called here \textit{ab-initio}.
-On the other side one can start from data and fit them into some general models (with fixed or not fixed number of parameters).
-This method will be called here as machine learning (ML) method.
-In the first case one have the advantage to have an almost full explanation of the meaning of the laws obtained and an overall idea of why nature works in this way.
-The toll that has to be paid in this case is that, since the phenomena are much more complicated with respect to the starting principles, a consistent part of them may be excluded by the assumption made at the beginning.
-Furthermore the calculation of the solution given the dynamical equation can be computationally expensive\footnote{This is the case for instance of Hartree-Fock method or of Density Functional Theory: such methods rewrite, with some approximations, the Schroedinger equation in to a much more computationally affordable way.
-However also in this case the solution for solids can be very expensive \cite{martin_2004}} .
-On the other side, the ML methods since they make much more general assumptions and work directly on data, are not limited to a particular class of phenomena.
-Therefore they do not require a development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation.
-The toll that has to be paid in this case it that the underlying mechanism by which a forecast is preferred to another one may be not so interpretable.
-Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century astronomers know that mercury follows an elliptical orbit that instead of being fixed rotates around the Sun as shown in Fig.
+On the other side, one can start from data and fit them into some general models (with fixed or not fixed number of parameters).
+These methods will be called here the machine learning (ML) methods.
+In the first case, one can have a full explanation of the laws obtained and an overall idea of why nature works in this way.
+The toll that has to be paid in this case is that, since the phenomena are much more complicated than the starting principles, a consistent part of them may be excluded by the assumption made at the beginning.
+Furthermore, the calculation of the solution given the dynamical equation can be computationally expensive\footnote{This is the case, for instance, of the Hartree-Fock method or Density Functional Theory: such methods rewrite, with some approximations, the Schroedinger equation into a much more computationally affordable way.
+However, also, in this case, the solution for solids can be costly \cite{martin_2004}}.
+On the other side, the ML methods, since they make much more general assumptions and work directly on data, are not limited to a particular class of phenomena.
+Therefore they do not require the development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation.
+The drawback of this approach is that the underlying mechanism by which a forecast is preferred to another one may not be interpretable.
+Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century, astronomers knew that Mercury follows an elliptical orbit that, instead of being fixed, rotates around the Sun as shown in Fig.
 \ref{Perihelion_precession2} Such phenomenon is called \textit{Precession of the perihelion of Mercury}.
 The crucial point is that there was no way to explain from the Newton gravity theory such phenomenon.
-At this point scholars had two paths: describe its motion with an empirical law (perhaps with an empirical modification of Newtons gravity law) or instead reformulate from scratch the Newton gravity.
+At this point, scholars had two paths: describe its motion with an empirical law (perhaps with an empirical modification of Newton's gravity law) or reformulate the Newton gravity from scratch.
 The second approach was the one followed by A.Einstein with his general relativity theory (further details are given in the A.Zee textbook \cite{zee2013einstein}).
-The graphical methods represent a good compromise  between these two approaches: they can be used for every physical phenomena and, since they provide the conditional dependencies/independences they also give a fine and intuitive representation  of their mechanisms.
-Therefore, if the theory of the process is known, one can use this class of algorithm and compare its finding with the theory.
-In this way one can not only evaluate the quality of the forecasts, but also of the model provided by the graphical methods.
-The aim of the present work is to perform such analysis for a dataset that contain the features and the hazardousness \footnote{For Earth} of asteroids.
+The graphical methods represent a good compromise between these two approaches: they can be used for every physical phenomenon, and, since they provide conditional dependencies/independences, they also give an elegant and intuitive representation of their mechanisms.
+Therefore, if the process theory is known, one can use this class of algorithms and compare its finding with the theory.
+In this way, one can evaluate the quality of the forecasts and the model provided by the graphical methods.
+The present work aims to perform such analysis for a dataset that contain the features and the hazardousness \footnote{For Earth} of asteroids.
 For this dataset, indeed, the relationships between the various features are known.
-The present work is organized as follows: first a brief introduction about the basic theoretical concepts of the statistical methods here used will be provided, then the dataset will be described and finally, the main results here obtained with graphical methods will be presented as compared with other type of machine learning algorithms.
-Furthermore an appendix about the theoretical concepts of the theory that explains and interconnects the dataset features is also provided at the end of the present work.
+The present work is organized as follows: first, a brief introduction about the basic theoretical concepts of the statistical methods here used will be provided, then the dataset will be described, and finally, the principal results here obtained with graphical methods will be presented as compared with other types of machine learning algorithms.
+Furthermore, a recap of the celestial mechanics will be provided in Appendix A.
+This theory explains the connections and the meaning of the different features contained in the dataset.
 

### 2021-11-29 Manuscript/manuscript.tex
PUB

@@
 \chapter{Introduction} The description and the forecast of physical \footnote{Where here physical should be intended as mesuarable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms, one can formulate theoretical models that provide equations that describe such phenomena.
-Such approach, which will be called here \textit{ab-initio}.
+Such approach will be called here \textit{ab-initio}.
 On the other side, one can start from data and fit them into some general models (with fixed or not fixed number of parameters).
 These methods will be called here the machine learning (ML) methods.
-In the first case, one can have a full explanation of the laws obtained and an overall idea of why nature works in this way.
+In the ab-initio case, one can fully explain the laws obtained and an overall idea of why nature works in this way.
 The toll that has to be paid in this case is that, since the phenomena are much more complicated than the starting principles, a consistent part of them may be excluded by the assumption made at the beginning.
+The drawback of this approach is that the underlying mechanism by which a forecast is preferred to another one may not be interpretable.
 Furthermore, the calculation of the solution given the dynamical equation can be computationally expensive\footnote{This is the case, for instance, of the Hartree-Fock method or Density Functional Theory: such methods rewrite, with some approximations, the Schroedinger equation into a much more computationally affordable way.
@@
 Therefore they do not require the development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation.
-The drawback of this approach is that the underlying mechanism by which a forecast is preferred to another one may not be interpretable.
 Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century, astronomers knew that Mercury follows an elliptical orbit that, instead of being fixed, rotates around the Sun as shown in Fig.
@@
 The graphical methods represent a good compromise between these two approaches: they can be used for every physical phenomenon, and, since they provide conditional dependencies/independences, they also give an elegant and intuitive representation of their mechanisms.
-Therefore, if the process theory is known, one can use this class of algorithms and compare its finding with the theory.
+Therefore, if the theory of the involved process is known, one can use this class of algorithms and compare its finding with the theory.
 In this way, one can evaluate the quality of the forecasts and the model provided by the graphical methods.
@@
 
-In this section we are going to review the theoretical concepts on which the probabilistic methods here used are based.
-We will expose them following the approaches of Murphy \cite{murphy2012machine},  Koller et al.
-\cite{koller2009probabilistic}, H{\o}jsgaardand \cite{hojsgaard2012graphical} et al.
-and Russel et al.
-\cite{russell2010artificial}.
-Furthermore we will provide also a rapid overview about the main concepts of information theory,  following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches,  since we used some of its concepts in the preliminary analysis of the dataset.
-Finally this chapter will be concluded with a rapid overview about the concept of algorithm intepretability.
-Such concept, together with the accuracy and the confusion matrix, is necessary for the assessment of a ML performances.
-On the other side the concepts related to the celestial mechanics here used will be described, following the Murray approach \cite{murray1999solar} into the Appendix A
-
-\section{Markov random fields}
-
 Lets start by supposing that we would represent compactly a joint distribution such as \cite{murphy2012machine}:
@@
 \end{equation}
-where V is the number of variables and 1:V stands for ${1,2,...,V}$.
-This decomposition makes explicit the conditional probability tables, or in other terms the transition probability tensors \cite{wu2017markov}.
-As one can point out, the number of parameters is cumbersome as the number of variables grows: indeed the number of parameter required scales as $\mathcal{O}(K^{V})$.
-Such formidable problem can be attacked by considering the concept of conditional independence.
+Where V is the number of variables and 1:V stands for ${1,2,...,V}$.
+This decomposition makes explicit the conditional probability tables, or in other terms, the transition probability tensors \cite{wu2017markov}.
+As one can point out, the number of parameters is cumbersome as the number of variables grows: indeed, the number of parameter required scales as $\mathcal{O}(K^{V})$.
+Such a challenging problem can be attacked by considering the concept of conditional independence.
 This is defined as\cite{murphy2012machine}:
@@
 
-In this case a first order Markov chain is obtained,  where the transition tensor is of second order \cite{wu2017markov}.
-Given this formalism we are interested in finding a smart way to plot such joint distribution into an intuitive way: the graph theory provide the answer to this quest.
-In particular the random variables can be represented by nodes and presence of conditional independence for two random variables by the lack of an edge that interconnects them.
-Bayesian networks consider directed edges,  while Markov random fields (MRF) only undirected.
-As consequence,  while the the concept of topological ordering,  by which the parents n nodes are labelled with a lower  with respect to their children,  is well defined for Bayesian network,  for MRF is not.
-In order to solve this issue it is useful to consider the Hammersley-Clifford theorem as stated in \cite{murphy2012machine}:
+In this case, a first-order Markov chain is obtained,  where the transition tensor is of second-order \cite{wu2017markov}.
+Given this formalism, we are interested in finding an intelligent way to plot such joint distribution intuitively: the graph theory answers this quest.
+In particular, the absence of an edge that interconnects the random variables can be represented by nodes and the presence of conditional independence for two random variables.
+Bayesian networks consider directed edges,  while Markov random fields (MRF) are only undirected.
+Consequently,  while the concept of a topological ordering,  by which the parents n nodes are labelled lower to their children,  is well defined for the Bayesian network, MRF is not.
+In order to solve this issue, it is helpful to consider the Hammersley-Clifford theorem as stated in \cite{murphy2012machine}:
 
@@
 
-Finally in order to reduce the computational cost,  one can consider only the pairwise interaction instead of the maximum clique.
-This is the analogue of what is usually performed in solid state physics (but surely not always)  when only the interaction between the first neighbour is considered.
+Finally, to reduce the computational cost,  one can consider only the pairwise interaction instead of the maximum clique.
+This is the analogue of what is usually performed in solid-state physics (but not always)  when only the interaction between the first neighbour is considered.
 Another example is the  Ising model: here we have a lattice of spins that can be or in $\ket{+}$ or in $\ket{-}$ and their interaction is modelled by\cite{murphy2012machine}:
@@
 
-where $w_{st}=J$ represent the coupling strength between two neighbour site.  The collective state is described by
+Where $w_{st}=J$ represent the coupling strength between two neighbour site.  The collective state is described by
 
@@
 
-where $\otimes$ is the tensor product.
-If this parameter is associated with a positive finite value we have an associative Markov network: basically collective states in which all sites have the same configuration is favoured.
-Thus we will have two collective states: one for which we have all $\ket{+}$ and another in which we have all $\ket{-}$  Such situation would model,  in principle,  the ferromagnet materials where the external magnetic field induce into the material a magnetic filed with the same direction.
-On the other side if the magnetization of the material is opposite with respect to the external field,  and thus $J<0$,  we have an anti-ferromagnetic system in which frustrated states are present.
+Where $\otimes$ is the tensor product, if this parameter is associated with a positive finite value, we have an associative Markov network: collectively, all sites with the same configuration are favoured.
+Thus we will have two collective states: one for which we have all $\ket{+}$ and another in which we have all $\ket{-}$  Such a situation would model,  in principle,  the ferromagnetic materials where the external magnetic field induces into the material a magnetic field with the same direction.
+On the other side, if the material's magnetisation is opposite to the external field,  and thus $J<0$,  we have an anti-ferromagnetic system in which frustrated states are present.
 Furthermore lets consider the unnormalized log probability of a collective state $\textbf{y}=\ket{i_{1},i_{2},...,i_{n}}$ \cite{murphy2012machine}:
@@
 
-But this is nothing more that the well know \footnote{In physics} Hamiltonian of an Ising system.
-This is not a simple coincidence:  indeed the Hamiltonian of a system represent,  rudely speaking,  its total energy.
-Thus according to the Boltzmann or Gibbs distribution we have \cite{murphy2012machine}:
+The previous expression is the Hamiltonian of an Ising system.
+This is not a simple coincidence:  indeed, the Hamiltonian of a system represent,  rudely speaking,  its total energy.
+Thus according to the Boltzmann or Gibbs distribution, we have \cite{murphy2012machine}:
 
@@
 
-From this expression it is clear why this method is called moment matching; it is worth nothing tha such computation is largely expensive from a computational point of view: thus scholar usually consider other techinques or at least stochastic gradient descent method.
+From this expression, it is clear why this method is called moment matching; it is worth noting that such computation is largely expensive from a computational point of view: thus, scholars usually consider other techniques or at least the stochastic gradient descent method.
 A full review can be found in \cite{murphy2012machine} and \cite{koller2009probabilistic}.
 Finally we consider, as for the dataset in analysed in this work, the case where we have both discrete and continuous variables i.e.
-$x=\left(i_{1},...,i_{d},y_{1},...,y_{q} \right)$ with d discrete variable and q continuos variables.
-The are called in the literature Mixed Interaction Models.
-In this case the following density has to be considered \cite{hojsgaard2012graphical}:
+$x=\left(i_{1},...,i_{d},y_{1},...,y_{q} \right)$ with d discrete variable and q continuous variables.
+They are called in the literature Mixed Interaction Models.
+In this case, the following density has to be considered \cite{hojsgaard2012graphical}:
 
@@
 
-Moreover one can further modify the previous form in order to obtain a particular factorial expansion: such models are referred as homogeneous mixed interaction models \cite{hojsgaard2012graphical}.
+Moreover, one can further modify the previous form in order to obtain a particular factorial expansion: such models are referred to as homogeneous mixed interaction models \cite{hojsgaard2012graphical}.
 
 \section{Information theory}
-Given an ensemble of random variable, we can qunaitify the amount of information that one variable contains of another one: such quantity is called mutual information and it is a key concept within the information theory.
-This approach, that was implemented by Claude Shannon decades before the probabilistic modelling, represent a complementary way by which one can attack the problem of conditional dependence between random variables.
-Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allow to properly define the concept of mutual information.
-The founding concept of information theory is the entropy.
-This quantity express the uncertainty of a random variable.
+Given an ensemble of random variables, we can quantify the amount of information that one variable contains: such quantity is called mutual information and is a key concept within information theory.
+This approach, which was implemented by Claude Shannon decades before the probabilistic modelling, represents a complementary way by which one can attack the problem of conditional dependence between random variables.
+Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allow defining the concept of mutual information properly.
+The founding concept of information theory is entropy.
+This quantity expresses the uncertainty of a random variable.
 Given a random variable $X$ with alphabet (the accessible states) $\cite{cover2006elements}$ and probability mass function $p(x)=Pr\left\lbrace X=x \rbrace\; x \in \chi$ we define the entropy of $X$ as $H(X)=-\sum_{x \in X} p(x)log p(x)$ where the logarithm has to be considered with basis 2.
@@
 \end{equation}
-Given a distribution q and another distribution p, one can quantify how inefficently the second one describe the first one using the concept of relative entropy or Kullback-Leibler distance \cite{cover2006elements,mackay2003information}:
+Given a distribution q and another distribution p, one can quantify how inefficiently the second one describes the first one using the concept of relative entropy or Kullback-Leibler distance \cite{cover2006elements,mackay2003information}:
 \begin{equation}
@@
 \end{equation}
-this quantity can not be negative: the entropy of a random variable associated to another cannot have a degree of uncertainty lower with respect to the quantity that its aimed to describe.
-On these basis we are now ready to introduce the concept of mutual information.
+this quantity can not be negative: the entropy of a random variable associated with another cannot have a degree of uncertainty lower to the quantity it aims to describe.
+On these bases, we are now ready to introduce the concept of mutual information.
 This is defined as  \cite{cover2006elements}:
@@
 \end{equation}
-As for the joint distribution also in this case we have a chain rule  \cite{cover2006elements}:
+As for the joint distribution also in this case, we have a chain rule  \cite{cover2006elements}:
 \begin{equation}
@@
 \end{equation}
-Finally we would report the data process inequality theorem that connects the information theory with the Markov chain: if we have a Markov chains, $X\rightarrow Y \rightarrow Z$  then $I(X;Y)\geq I(X;Z) $.
+Finally, we would report the data process inequality theorem that connects the information theory with the Markov chain: if we have a Markov chain, $X\rightarrow Y \rightarrow Z$  then $I(X;Y)\geq I(X;Z) $.
 As for the Gibbs inequality, the underlying idea is that no clever manipulation of the data can improve the inference that can be made on them \cite{cover2006elements,mackay2003information}.
-Otherwise we would have a clear violation of the second principle of thermodynamics (see for instance the Maxwell's demon \cite{feynman2018feynman})
+Otherwise, we would have a clear violation of the second principle of thermodynamics (see for instance, the Maxwell's demon \cite{feynman2018feynman}).
 
@@
 
-Beside the accuracy and other features that are connected with the confusion matrix, another important feature is the algorithm intepretability.
-This feature, at a first instance, is connected on how much the algorithm provide explanation to a human on how it choose a forecast instead of another one, or in other word how much the algorithm is not a blackbox.
-Apart from all legal problems connected with an algorithm that is a blackbox\footnote{Think for instance the conditions imposed by the GDPR (e.g the right to explanation)}, in the author opinion results provided by such algorithm cannot be considered scientific, since the scientific method also requires to provide an explanation and not simply almost correct forecast.
+Besides the accuracy and other features connected with the confusion matrix, another critical feature is algorithm interpretability.
+This feature represents how much the algorithm explains its predictions and, in general, its mechanics.
+In other words, how much the algorithm is not a black box.
+Apart from all legal problems connected with an algorithm that is a blackbox\footnote{Think for instance the conditions imposed by the GDPR (e.g.
+right to explanation)}, in the author opinion, results provided by such an algorithm cannot be considered scientific since the scientific method also requires to explain and not simply almost correct forecast.
 Following the Tarski et al.
-\cite{tarski1953undecidable} argument, which is based on mathematical logic, it can be said that the a formal theory T can be translated into S if and only if S is able to proof the theorem of T in its own language.
-On the other side we would also that an algorithm, as an explanation, is complete: we expect that it will be also able to make correct forecasts for all available data.
-However as shown by \cite{doshi2017towards} and by \cite{gilpin2018explaining} the intepretability of an algorithm is linked with its incompleteness.
-This is something that is nested both in formal system as algorithms or scientific theories.
-For instance lets consider the fall of object: at first glance one can consider the only the items on Earth, in this case the acceleration is constant and it is a very simple theory; but as one moves to consider also the interaction within planets a much more complicate law should be considered (in which the previous case is a particular one).
-The first one is fully explainable but poorly complete, the second one more difficult to explain but much more complete.
-Therefore, in principle, the evaluation of an algorithm performances should be considered not only in to one dimension, the accuracy, but also by considering its intepretability as shown in Fig.
+\cite{tarski1953undecidable} argument, which is based on mathematical logic, it can be said that the formal theory T can be translated into S if and only if S can prove the theorem of T in its language.
+On the other side, we would also know that an algorithm, as an explanation, is complete: we expect that it will also be able to make correct forecasts for all available data.
+However, as shown by \cite{doshi2017towards}, and by \cite{gilpin2018explaining} the interpretability of an algorithm is linked with its incompleteness.
+This is something that is nested both in formal systems as algorithms or scientific theories.
+For instance, let us consider the fall of object: at first glance, one can consider the only the items on Earth, in this case, the acceleration is constant, and it is a straightforward theory; but as one moves to consider also the interaction within planets a much more complicate law should be considered (in which the previous case is a particular one).
+The first one is fully explainable but poorly complete; the second one is more difficult to explain but much more complete.
+Therefore, in principle, the evaluation of algorithm performances should be considered not only into one dimension, the accuracy but also by considering its interpretability as shown in Fig.
 \ref{ML_intepretability}
@@
 
-The asteroid dataset was retrieved from Kaggle \cite{kaggle_dataset}, which reports into an more user-friendly mode the dataset of The Center for Near-Earth Object Studies (CNEOS) \cite{cneos+nasa}, a NASA research center.
-Among the 40 features that were present in the dataset we excluded the ones that were clearly redundant (such as the distance quantity evaluated in miles instead of kilometres), the ones that were connected only to the other name of the asteroid, the one connected to the name of the orbit and the one connected with the orbiting planet, since for all it was the Earth.
+The asteroid dataset was retrieved from Kaggle \cite{kaggle_dataset}, which reports into a more user-friendly mode the dataset of The Center for Near-Earth Object Studies (CNEOS) \cite{cneos+nasa}, a NASA research centre.
+Among the 40 features that were present in the dataset,  the following cases were excluded: the redundant features (e.g.
+the distance quantity evaluated in miles instead of kilometres), the features connected only to the other name of the asteroid, the features connected to the name of the orbit and the one connected with the orbiting planet ( since for all it was the Earth).
 Thus the features obtained were therefore reduced to 22.
-Their description will be postponed to Appendix A, since a description of them cannot be disjoint with celestial mechanics concepts.
+Their description will be postponed to Appendix A since their description cannot be disjoint with celestial mechanics concepts.
 Their enumeration is provided in Tab.
 \ref{tab_features}.
-Since only the 16 \% of the asteroids were hazardous, we considered to reduce the number of the non-hazardous in order to improve the portion of the hazardous ones.
-For this purpose we constructed a database in which the proportion hazardous/not hazardous was 1:5: thus all the hazardous one were included, while the not-hazardous were randomly selected \footnote{The author is aware that in principle the proportion should be much less unbalanced, however the tool that has to be paid for this operation is an overall reduction of the cases in the dataset that lowers the performances of the algorithms here used.
-The proportion here used was a trade-off for these two contrasting facts}.
+Since only 16 \% of the asteroids were hazardous, we considered reducing the number of the non-hazardous to improve the portion of the hazardous ones.
+For this purpose, we constructed a database in which the proportion hazardous/not hazardous was 1:5: thus, all the hazardous ones were included, while the not-hazardous were randomly selected \footnote{The author is aware that in principle, the proportion should be much less unbalanced, however, the tool that has to be paid for this operation is an overall reduction of the cases in the dataset that lowers the performances of the algorithms here used.
+The proportion here is used as a trade-off for these two contrasting facts}.
 Furthermore, concerning the continuous measures, the dataset was standardised and demeaned.
@@
 \begin{table}[]
-\caption{The features used for the present analysis. The unit of measure are reported in the original dataset \cite{kaggle_dataset}.}
+\caption{The features used for the present analysis. The units of measure are reported in the original dataset \cite{kaggle_dataset}. The explanation of these features is provided in Appendix A. }
 \begin{center}
@@
 
-This chapter is organized in the following way: first we are going to report the results concerning the preliminary analysis performed on the dataset.
-This include the factor analysis of mixed data and the mutual information analysis of the continuous variables of asteroids vs their hazard.
-Then it will follow the analysis of the dataset performed with the probabilistic methods: after a preliminary analysis on the continuous variables, the mixed interaction model as well the minforest model obtained for the whole dataset will be presented and discussed.
-Finally the probabilistic models previously obtained will be compared with the outputs and the performances of four machine learning algorithm (Random Forest, Support vector machines Quadratic Discriminant Analysis and Logistic regression).
+This chapter is organized in the following way: first, we will report the results concerning the preliminary analysis performed on the dataset.
+The factor analysis of mixed data and the mutual information analysis of the continuous variables of asteroids vs their hazard will be reported.
+Then it will follow the analysis of the dataset performed with the probabilistic methods.
+After a preliminary analysis on the continuous variables, the mixed interaction model and the minforest model obtained for the whole dataset will be presented and discussed.
+Finally, the probabilistic models previously obtained will be compared with the outputs and the performances of four machine learning algorithms (Random Forest, Support vector machines, Quadratic Discriminant Analysis and Logistic regression).
 
@@
 
-We started the analysis with the graphical models by inspecting the relations between continuous variables in the dataset: thus in this first step the binary variable \textit{Hazardous} was excluded.
+The analysis with the graphical models was started by inspecting the relations between continuous variables in the dataset.
+Therefore, in this first step, the binary variable \textit{Hazardous} was excluded.
 For this purpose, among the different methods available, we considered the \textit{Graphical least absolute shrinkage and selection operator} GLASSO as implemented in the \textit{glasso} R package \cite{friedman2008sparse,glasso}.
 After different tests reported in Fig.
-\ref{GLASSO_convergence}, we considered as final result the graph obtained with the value of $\rho$ (the one that penalize further connections) equal to $0.3$.
-This is because, in this plot, we see that the conditional dependences/independences stated by the Celestial Mechanics are correctly reproduced: for instance we see that the diameter of the asteroids is independent from all features related to its motion.
-This is definitely meaningful since the mass of the asteroids is largely lower with respect to the mass of earth: thus there is no way by which the asteroid orbit can be modified by its mass as explained in Appendix A.
-Furthermore we see that the Close approach Date and the Epoch date are dependant each other but in no way from the other features: this is right since the date and epoch are set with an arbitrary scale.
-Also the Perihelion Epoch and Osculation Time are dependant, as expected, but there is no dependence with the orbit parameters.
-We see that the mean motion is conditionally independent with respect to relative velocity, but correctly this is dependent from perihelion distance.
+\ref{GLASSO_convergence}, we considered as the final result the graph obtained with the value of $\rho$ (the one that penalizes further connections) equal to $0.3$.
+Such choice is motivated by the fact that the conditional dependences/independences stated by the Celestial Mechanics are correctly reproduced with this parameter.
+For instance, we see that the diameter of the asteroids is independent of all features related to their motion.
+Such a result is meaningful since the volume/mass of the asteroids is largely lower than the mass of the earth.
+Thus, as explained in Appendix A, there is no way the asteroid orbit can be modified by its mass.
+In addition, we see that the Close approach Date and the Epoch date are linked to each other but in no way from the other features.
+This result is correct since the date and epoch are set with an arbitrary scale.
+Also, as expected, the Perihelion Epoch and Osculation Time are linked, but there is no dependence on the orbit parameters.
+Furthermore, the mean motion is conditionally independent with relative velocity, but correctly, this is dependent on perihelion distance.
+
 \begin{figure}[ht]
@@
 \end{figure}
-Lets now move on the mixed interaction model: first we considered its implementation in the mgm package \cite{mgm,haslbeck2015mgm}.
-We choose as $k$ parameter 2 and a cross validation (CV) with ten folds.
+
+Let us now move on to the mixed interaction model, where the hazardousness feature was also considered.
+The first model was calculated with the mgm package  \cite{mgm,haslbeck2015mgm} where the nodewise regression is used \cite{meinshausen2006high}.
+The $k$ parameter was set equal to two, and a cross-validation (CV) with ten folds was considered.
 The result is reported in Fig.
 \ref{mgm}.
-From this figure we see that the features that are conditionally dependant to the Hazardous features are: the absolute magnitude, the min orbit intersection, and the eccentricity.
-All these connections are supported by the theory: the min orbit intersection is the main parameter for the hazardous value, the eccentricity on the other side can be tough as a parameter that describes how near the celestial body moves at the perihelion (indeed this parameter is correctly linked with the perihelion distance) and the Absolute Magnitude is also a meaningful parameter for the hazard evaluation since, if the asteroid is too small it will be destroyed by the earth atmosphere.
-It is worth mentioning that this parameter is correctly not connected with all quantities related to the orbital parameter except for the mean anomaly.
-Furthermore the parameters related to the diameters are also not connected with the orbital parameters.
-In addition we see also that the mean velocity is not directly related to the mean motion but there are the orbital parameters within them as expected.
+From this figure, it can be seen that the features that are conditionally dependant on the Hazardous features are: the absolute magnitude, the min orbit intersection, and the eccentricity.
+The theory supports all these connections.
+The min orbit intersection is the main parameter for the hazardous value.
+On the other side, the eccentricity can be tough as a parameter that describes how near the celestial body moves at the perihelion (indeed, this parameter is correctly linked with the perihelion distance).
+The Absolute Magnitude is also a meaningful parameter for the hazard evaluation since, if the asteroid is too small, it will be destroyed by the earth atmosphere.
+Furthermore, the parameters related to the diameters are also not connected with the orbital parameters.
+In addition, the mean velocity is not directly related to the mean motion, but there are the orbital parameters within as expected.
 Thus it can be said that the model produced is almost consistent with the astronomic laws.
-On this basis, we can evaluate the performances of the model in term of confusion matrix, ROC curve and $\phi$ parameters (we used for this purpose the Caret R package \cite{kuhn2008building,caret}).
+On this basis, the performances of the model in terms of confusion matrix, ROC curve and $\phi$ parameters can be evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
 The confusion matrix obtained is reported in Fig.
@@
 \ref{ROC_mgm} and it corresponds to a $\phi=0.6$.
-We will comment these performances in a few when we will present the performances of other algorithms for the same dataset.
-Before this we would show the graphical models obtained with different implementation of mixed interaction model: we considered the function \textit{mmod()} of gRim package \cite{hojsgaard2012graphical} and the \textit{minforest()} of gRapHD (which uses the minForest method) \cite{de2009high}.
-Both were obtained with a stepwise algorithm.
+These performances will be commented on in the next section when the performances of other ML algorithms for the same dataset will be reported.
+Beside the mgm algorithm/package other approaches were also considered for the evaluation of the mixed interaction graphical model: the function \textit{mmod()} of gRim package \cite{hojsgaard2012graphical} and the \textit{minforest()} of gRapHD (which uses the minForest method) \cite{de2009high}.
+For both cases, a stepwise algorithm was used.
 Their result are reported in Fig.
 \ref{mmod} and \ref{minforest}.
-We see that in these models the minimum diameter is linked with the hazordous feature, and this is correct, but what is without a physical meaning is that this quantity is also linked to the features connected with the orbital parameters.
-Indeed, while for the magnitude (as for mgm model) a weak relationship can be admitted since this depends also on the distance between the asteroids and the observer and from the orbit, a volume and thus a mass dependence is not acceptable.
-In principle one can set these link as forbidden in the algorithm (blacklist), but in the author view it is preferable a model were the result is obtained without boundaries, instead of a one obtained with a large number of boundaries.
-Thus since a correct model was obtained without the imposition a blacklist, the mgm, we simply reject the mmod and minforest ones in favour of the first one.
-As consequence this is the reason why we report the performances of mgm model only.
+It can be seen that in these models, the minimum diameter is linked with the hazardous feature, which is correct, but what is without a physical meaning is that this quantity is also linked to the features connected with the orbital parameters.
+A weak link with the orbital parameters is meaningful for the magnitude (as the mgm model for the mean anomaly).
+The magnitude depends on the distance between the asteroids and the observer.
+However, a direct link from the volume parameters (and thus from the mass) with the orbital ones is not acceptable.
+In principle, one can set these links as forbidden in the algorithm (blacklist), but in the author view, a model where the result is obtained without boundaries should be preferred to those obtained with a large number of boundaries.
+Therefore we reject the mmod and minforest ones favouring the mgm, which was obtained without boundaries.
 
@@
 
-Given the mgm model as the one obtained with the probabilistic graphical methods, we would compare its performances with the ones obtained from the following machine learning algorithms: the random-forest (as implemented in the RandomForest package \cite{rfor}), the support vector machines (as implemented in the e1071 package \cite{dimitriadou2008misc}) , the quadratic discriminant analysis (as implemented in the MASS package \cite{MASS} and the logistic regression (as implented in stats package \cite{stats}).
+The performances of mgm model will be now compared with the obtained from other ML algorithms.
+These include the random-forest (as implemented in the RandomForest package \cite{rfor}), the support vector machines (as implemented in the e1071 package \cite{dimitriadou2008misc}) , the quadratic discriminant analysis (as implemented in the MASS package \cite{MASS} and the logistic regression (as implemented in stats package \cite{stats}).
 Their performances are reported in Fig.
@@
 \ref{phi_values}.
-From these comparison we can see that most of these algorithms, expect for the QDA outperform the graphical method mgm or have similar performances as for the logistic algorithm.
-Thus at this point one can ask what is the advantage of using a graphical method instead of a random forest or a SVM since its performances seems lower.
-The answer is the interpretability of the model provided: the random forest, at lest, can provide a variable rank importance as reported in Fig.
-\ref{RF_Importance} (note that apart from the Min Orbit intersection, there is a slight reshuffling in the feature importance with respect to the one established by the MI in fig \ref{Mutual_information}; however apart from this is a black-box as the SVM, the logistic regression and the QDA.
-Differently the probabilistic graph, providing the conditional independences, give to the user a interpretable model whose properties can be also compared, discussed and validated with the theory.
-Thus the model developed in this way allows a more scientific evaluation with respect to the black-box one.
-In the author view this characteristic definitely compensates the lack of predictive power with respect to the Random Forest or the SVM methods.
+This comparison shows that the RF and the SVM outperform the mgm graphical method while the logistic one has similar performances.
+Thus, one can ask the advantage of using a graphical method instead of a random forest or an SVM since its performances seem lower.
+The answer is the interpretability of the model provided: the random forest, at least, can provide variable rank importance as reported in Fig.
+\ref{RF_Importance} (note that apart from the Min Orbit intersection, there is a slight reshuffling in the feature importance to the one established by the MI in fig \ref{Mutual_information}; however apart from this is a black-box as the SVM, the logistic regression and the QDA.
+The probabilistic graph, providing conditional independences, gives the user an interpretable model whose properties can also be compared, discussed, and validated with the theory.
+Thus the model developed in this way allows a more scientific evaluation of the black-box ones.
+In the author's view, this characteristic compensates for the lack of predictive power regarding the RF or the SVM methods.
 
@@
 
-The Glasso (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that correctly capture the physical connections of the dataset feature.
-In particular for both models we saw that the no one of the orbital parameters are conditionally dependent with respect to the diameter of the asteroid, as stated by theory.
-Furthermore the mgm model is able to provide forecasts about the hazardousness of the asteroids with performances that are almost equal or better  than the logistic and QDA algorithms.
-On the other side the fact that its performances are lower with respect to the RF and SVM is definitely compensated by its interpretability.
+The GLASSO (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that correctly capture the physical connections of the dataset feature.
+In particular, for both models, it can be seen that no one of the orbital parameters is conditionally dependent on the diameter of the asteroid, as stated by theory.
+Furthermore, the mgm model can provide forecasts about the hazardousness of the asteroids with almost equal or better performances than the logistic and QDA algorithms.
+On the other hand, despite the mgm performances being lower than the RF and SVM, its interpretability is higher.
+Such a feature compensates the mgm forecast performances gap.
 
@@
 
-This Appendix is dedicated to review the basics concepts of astronomy that are needed to understand the features of the dataset and the theoretical framework in which these are involved.
-First the basic concepts of celestial mechanics will be recap following the Murray approach \cite{murray1999solar}: these will provide a description of the orbital parameters of the dataset.
+This Appendix is dedicated to reviewing the basic concepts of astronomy needed to understand the dataset's features and how these are interconnected.
+First, the basic concepts of celestial mechanics will be recapped following the Murray approach \cite{murray1999solar}: these will describe the orbital parameters of the dataset.
 Then the concepts connected to the observation of the asteroids will be reviewed following the Burbine textboox \cite{burbine2016asteroids}.
-Finally, using the previous definition, the classification of the asteroids and in particular the definition of their hazardousness will be provided following the Center for Near-Earth Object Studies (CNEOS) statements \cite{nasa_classification}.
+Finally, using the previous definition, the classification of the asteroids and in particular, the definition of their hazardousness will be provided following the Center for Near-Earth Object Studies (CNEOS) statements \cite{nasa_classification}.
 
@@
 
-Lets start by considering two masses $m_{1}$ and $m_{2}$, which in the present case will be respectively the planet Earth and the asteroid.
+Let us start by considering two masses $m_{1}$ and $m_{2}$, which in the present case will be respectively the planet Earth and the asteroid.
 Their position will be given, respectively, by two vectors $\textbf{r}_{1}$ and $\textbf{r}_{2}$ considering the origin $O$ bounded in to an inertial space.
 Furthermore we can define also the relative position with the vector $\textbf{r}=\textbf{r}_{2}-\textbf{r}_{1}$.
-Since we suppose that the masses are not interacting with the electromagnetic force, these will be bounded only by the gravitational interaction which is given by the Newton law \cite{murray1999solar}:
+Since we suppose that the masses are not interacting with the electromagnetic force, these will be bounded only by the gravitational interaction, which is given by the Newton law \cite{murray1999solar}:
 
@@
 
-in which $a$ is the semi-major axis of the conic.
+In which $a$ is the semi-major axis of the conic.
 The shape of these orbits is reported in Fig.
 \ref{Elliptical_orbit}.
-All the asteroids considered here an eccentricity $0<e<1$, therefore they have an elliptical orbits in which the earth lies in to one of the two focal point.
-It is worth nothing that this is the first Kepler law.
+All the asteroids considered here have an eccentricity $0<e<1$, they have elliptical orbits in which the Earth lies in one of the two focal points.
+It is worth noting that this is the first Kepler law.
 Looking to Fig.
-\ref{Elliptical_orbit} we can define the point of minimum distance between $m_{1}$ and the orbiting body as the pericentre or \textcolor{red}{perihelion}, and the maximum distance as the apocentre or the aphelion.
-The \textcolor{red}{semi-major axis}, here denoted as $b$ on the other side is defined as the distance between the pericentre and the apocentre.
+\ref{Elliptical_orbit} we can define the point of the minimum distance between $m_{1}$ and the orbiting body as the pericentre or \textcolor{red}{perihelion}, and the maximum distance as the apocentre or the aphelion.
+The \textcolor{red}{semi-major axis}, here denoted as $b$ on the other side, is defined as the distance between the pericentre and the apocentre.
 Using the following identity \cite{murray1999solar}:
@@
 \begin{equation}
-r=\frac{a(1-e^{2}}{1+e\cdot cos(\theta-\phi}
+r=\frac{a(1-e^{2}}{1+e\cdot cos(\theta-\phi)}
 \label{eq-mot}
@@
 
-Furthermore the third Kepler law, can be quickly obtained considering the area swept in one \textcolor{red}{orbital period} T (the time needed to complete a full round of the orbit) $A=\pi ab$.
+Furthermore, the third Kepler law can be quickly obtained considering the area swept in one \textcolor{red}{orbital period} T (the time needed to complete a full round of the orbit) $A=\pi ab$.
 Since we know that this area is equal to $hT/2$ and $h^{2}=\mu a(1-e^{2})$ \cite{murray1999solar}:
@@
 
-We see that the two orbital parameters $T'$ and $a'$ are independent with respect to orbiting mass.
+We see that the two orbital parameters $T'$ and $a'$ are independent to orbiting mass.
 This statement can be extended to the other orbital parameters by considering the previous approximation $m$,$m'<<m_{c}$.
-This is why we expect that the mass/volume of the asteoroid can not be conditionally dependant to the orbital parameters.
+This is why we expect that the mass/volume of the asteroid can not be conditionally dependant on the orbital parameters.
 It is useful to define also the \textcolor{red}{mean motion} (feature that is also present in the asteroids dataset) as \cite{murray1999solar}:
@@
 Such boundaries should be intended as periodic for multiple of the orbital period T.
-The geometrical interpretation of the angle associated to the mean anomaly is given in Fig.
+The geometrical interpretation of the angle associated with the mean anomaly is given in Fig.
 \ref{Mean_anomaly}.
-It can be proven (see \cite{murray1999solar}) that the value of this angle, which describe the position of the orbiting item, is given by the following expression, know as Kepler equation \cite{murray1999solar}:
+It can be proven (see \cite{murray1999solar}) that the value of this angle, which describes the position of the orbiting item, is given by the following expression, known as Kepler equation \cite{murray1999solar}:
 
@@
 
-Finally as move on to space, other two angles are necessary for the description of an orbit: these are shown in Fig.
+Finally, as one move to space, other two angles are necessary for the description of an orbit: these are shown in Fig.
 \ref{Inclination} and are the \textcolor{red}{inclination} of the orbit (I) and the \textcolor{red}{longitude of the ascending node} $\Omega$.
@@
 The luminosity of an asteroid can be quantified with the concepts of magnitude.
-It is worth pointing out that asteroids have no intrinsic luminosity, but instead reflects the radiation, usually not uniformly since they have no atmosphere, of stars or other celestial bodies with intrinsic luminosity.
-First of all, before defining the concept of magnitude is useful to provide the concept of radiation flux.
-In our case it can be though as the number of photons \footnote{One can think, at first approximation, photons as packet of energy associated to the emitted light.
-A more formal and complete description can be found in \cite{feynman2018feynman} } that moves across a sphere centred on the light source and with radius equal to the observer distance \cite{burbine2016asteroids}:
+It is worth pointing out that asteroids have no intrinsic luminosity but instead reflects the radiation, usually not uniformly, since they have no atmosphere of stars or other celestial bodies with intrinsic luminosity.
+First of all, before defining the concept of magnitude is helpful to provide the concept of radiation flux.
+In our case, it can be thought of as the number of photons \footnote{One can think, at first approximation, photons as a packet of energy associated with the emitted light.
+A more formal and complete description can be found in \cite{feynman2018feynman} } that moves across a sphere centred on the light source and with a radius equal to the observer distance \cite{burbine2016asteroids}:
 
@@
 
-As one can note there is a dependence from $1/r^{2}$ which is given by the fact that the electromagnetic radiation, as for gravitational force, is spherically symmetric\footnote{A formal argument of this point can be found in \cite{zee2013einstein}}.
+As one can note, there is a dependence from $1/r^{2}$ which is given by the fact that the electromagnetic radiation, as for gravitational force, is spherically symmetric\footnote{A formal argument of this point can be found in \cite{zee2013einstein}}.
 On this basis, it is possible to define the relative magnitude as \cite{burbine2016asteroids}:
@@
 
-where $d$ is the light source distance in parsec. Rearranging the terms we have \cite{burbine2016asteroids}:
+where $d$ is the light source distance in parsec. Rearranging the terms, we have \cite{burbine2016asteroids}:
 
@@
 
-The solar system is composed not only by planets and the Sun, but also, as shown in Fig.
-\ref{Orbit_asteorids}, by a plethora of small bodies\footnote{Small with respect to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}.
-In particular, if the orbit of these items is near less then 1.3 AU such items are called near-Earth object (NEO).
-Such items are of three type: comets, meteoroids and asteroids.
-The first one are icy bodies that moving near the sun melts and release gases giving coloured tails.
-Meteroids are small \footnote{With respect to asteroids} rocky/metallic items with a diameter less than 1 meter.
+The solar system is composed not only of planets and the Sun but also, as shown in Fig.
+\ref{Orbit_asteorids}, by a plethora of small bodies\footnote{Small to the size of planets} whose orbit can be near to the Earth \cite{burbine2016asteroids}.
+In particular, if the orbit of these items is near less than 1.3 AU, such items are called near-Earth objects (NEO).
+Such items are of three types: comets, meteoroids and asteroids.
+The first ones are icy bodies that, as they move near the Sun, melt and release gases giving coloured tails.
+Meteoroids are minor \footnote{To asteroids} rocky/metallic items with a diameter less than 1 meter.
 Asteroids, on the other side, are items with a diameter larger than 1 m \cite{burbine2016asteroids}.
 The first two bodies are excluded from the present analysis.
-The near asteroids, which are mostly located on the sites reported in Fig.
+The near asteroids, which are mainly located on the sites reported in Fig.
 \ref{heic1715c}, are classified, as shown in Fig.
@@
 
-Beside this classification, there is the one which involve the hazardousness of an asteroid \cite{nasa_classification}:
+Besides this classification, there is the one that involve the hazardousness of an asteroid \cite{nasa_classification}:
 

### 2021-11-29 Manuscript/Script_FINAL.R

@@ -135,6 +135,13 @@ mbG<-stepw(model=bf,data=Asteroids_FINAL_double)
 plot(mbG,cex.vert.label=0.6,numIter=6000,col.labels=c("red"),vert.hl=c(21),col.hl=c("blue"),energy=TRUE)
 
 
+#mmod model
+mmod_model<-mmod(~.^1.,data=Asteroids_FINAL[,2:22])
+mmod_model_st<-stepwise(mmod_model,k=log(nrow(Asteroids_FINAL[,2:22])),direction="forward",details=1)
+plot(as(mmod_model_st ,"igraph"),vertex.color="red",vertex.label.dist=2,alpha=TRUE,vertex.label.cex=0.5,vertex.size=10,edge.color="black",edge.size=2,layout=layout_with_fr(as(mmod_model_st,'igraph')), vertex.label.family='Helvetica')
+
+
+
 #Mixed interaction analysis with mgm
 
 

### 2021-11-29 Manuscript/manuscript.tex
PUB

@@
 \section*{Abstract} The machine learning algorithms provide a promising approach to classify and predict natural or social phenomena.
-Differently to the oldest theoretical approach in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynmanlectures}), the machine learning algorithms start from very general assumptions/expressions and then their parameters, are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}.
+Differently to the theoretical approach, in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., a symmetry, the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynmanlectures}), the machine learning algorithms start from very general assumptions/expressions and then their parameters, are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}.
 The first approach allows obtaining interpretable predictions, the second one good forecast with a reduced effort.
@@
 The toll that has to be paid in this case is that, since the phenomena are much more complicated than the starting principles, a consistent part of them may be excluded by the assumption made at the beginning.
-The drawback of this approach is that the underlying mechanism by which a forecast is preferred to another one may not be interpretable.
 Furthermore, the calculation of the solution given the dynamical equation can be computationally expensive\footnote{This is the case, for instance, of the Hartree-Fock method or Density Functional Theory: such methods rewrite, with some approximations, the Schroedinger equation into a much more computationally affordable way.
@@
 Therefore they do not require the development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation.
+The drawback of this approach is that the underlying mechanism by which a forecast is preferred to another one may not be interpretable.
 Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century, astronomers knew that Mercury follows an elliptical orbit that, instead of being fixed, rotates around the Sun as shown in Fig.
@@
 This theory explains the connections and the meaning of the different features contained in the dataset.
+Finally, the code used to produce the results here reported if provided in Appendix B.
 
@@
 \chapter{Theoretical Framework}
+
+In this section, we are going to review the theoretical concepts
+that underlie the probabilistic methods here used: these will be
+exposed following the approaches of Murphy \cite{murphy2012machine} , Koller
+et al.
+\citep{koller2009probabilistic}, Højsgaardand et al.
+\citep{hojsgaard2012graphical} and Russel et al.
+\citep{russell2010artificial}.
+Furthermore, a rapid overview of the main concepts of information theory will also be provided following the Cover \citep{cover2006elements}, and MacKay \citep{mackay2003information} approaches.
+This is because some of its concepts in the preliminary analysis of the dataset were used.
+Finally, we concluded this chapter with a rapid overview of algorithm interpretability, which, together with the accuracy and the
+confusion matrix, is necessary for the assessment of a ML performances. On the other side, the concepts related to the celestial
+mechanics here used will be described, following the Murray
+approach \cite{murray1999solar} into the Appendix A.
+
+\section{Probabilistic models}
 
@@
 \section{Information theory}
-Given an ensemble of random variables, we can quantify the amount of information that one variable contains: such quantity is called mutual information and is a key concept within information theory.
+Given an ensemble of random variables, we can quantify the amount of information that one variable contains: such quantity is called mutual information and is a crucial concept within information theory.
 This approach, which was implemented by Claude Shannon decades before the probabilistic modelling, represents a complementary way by which one can attack the problem of conditional dependence between random variables.
+Furthermore, as shown in Fig.
+\ref{Information_theory_connections}, this theory provides a formidable contribution to different scientific fields.
 Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allow defining the concept of mutual information properly.
@@
 \end{equation}
+The concepts here reviewed, and the relations that interconnect them can be represented as shown in Fig.
+\ref{Entropy_MI}.
 Finally, we would report the data process inequality theorem that connects the information theory with the Markov chain: if we have a Markov chain, $X\rightarrow Y \rightarrow Z$  then $I(X;Y)\geq I(X;Z) $.
@@
 
-\section{Algorithm intepretability}
+\section{Algorithm interpretability}
 
@@
 \ref{ML_intepretability}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Information_theory_connections.jpg}
+\caption{The connections of information theory with different scientific fields. Image taken from \cite{cover2006elements}}
+\label{Information_theory_connections}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Figures/Entropy_MI.png}
+\caption{The relations between the entropy, conditional entropy and mutual information. Image taken from \cite{cover2006elements}}
+\label{Entropy_MI}
+\end{center}
+\end{figure}
 
@@
 
-\section{Probabilistic models}
-
-The analysis with the graphical models was started by inspecting the relations between continuous variables in the dataset.
+\section{Probabilistic models} The analysis with the graphical models was started by inspecting the relations between continuous variables in the dataset.
 Therefore, in this first step, the binary variable \textit{Hazardous} was excluded.
@@
 
-\section{Machine learning algorithms}
-
-The performances of mgm model will be now compared with the obtained from other ML algorithms.
-These include the random-forest (as implemented in the RandomForest package \cite{rfor}), the support vector machines (as implemented in the e1071 package \cite{dimitriadou2008misc}) , the quadratic discriminant analysis (as implemented in the MASS package \cite{MASS} and the logistic regression (as implemented in stats package \cite{stats}).
+\section{Machine learning algorithms} The performances of mgm model will be now compared with the one obtained from other ML algorithms.
+These include the random-forest (RF, as implemented in the RandomForest package \cite{rfor}), the support vector machines (SVM, as implemented in the e1071 package \cite{dimitriadou2008misc}), the quadratic discriminant anchiedersialysis (QDA, as implemented in the MASS package \cite{MASin reverseS} and the logistic regression (as implemented in stats package \cite{stats}).
 Their performances are reported in Fig.
@@
 This comparison shows that the RF and the SVM outperform the mgm graphical method while the logistic one has similar performances.
-Thus, one can ask the advantage of using a graphical method instead of a random forest or an SVM since its performances seem lower.
-The answer is the interpretability of the model provided: the random forest, at least, can provide variable rank importance as reported in Fig.
-\ref{RF_Importance} (note that apart from the Min Orbit intersection, there is a slight reshuffling in the feature importance to the one established by the MI in fig \ref{Mutual_information}; however apart from this is a black-box as the SVM, the logistic regression and the QDA.
-The probabilistic graph, providing conditional independences, gives the user an interpretable model whose properties can also be compared, discussed, and validated with the theory.
-Thus the model developed in this way allows a more scientific evaluation of the black-box ones.
+Thus, one can wonder about the advantage of using a graphical method instead of a random forest or an SVM since its performances seem lower.
+The answer is the interpretability of the model provided: the RF, at least, can provide variable rank importance as reported in Fig.
+\ref{RF_Importance} (note that apart from the Min Orbit intersection, there is a slight reshuffling in the feature importance to the one estain reverseblished by the MI in fig \ref{Mutual_information}.
+However, apart from this intepretation, the RF is a black-box as the SVM, the logistic regression and the QDA.
+Conversely, the probabilistic graphs, providing conditional independences, give the user an interpretable model whose properties can also be compared, discussed, and validated with the theory.
+Thus the model developed in this way allows a more scientific evaluation with respect to the black-box ones.
 In the author's view, this characteristic compensates for the lack of predictive power regarding the RF or the SVM methods.
@@
   \end{minipage}%%
-\caption{ROC curves for a selected set of ML algorithms. Plot obtained with caret and ggplot2 package \cite{kuhn2008building},ggplot2}}
+\caption{ROC curves for a selected set of ML algorithms. Plot obtained with the caret and the ggplot2 package \cite{kuhn2008building},ggplot2}}
 \label{ROC_ML}
@@
 \includegraphics[width=1\textwidth]{Figures/RF_Importance.pdf}
-\caption{Variable importance according to the random forest algorithm as implemented in \cite{rfor} package. Plot obtained with ggplot2 package \cite{ggplot2}}
+\caption{Variable importance according to the random forest algorithm as implemented in \cite{rfor} package. Plot obtained with the ggplot2 package \cite{ggplot2}}
 \label{RF_Importance}

### 2021-11-30 Manuscript/manuscript.tex
PUB

@@
 For this dataset, indeed, the relationships between the various features are known.
-The present work is organized as follows: first, a brief introduction about the basic theoretical concepts of the statistical methods here used will be provided, then the dataset will be described, and finally, the principal results here obtained with graphical methods will be presented as compared with other types of machine learning algorithms.
+This paper is organized as follows: first, a brief introduction about the basic theoretical concepts of the statistical methods here used will be provided, then the dataset will be described, and finally, the principal results here obtained with graphical methods will be presented as compared with other types of machine learning algorithms.
 Furthermore, a recap of the celestial mechanics will be provided in Appendix A.
@@
 
-In this section, we are going to review the theoretical concepts
+In this section, I am going to review the theoretical concepts
 that underlie the probabilistic methods here used: these will be
@@
 This is because some of its concepts in the preliminary analysis of the dataset were used.
-Finally, we concluded this chapter with a rapid overview of algorithm interpretability, which, together with the accuracy and the
+Finally, I concluded this chapter with a rapid overview of algorithm interpretability, which, together with the accuracy and the
 confusion matrix, is necessary for the assessment of a ML performances. On the other side, the concepts related to the celestial
@@
 \ref{Information_theory_connections}, this theory provides a formidable contribution to different scientific fields.
-Here we would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allow defining the concept of mutual information properly.
+Here I would provide some basic concepts of this theory, following the Cover \cite{cover2006elements} and MacKay \cite{mackay2003information} approaches, that allow defining the concept of mutual information properly.
 The founding concept of information theory is entropy.
@@
 \ref{tab_features}.
-Since only 16 \% of the asteroids were hazardous, we considered reducing the number of the non-hazardous to improve the portion of the hazardous ones.
-For this purpose, we constructed a database in which the proportion hazardous/not hazardous was 1:5: thus, all the hazardous ones were included, while the not-hazardous were randomly selected \footnote{The author is aware that in principle, the proportion should be much less unbalanced, however, the tool that has to be paid for this operation is an overall reduction of the cases in the dataset that lowers the performances of the algorithms here used.
+Since only 16 \% of the asteroids were hazardous, I considered reducing the number of the non-hazardous to improve the portion of the hazardous ones.
+For this purpose, I constructed a database in which the proportion hazardous/not hazardous was 1:5: thus, all the hazardous ones were included, while the not-hazardous were randomly selected \footnote{The author is aware that in principle, the proportion should be much less unbalanced, however, the tool that has to be paid for this operation is an overall reduction of the cases in the dataset that lowers the performances of the algorithms here used.
 The proportion here is used as a trade-off for these two contrasting facts}.
@@
 Thus it can be said that the model produced is almost consistent with the astronomic laws.
-On this basis, the performances of the model in terms of confusion matrix, ROC curve and $\phi$ parameters can be evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
+On this basis, the performances of the model in terms of confusion matrix, ROC curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) can be evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
 The confusion matrix obtained is reported in Fig.

### 2021-11-30 Manuscript/manuscript.tex
PUB

@@
 \usepackage[utf8]{inputenc}
+\usepackage{csquotes}
 
@@
 \section*{Abstract} The machine learning algorithms provide a promising approach to classify and predict natural or social phenomena.
-Differently to the theoretical approach, in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., a symmetry, the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynmanlectures}), the machine learning algorithms start from very general assumptions/expressions and then their parameters, are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}.
+Differently to the theoretical approach, in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., a symmetry, the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynmanlectures}), the machine learning algorithms start from very general assumptions/expressions and then their parameters are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}.
 The first approach allows obtaining interpretable predictions, the second one good forecast with a reduced effort.
@@
 Since they provide the list of conditional independences/dependencies, the graphical methods represent a good compromise between the need for interpretability and the forecasts obtained without developing a general theory.
-In order to prove the validity of this statement, I considered a dataset for which the theory that interconnects its feature was known: in particular, I chose the asteroids hazardousness as provided by CNEOS \cite{cneos+nasa} and published on Kaggle \cite{kaggle_dataset}.
+In order to prove the validity of this statement, I considered a dataset for which the theory that interconnects its features was known: in particular, I chose the asteroids hazardousness as provided by CNEOS \cite{cneos+nasa} and published on Kaggle \cite{kaggle_dataset}.
 The outcomes of the probabilistic methods for predicting the asteroid hazardousness were compared with the ones provided by the theory and with the ones of Random Forest (RF), Support Vector Machines (SVM), and Logistic Regression and Quadric Discriminant analysis (QDA).
@@
 Therefore they do not require the development of a specific theory as the \textit{ab-initio} one: the same model can be applied to exoplanets habitability as well to credit risk evaluation.
-The drawback of this approach is that the underlying mechanism by which a forecast is preferred to another one may not be interpretable.
+The drawback of this approach is that the underlying mechanism,  by which a forecast is preferred to another one, may not be interpretable.
 Such dichotomy between these two methods can be explained with an example taken from astronomy: at the beginning of XX century, astronomers knew that Mercury follows an elliptical orbit that, instead of being fixed, rotates around the Sun as shown in Fig.
@@
 In this way, one can evaluate the quality of the forecasts and the model provided by the graphical methods.
-The present work aims to perform such analysis for a dataset that contain the features and the hazardousness \footnote{For Earth} of asteroids.
+The present work aims to perform such analysis for a dataset that contains the features and the hazardousness \footnote{For Earth} of asteroids.
 For this dataset, indeed, the relationships between the various features are known.
@@
 Furthermore, a rapid overview of the main concepts of information theory will also be provided following the Cover \citep{cover2006elements}, and MacKay \citep{mackay2003information} approaches.
-This is because some of its concepts in the preliminary analysis of the dataset were used.
+This is because some of its quantities in the preliminary analysis of the dataset were used.
 Finally, I concluded this chapter with a rapid overview of algorithm interpretability, which, together with the accuracy and the
@@
 \end{equation}
-Where V is the number of variables and 1:V stands for ${1,2,...,V}$.
+Where V is the number of variables and $1:V$ stands for ${1,2,...,V}$.
 This decomposition makes explicit the conditional probability tables, or in other terms, the transition probability tensors \cite{wu2017markov}.
@@
 
-In the first term $\textbf{y}$ is fixed to its observed values while in the second it is free. Such expression can be recasted in to a more explicative form \cite{murphy2012machine}:
+In the first term $\textbf{y}$ is fixed to its observed values while in the second it is free. Such expression can be recast in to a more explicative form \cite{murphy2012machine}:
 
@@
 
-Therfore at the optimum we will have \cite{murphy2012machine}:
+Therefore at the optimum we will have \cite{murphy2012machine}:
 
@@
 \ref{Entropy_MI}.
-Finally, we would report the data process inequality theorem that connects the information theory with the Markov chain: if we have a Markov chain, $X\rightarrow Y \rightarrow Z$  then $I(X;Y)\geq I(X;Z) $.
+Finally, I would report the data process inequality theorem that connects the information theory with the Markov chain: if we have a Markov chain, $X\rightarrow Y \rightarrow Z$  then $I(X;Y)\geq I(X;Z) $.
 As for the Gibbs inequality, the underlying idea is that no clever manipulation of the data can improve the inference that can be made on them \cite{cover2006elements,mackay2003information}.
@@
 Besides the accuracy and other features connected with the confusion matrix, another critical feature is algorithm interpretability.
-This feature represents how much the algorithm explains its predictions and, in general, its mechanics.
+This concept represents how much the algorithm explains its predictions and, in general, its mechanics.
 In other words, how much the algorithm is not a black box.
@@
 This is something that is nested both in formal systems as algorithms or scientific theories.
-For instance, let us consider the fall of object: at first glance, one can consider the only the items on Earth, in this case, the acceleration is constant, and it is a straightforward theory; but as one moves to consider also the interaction within planets a much more complicate law should be considered (in which the previous case is a particular one).
+For instance, let us consider the fall of object: at first glance, one can consider only the items on Earth.
+In this case, the acceleration is constant, and a straightforward theory will be obtained.
+As one moves to consider also the interaction within planets and stars a much more complicate law should be considered (in which the previous case is a particular one).
 The first one is fully explainable but poorly complete; the second one is more difficult to explain but much more complete.
+This is something that was elegantly stated by A.Einstein \cite{physics-reality}:
+\begin{displayquote}
+\textit{Science uses the totality of the primary concepts, i.e., concepts directly connected with sense experiences, and propositions connecting them.
+In its first stage of development, science does not contain anything else.
+Our everyday thinking is satisfied on the whole with this level.
+Such a state of affairs cannot, however, satisfy a spirit which is really scientifically minded; because the totality of concepts and relations obtained in this manner is utterly lacking in logical unity.
+In order to supplement this deficiency, one invents a system poorer in concepts and relations, a system retaining the primary concepts and relations of the “first layer” as logically derived concepts and relations.
+This new “secondary system” pays for its higher logical unity by having elementary concepts (concepts of the second layer), which are no longer directly connected with complexes of sense experiences.
+Further striving for logical unity brings us to a tertiary system, still poorer in concepts and relations, for the deduction of the concepts and relations of the secondary (and so indirectly of the primary) layer.
+Thus the story goes on until we have arrived at a system of the greatest conceivable unity, and of the greatest poverty of concepts of the logical foundations, which is still compatible with the observations made by our senses.
+We do not know whether or not this ambition will ever result in a definitive system.
+If one is asked for his opinion, he is inclined to answer no.
+While wrestling with the problems, however, one will never give up hope that this greatest of all aims can really be attained to a very high degree [...]The essential thing is the aim to represent the multitude of concepts and propositions, close to experience, as propositions, logically deduced from a basis, as narrow as possible, of fundamental concepts and fundamental relations which themselves can be chosen freely (axioms).
+The liberty of choice, however, is of a special kind; it is not in any way similar to the liberty of a writer of fiction.
+Rather, it is similar to that of a man engaged in solving a well-designed word puzzle.
+He may, it is true, propose any word as the solution; but, there is only one word which really solves the puzzle in all its parts.
+It is a matter of faith that nature– as she is perceptible to our five senses– takes the character of such a well-formulated puzzle.
+The successes reaped up to now by science do, it is true, give a certain encouragement for this faith.}
+\end{displayquote}
 Therefore, in principle, the evaluation of algorithm performances should be considered not only into one dimension, the accuracy but also by considering its interpretability as shown in Fig.
@@
 Thus the features obtained were therefore reduced to 22.
-Their description will be postponed to Appendix A since their description cannot be disjoint with celestial mechanics concepts.
+Their description will be postponed to Appendix A since tit cannot be disjoint with celestial mechanics concepts.
 Their enumeration is provided in Tab.
@@
 
-This chapter is organized in the following way: first, we will report the results concerning the preliminary analysis performed on the dataset.
+This chapter is organized in the following way: first, I will report the results concerning the preliminary analysis performed on the dataset.
 The factor analysis of mixed data and the mutual information analysis of the continuous variables of asteroids vs their hazard will be reported.
-Then it will follow the analysis of the dataset performed with the probabilistic methods.
+Then it will follows the analysis of the dataset performed with the probabilistic methods.
 After a preliminary analysis on the continuous variables, the mixed interaction model and the minforest model obtained for the whole dataset will be presented and discussed.
@@
 \ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig.
-\ref{FAMD_Quantitative variables} the different correlations that are given by the celestial mechanics laws can be recognized: for instance, see that there is strong correlation of the mean motion with the semi-major axis.
+\ref{FAMD_Quantitative variables} the different correlations that are given by the celestial mechanics laws can be recognized: for instance, see that there is strong anti-correlation of the mean motion with the semi-major axis.
 This is due to the Kepler law discussed in Appendix A.
@@
 et al approach \cite{peng2005feature}.
-We see that the first place in the mutual information ranking, looking to the diagonal element, is taken by minimum orbit intersection: this is correct since this is one parameter used by NASA for deciding if an asteroid is hazardous or not (see Appendix A).
+It can be seen that the first place in the mutual information ranking, looking to the diagonal element, is taken by minimum orbit intersection: this is correct since this is one parameter used by NASA for deciding if an asteroid is hazardous or not (see Appendix A).
 The second place is occupied by Epoch date close approach.
 The third place is the eccentricity value: this parameter is entangled with the minimum orbit intersection and thus it is reasonable that it is important.
-Then we have two parameter that are related to the dimension of the asteroid: this is meaningful since if the asteroid has a too reduced volume it will be destroyed by the Earth atmosphere.
-On the other side the other parameters seems to have a too low MI for being interesting in this preliminary analysis.
-The results obtained for FAMD and MI will provide a useful path-guide, together with the celestial mechanics theory, for the interpretation of the results obtained in the following section
+Then two parameters that are related to the dimension of the asteroid are present: this is meaningful since if the asteroid has a too reduced volume it will be destroyed by the Earth atmosphere.
+On the other hand the other parameters seems to have a too low MI for being interesting in this preliminary analysis.
+The results obtained for FAMD and MI will provide a useful path-guide, together with the celestial mechanics theory, for the interpretation of the results obtained in the following section.
 
@@
 Therefore, in this first step, the binary variable \textit{Hazardous} was excluded.
-For this purpose, among the different methods available, we considered the \textit{Graphical least absolute shrinkage and selection operator} GLASSO as implemented in the \textit{glasso} R package \cite{friedman2008sparse,glasso}.
+For this purpose, among the different methods available, I considered the \textit{Graphical least absolute shrinkage and selection operator} GLASSO as implemented in the \textit{glasso} R package \cite{friedman2008sparse,glasso}.
 After different tests reported in Fig.
-\ref{GLASSO_convergence}, we considered as the final result the graph obtained with the value of $\rho$ (the one that penalizes further connections) equal to $0.3$.
+\ref{GLASSO_convergence}, I considered as the final result the graph obtained with the value of $\rho$ (the one that penalizes further connections) equal to $0.3$.
 Such choice is motivated by the fact that the conditional dependences/independences stated by the Celestial Mechanics are correctly reproduced with this parameter.
-For instance, we see that the diameter of the asteroids is independent of all features related to their motion.
+For instance, it can be seen that the diameter of the asteroids is independent of all features related to their motion.
 Such a result is meaningful since the volume/mass of the asteroids is largely lower than the mass of the earth.
 Thus, as explained in Appendix A, there is no way the asteroid orbit can be modified by its mass.
-In addition, we see that the Close approach Date and the Epoch date are linked to each other but in no way from the other features.
+In addition, it can be seen that the Close approach Date and the Epoch date are linked to each other but in no way from the other features.
 This result is correct since the date and epoch are set with an arbitrary scale.
@@
 Thus it can be said that the model produced is almost consistent with the astronomic laws.
-On this basis, the performances of the model in terms of confusion matrix, ROC curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) can be evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
+On this basis, the performances of the model in terms of confusion matrix, ROC curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) were evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
 The confusion matrix obtained is reported in Fig.

### 2021-11-30 Manuscript/manuscript.tex
PUB

@@
 
-\chapter{Appendix B: R code}
+\chapter{Appendix B: R code} Here I provide the code which I run to obtain the results previously showed.
+For its writing I took in consideration the examples provided in the  packages documentation, the Prof.Nicolussi lectures and he examples provided in the Hojsgaard textbook  \cite{hojsgaard2012graphical}.
 

### 2021-11-30 Manuscript/manuscript.tex
PUB

@@
 \textit{This day may possibly be my last: but the laws of probability, so true in general, so fallacious in particular, still allow about fifteen years. }\\Edward Gibbon (1737-1794)
+}
+
+\epigraph{
+\textit{In its efforts to learn as much as possible about nature, modern physics has found that certain things can never be “known” with certainty.
+Much of our knowledge must always remain uncertain.
+The most we can know is in terms of probabilities.}\\Richard P.
+Feynman (1918-1988)
 }
@@
 \ref{FAMD_Quantitative variables} the different correlations that are given by the celestial mechanics laws can be recognized: for instance, see that there is strong anti-correlation of the mean motion with the semi-major axis.
-This is due to the Kepler law discussed in Appendix A.
+This is due to the Kepler laws discussed in Appendix A.
 Furthermore we see that that the mean motion is correctly almost independent with respect to the the diameter (max or min) of  the asteroid.
@@
     \subcaption{ \begin{center}
-    d) Semi Major Axis
+    d) Absolute magnitude
     \end{center}}
@@
 
-Therefore \cite{murray1999solar}:
-
-\cite{murray1999solar}:
-
 We see that the two orbital parameters $T'$ and $a'$ are independent to orbiting mass.

### 2021-12-02 Manuscript/manuscript.tex
PUB

@@
 Differently to the theoretical approach, in which the laws that describe particular phenomena are derived from general assumptions using a mathematical language (e.g., a symmetry, the principle of energy conservation, the three Newton laws, the second principle of thermodynamics \cite{feynmanlectures}), the machine learning algorithms start from very general assumptions/expressions and then their parameters are optimized with the likehood maximization\cite{russell2010artificial,murphy2012machine}.
-The first approach allows obtaining interpretable predictions, the second one good forecast with a reduced effort.
+The first approach allows obtaining interpretable predictions, the second one good forecasts with a reduced effort.
 However, the toll that has to be paid in this case is the low interpretability \cite{russell2010artificial,murphy2012machine}.
@@
 
-\chapter{Introduction} The description and the forecast of physical \footnote{Where here physical should be intended as mesuarable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms, one can formulate theoretical models that provide equations that describe such phenomena.
+\chapter{Introduction} The description and the forecast of physical \footnote{Where here physical should be intended as measurable} phenomena can be attacked with two different approaches: starting from a restricted set of principles/axioms, one can formulate theoretical models that provide equations that describe such phenomena.
 Such approach will be called here \textit{ab-initio}.
@@
 Finally, to reduce the computational cost,  one can consider only the pairwise interaction instead of the maximum clique.
-This is the analogue of what is usually performed in solid-state physics (but not always)  when only the interaction between the first neighbour is considered.
+This is the analogue of what is usually performed in solid-state physics (but not always)  when only the interaction between the first neighbour atoms is considered.
 Another example is the  Ising model: here we have a lattice of spins that can be or in $\ket{+}$ or in $\ket{-}$ and their interaction is modelled by\cite{murphy2012machine}:
@@
 
-(this last expression can be reconducted to the multivariate gassian if one consider $\boldsymbol{\Delta}=\boldsymbol{\Sigma}^{-1}$ and $\boldsymbol{\eta}=\boldsymbol{\Delta}\boldsymbol{\mu}$.
-Given the network, we would now move on how the parameters can be achieved.
+(this last expression can be reconducted to the multivariate Gaussian if one consider $\boldsymbol{\Delta}=\boldsymbol{\Sigma}^{-1}$ and $\boldsymbol{\eta}=\boldsymbol{\Delta}\boldsymbol{\mu}$.
+Given the network, we would now move on how the parameters can be calculated.
 Lets start from a Markov random field in log-linear form \cite{murphy2012machine}:
@@
 Apart from all legal problems connected with an algorithm that is a blackbox\footnote{Think for instance the conditions imposed by the GDPR (e.g.
-right to explanation)}, in the author opinion, results provided by such an algorithm cannot be considered scientific since the scientific method also requires to explain and not simply almost correct forecast.
+right to explanation)}, in the author opinion, results provided by such an algorithm cannot be considered scientific: the scientific method also requires to explain and not simply to provide correct forecast.
 Following the Tarski et al.
@@
 In this case, the acceleration is constant, and a straightforward theory will be obtained.
-As one moves to consider also the interaction within planets and stars a much more complicate law should be considered (in which the previous case is a particular one).
+As one moves to consider also the interaction within planets and stars, a much more complicate law should be considered (in which the previous case is a particular one).
 The first one is fully explainable but poorly complete; the second one is more difficult to explain but much more complete.
@@
 \end{displayquote}
-Therefore, in principle, the evaluation of algorithm performances should be considered not only into one dimension, the accuracy but also by considering its interpretability as shown in Fig.
-\ref{ML_intepretability}
+Therefore the evaluation of algorithm performances should be considered not only into one dimension, the accuracy but also by considering its interpretability as shown in Fig. \ref{ML_intepretability}
 
@@
 
-The asteroid dataset was retrieved from Kaggle \cite{kaggle_dataset}, which reports into a more user-friendly mode the dataset of The Center for Near-Earth Object Studies (CNEOS) \cite{cneos+nasa}, a NASA research centre.
+The asteroid dataset was retrieved from Kaggle \cite{kaggle_dataset}, which reports into a more machine readable form the dataset of The Center for Near-Earth Object Studies (CNEOS) \cite{cneos+nasa}, a NASA research centre.
 Among the 40 features that were present in the dataset,  the following cases were excluded: the redundant features (e.g.
@@
 Since only 16 \% of the asteroids were hazardous, I considered reducing the number of the non-hazardous to improve the portion of the hazardous ones.
-For this purpose, I constructed a database in which the proportion hazardous/not hazardous was 1:5: thus, all the hazardous ones were included, while the not-hazardous were randomly selected \footnote{The author is aware that in principle, the proportion should be much less unbalanced, however, the tool that has to be paid for this operation is an overall reduction of the cases in the dataset that lowers the performances of the algorithms here used.
+For this purpose, I constructed a database in which the proportion hazardous/not hazardous was 1:5: thus, all the hazardous ones were included, while the not-hazardous were randomly extracted \footnote{The author is aware that in principle, the proportion should be much less unbalanced, however, the tool that has to be paid for this operation is an overall reduction of the cases in the dataset that lowers the performances of the algorithms here used.
 The proportion here is used as a trade-off for these two contrasting facts}.
@@
 \ref{Density_relevant}.
-Next we moved on a more systematic analysis with FADM and mutual information.
+It can be seen that there is an evident distinction for the min orbit intersection, perihelion distance and eccentricity (note that these three parameters are the different faces of the same medal).
+On the other side, concerning the Absolute magnitude, the distinction seems not so clear.
+This result, together with the overlap in the three previous features, is because, according to the CNEOS definition (see Appendix A), a hazardous asteroid should have a min orbit intersection lower than a fixed parameter and an absolute magnitude higher to a certain threshold.
+The fulfilment of both these conditions makes the asteroid dangerous.
+Next I moved on a more systematic analysis with FADM and mutual information.
 In Fig.
 \ref{FADM},\ref{FAMD_Quantitative variables} and \ref{FAMD_Individuals_(c)} are reported the main achivements of FADM (performed with FactoMineR package \cite{le2008factominer}): in particular from the correlation circle in Fig.
-\ref{FAMD_Quantitative variables} the different correlations that are given by the celestial mechanics laws can be recognized: for instance, see that there is strong anti-correlation of the mean motion with the semi-major axis.
+\ref{FAMD_Quantitative variables} the different correlations that are given by the celestial mechanics laws can be recognized.
+For instance, see that there is strong anti-correlation of the mean motion with the semi-major axis.
 This is due to the Kepler laws discussed in Appendix A.
 Furthermore we see that that the mean motion is correctly almost independent with respect to the the diameter (max or min) of  the asteroid.
-As explained in Appendix A this is an another result of Newton gravitation theory for a two body interaction: the mean motion of an asteroid, as long as the the mass of it is much lower with respect to the sun, is independent from its mass.
-On the other side the fact that relative velocity is correlated with the diameter is a spurios correlation.
-At this point one can ask why the mean motion and the relative velocity are orthogonal: this point will be clarified from a theoretical point of view in appendix A and also with graphical models.
-Briefly the motion of an object on an ellipse as seen from a focus is not uniform: this is faster as the two bodies approaches.
+As explained in Appendix A this is an another result of Newton gravitation theory for a two body interaction: the mean motion of an asteroid, as long as the the mass of it is much lower with respect to the earth, is independent from its mass.
+On the other side the fact that relative velocity is correlated with the diameter is a spurious correlation.
+At this point one can ask why the mean motion and the relative velocity are orthogonal: this outcome will be clarified from a theoretical point of view in appendix A and also with graphical models.
+Briefly the motion of an object on an ellipse as seen from a focus (the Earth) is not uniform: this is faster as the two bodies approaches.
 Beside this inspection a further analysis based on the concept of mutual information (summarized in the previous chapter) was performed: its result is reported in Fig.
@@
 The third place is the eccentricity value: this parameter is entangled with the minimum orbit intersection and thus it is reasonable that it is important.
-Then two parameters that are related to the dimension of the asteroid are present: this is meaningful since if the asteroid has a too reduced volume it will be destroyed by the Earth atmosphere.
-On the other hand the other parameters seems to have a too low MI for being interesting in this preliminary analysis.
-The results obtained for FAMD and MI will provide a useful path-guide, together with the celestial mechanics theory, for the interpretation of the results obtained in the following section.
+Then two parameters that are related to the dimension of the asteroid are present: the Estimated min diameter and Absolute Magnitude.
+This fact is meaningful since if the asteroid has a too reduced volume it will be destroyed by the Earth atmosphere.
+Indeed the Absolute magnitude and the Min orbit intersection are the features used by CNEOS to classify if an asteroid is hazardous or not (see Appendix A).
+On the other hand the other parameters seems to have a too low MI score for being interesting in this preliminary analysis.
 
@@
 Thus it can be said that the model produced is almost consistent with the astronomic laws.
-On this basis, the performances of the model in terms of confusion matrix, ROC curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) were evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
+On this basis, the performances of the model in terms of confusion matrix, ROC (Receiver operating characteristic) curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) were evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
 The confusion matrix obtained is reported in Fig.
@@
 \includegraphics[width=1\textwidth]{Figures/ROC_mgm.pdf}
-\caption{The ROC curve as obtained from the ROCR package \cite{sing2005rocr} and ggplot2 \cite{ggplot2}. The corresponding $\phi$ value is $0.6$ } 
+\caption{The ROC (Receiver operating characteristic) curve as obtained from the ROCR package \cite{sing2005rocr} and ggplot2 \cite{ggplot2}. The corresponding $\phi$ value is $0.6$ } 
 \label{ROC_mgm}
@@
 \begin{table}[]
-\caption{$\phi$ factor for a selected set of ML algorithms as compared with the mgm}
+\caption{$\phi$ coefficient (also known as Matthews correlation coefficient ) for a selected set of ML algorithms as compared with the mgm}
 \begin{center}
@@
 
-\chapter{Conclusions}
-
-The GLASSO (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that correctly capture the physical connections of the dataset feature.
+\chapter{Conclusions and Outlook}
+
+The GLASSO (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that correctly reproduce the physical connections of the dataset feature.
 In particular, for both models, it can be seen that no one of the orbital parameters is conditionally dependent on the diameter of the asteroid, as stated by theory.
@@
 Such a feature compensates the mgm forecast performances gap.
+As a future outlook I consider to extend this analysis for physical phenomena for which there is no deterministic theory but only probabilistic one\footnote{Following the I.Prigogine \cite{prigogine2017non,prigogine1997end,nicolis1989exploring,prigogine1978time} argument it must be said that a pure deterministic theory, if exist, cannot be fully applied.
+Therefore, in principle, all possible theories should be considered probabilistic.
+This is because the initial conditions can not be known with a infinite precision.
+Therefore, if one also perfectly knows the equation of motion and can also solve them exactly (and in most of cases it is not possible) a source of error still persist and it propagates with the Lyapunov exponent.
+However as it was shown by Prigogine it a system is at equilibrium this instability can neglected.
+In this case a deterministic approach is meaningful }.
+For instance one can consider Earthquakes and Vulcanos Eruptions \cite{sornette1989self,sparks2003forecasting,sornette2006critical}
 
@@
 \begin{equation}
-(a/a')^{3}\approx(T/T')^{2}
+\frac{m_{c}+m'}{m_{c}+m}\approx\dfrac{m}{m_{c}}=\left(a/a'\right)^{3}\left( T/T'\right)^{2}
 \end{equation}

### 2021-12-03 Manuscript/manuscript.tex
PUB

@@
 However, the toll that has to be paid in this case is the low interpretability \cite{russell2010artificial,murphy2012machine}.
-Since they provide the list of conditional independences/dependencies, the graphical methods represent a good compromise between the need for interpretability and the forecasts obtained without developing a general theory.
+Since they provide the list of connections within the features, the graphical methods represent a good compromise between the need for interpretability and the forecasts obtained without developing a general theory.
 In order to prove the validity of this statement, I considered a dataset for which the theory that interconnects its features was known: in particular, I chose the asteroids hazardousness as provided by CNEOS \cite{cneos+nasa} and published on Kaggle \cite{kaggle_dataset}.
@@
 The results show that the forecast performances of probabilistic methods are better than QDA and almost equal to the logistic regression but lower to RF and SVM.
-However, since the list of conditional dependencies correctly reflects the laws of celestial mechanics \cite{murray1999solar}, it can be said that in this case, the probabilistic methods provide an interpretable and correct explanation of their mechanism.
+However, since the list of connections correctly reflects the laws of celestial mechanics \cite{murray1999solar}, it can be said that in this case, the probabilistic methods provide an interpretable and correct explanation of their mechanism.
 Therefore, contrary to other machine learning algorithms, such methods can be fully validated scientifically.
@@
 The second approach was the one followed by A.Einstein with his general relativity theory (further details are given in the A.Zee textbook \cite{zee2013einstein}).
-The graphical methods represent a good compromise between these two approaches: they can be used for every physical phenomenon, and, since they provide conditional dependencies/independences, they also give an elegant and intuitive representation of their mechanisms.
+The graphical methods represent a good compromise between these two approaches: they can be used for every physical phenomenon, and, since they provide the connection list, they also give an elegant and intuitive representation of their mechanisms.
 Therefore, if the theory of the involved process is known, one can use this class of algorithms and compare its finding with the theory.
@@
 Given this formalism, we are interested in finding an intelligent way to plot such joint distribution intuitively: the graph theory answers this quest.
-In particular, the absence of an edge that interconnects the random variables can be represented by nodes and the presence of conditional independence for two random variables.
+The nodes can be used to represent the random variables while the presence or the lack of edges can be used to represent the conditional indipendence \cite{murphy2012machine}.
 Bayesian networks consider directed edges,  while Markov random fields (MRF) are only undirected.
@@
 \ref{GLASSO_convergence}, I considered as the final result the graph obtained with the value of $\rho$ (the one that penalizes further connections) equal to $0.3$.
-Such choice is motivated by the fact that the conditional dependences/independences stated by the Celestial Mechanics are correctly reproduced with this parameter.
-For instance, it can be seen that the diameter of the asteroids is independent of all features related to their motion.
+Such choice is motivated by the fact that the connections stated by the Celestial Mechanics are correctly reproduced with this parameter.
+For instance, it can be seen that the diameter of the asteroids is not connected to all features related to their motion.
 Such a result is meaningful since the volume/mass of the asteroids is largely lower than the mass of the earth.
@@
 Also, as expected, the Perihelion Epoch and Osculation Time are linked, but there is no dependence on the orbit parameters.
-Furthermore, the mean motion is conditionally independent with relative velocity, but correctly, this is dependent on perihelion distance.
+Furthermore, the mean motion is not connected directly to relative velocity, but correctly, both are connected to the perihelion distance.
 
@@
 \ref{mgm}.
-From this figure, it can be seen that the features that are conditionally dependant on the Hazardous features are: the absolute magnitude, the min orbit intersection, and the eccentricity.
+From this figure, it can be seen that the features that are directly connected to the Hazardous feature are: the absolute magnitude, the min orbit intersection, and the eccentricity.
 The theory supports all these connections.
@@
 Furthermore, the parameters related to the diameters are also not connected with the orbital parameters.
-In addition, the mean velocity is not directly related to the mean motion, but there are the orbital parameters within as expected.
+In addition, the relative velocity is not directly related to the mean motion, but there are the orbital parameters within as expected.
 Thus it can be said that the model produced is almost consistent with the astronomic laws.
@@
 However, apart from this intepretation, the RF is a black-box as the SVM, the logistic regression and the QDA.
-Conversely, the probabilistic graphs, providing conditional independences, give the user an interpretable model whose properties can also be compared, discussed, and validated with the theory.
+Conversely, the probabilistic graphs, providing the list of connections among the random variables, give the user an interpretable model whose properties can also be compared, discussed, and validated with the theory.
 Thus the model developed in this way allows a more scientific evaluation with respect to the black-box ones.
@@
 The GLASSO (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that correctly reproduce the physical connections of the dataset feature.
-In particular, for both models, it can be seen that no one of the orbital parameters is conditionally dependent on the diameter of the asteroid, as stated by theory.
+In particular, for both models, it can be seen that no one of the orbital parameters is conneted to the diameter of the asteroid, as stated by theory.
 Furthermore, the mgm model can provide forecasts about the hazardousness of the asteroids with almost equal or better performances than the logistic and QDA algorithms.

### 2021-12-03 Manuscript/manuscript.tex
PUB

@@
 In addition, the relative velocity is not directly related to the mean motion, but there are the orbital parameters within as expected.
-Thus it can be said that the model produced is almost consistent with the astronomic laws.
+Moving to the relation of orbital parameters, it can be seen that there is a negative relation between the eccentricity and perihelion distance: this is meaningful since it comes from the definition of eccentricity.
+The other side of the coin of this connection is the positive relationships within the semimajor axis and the eccentricity.
+Furthermore, the positive relationship between the mean motion and the semimajor axis is correct: this is the result of the 3$^{th}$ Kepler law.
+Finally the connections with the Jupiter Tisserand Invariant comes directly from the definition of this parameter.
+A full explanation of these connections is provided in the Appendix A.
+Thus it can be said that the model produced is consistent with the astronomic laws.
 On this basis, the performances of the model in terms of confusion matrix, ROC (Receiver operating characteristic) curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) were evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
@@
 This statement can be extended to the other orbital parameters by considering the previous approximation $m$,$m'<<m_{c}$.
-This is why we expect that the mass/volume of the asteroid can not be conditionally dependant on the orbital parameters.
+This is why we expect that the mass/volume of the asteroid can not be  dependant on the orbital parameters.
 It is useful to define also the \textcolor{red}{mean motion} (feature that is also present in the asteroids dataset) as \cite{murray1999solar}:

### 2021-12-07 Presentation/Presentation.tex
PUB

\documentclass{beamer}
%
% Choose how your presentation looks.
%
% For more themes, color themes and font themes, see:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
%

\mode<presentation>
{
  \usetheme{Darmstadt}      % or try Darmstadt, Madrid, Warsaw, ...
  \usecolortheme{beaver} % or try albatross, beaver, crane, ...
  \usefonttheme{default}  % or try serif, structurebold, ...
  \setbeamertemplate{navigation symbols}{}
  \setbeamertemplate{caption}[numbered]
}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{Pic/Intro.jpg}}
\title[Hazardous asteroids forecast via Markov random fields]{Hazardous asteroids forecast via Markov random fields}
\subtitle{Project for the exam: Probabilistic Modelling (DSE)}
\author{Marzio De Corato}
\date{\today}

\begin{document}

\begin{frame}
\vspace{+4 cm}  \titlepage
\end{frame}

\usebackgroundtemplate{ }

% Uncomment these lines for an automatically generated outline.
%\begin{frame}{Outline}
%  \tableofcontents
%\end{frame}

\section{Introduction}

\begin{frame}{Introduction}

%\begin{itemize}
Our lab has been working on analyzing data on mast cell cancer in Guiding Eyes for the Blind dogs using R programming. This presentation will illustrate the work we have done so far and what we hope to accomplish in the upcoming weeks.

%\end{itemize}

\end{frame}

\end{document}

### 2021-12-07 Presentation/Presentation.tex
PUB

@@
 }
+
+\usepackage[backend=bibtex]{biblatex}
+\addbibresource{bibliography.bib}
 
@@
 
-%\begin{itemize}
-Our lab has been working on analyzing data on mast cell cancer in Guiding Eyes for the Blind dogs using R programming.
-This presentation will illustrate the work we have done so far and what we hope to accomplish in the upcoming weeks.
+\begin{itemize}
+\item \textbf{Final Goal} Assessment of forecasts and interpretability for different machine learning algorithms, including the probabilistic models
+\item \textbf{Method} Use a dataset for which the laws that interconnects the different features are known from general principles
+\item \textbf{Dataset} CNEOS asteroids dataset for more than 3500 asteroids
+\item \textbf{Theoretical laws} Celestial mechanics
+\item \textbf{Algorithms involved - probabilistic models} GLASSO, mgm, minforest, mmod
+\item \textbf{Algorithms involved - others} Random forest, Support Vector Machines, Quadratic Discriminant Analysis, Logistic Regression
 
-%\end{itemize}
+\end{itemize}
 
@@
 
+\section{Celestial mechanics}
+
+\begin{frame}{Celestial mechanics}
+\begin{equation}
+\textbf{F}_{1}=\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{1}
+\end{equation}
+
+\begin{equation}
+\textbf{F}_{2}=-\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{2}
+\end{equation}
+
+If we consider the motion of the second item with respect to the first one
+
+\begin{equation}
+\ddot{\textbf{r}}=\ddot{\textbf{r}}_{2}-\ddot{\textbf{r}}_{1} \quad \mu=\mathcal{G}(m_{1}+m_{2})
+\end{equation}
+
+\begin{equation}
+\dfrac{d^{2}\textbf{r}}{dt^{2}}+\mu\dfrac{\textbf{r}}{r^{3}}=0
+\end{equation}
+
+$\textbf{r} \times \ddot{ \textbf{r}}=0  \implies $ $\textbf{r}$  and $\dot{\textbf{r}}$ lies in the same plane
+
+\end{frame}
+
+\begin{frame}{Celestial mechanics}
+With polar coordinates  $\hat{\textbf{r}}$ and $\hat{\boldsymbol{\theta}}$
+
+\begin{equation}
+\textbf{r}=r\hat{\textbf{r}}
+\end{equation}
+
+\begin{equation}
+\dot{\textbf{r}}=\dot{r}\hat{\textbf{r}}+r\dot{\theta}\hat{\boldsymbol{\theta}}
+\label{eq_dyn_nop}
+\end{equation}
+
+\begin{equation}
+\ddot{\textbf{r}}=\left(\ddot{r}-r\dot{\theta}^{2}\right)\hat{\textbf{r}}+\left[\dfrac{1}{r}\frac{d}{dt}\left(r^{2}\dot{\theta}\right)\right]\hat{\boldsymbol{\theta}}
+\end{equation}
+
+\begin{equation}
+\textbf{h}=r^{2}\dot{\theta}\hat{\textbf{z}}
+\end{equation}
+
+\begin{equation}
+h=r^{2}\dot{\theta}
+\end{equation}
+
+\end{frame}
+
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: $2^{th}$ Kepler law}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.3\textwidth]{Pic/Area_dynamics.png}
+\caption{\cite{murray1999solar}}
+\label{Area_dynamics}
+\end{center}
+\end{figure}
+
+\begin{equation}
+\delta A \approx \dfrac{1}{2} r(r+dr)\sin(\delta\theta) \approx  \dfrac{1}{2} r^{2}\delta\theta
+\end{equation}
+
+\begin{equation}
+\dfrac{dA}{dt}=\dfrac{1}{2}r^{2}\dfrac{d\theta}{dt}=\dfrac{1}{2}h
+\end{equation}
+
+\begin{center}
+$h$ is constant $\implies$ $2^{th}$ Kepler law
+\end{center}
+
+\end{frame}
+
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: $1^{th}$ Kepler law}
+\begin{center}
+Using the substitution $u=\dfrac{1}{r}$ $h=r^{2}\dot{\theta}$
+\end{center}
+
+\begin{equation}
+\dot{r}=-\frac{1}{u}\dfrac{du}{d\theta}\dot{\theta}=-h\frac{du}{d\theta}
+\end{equation}
+
+\begin{equation}
+\ddot{r}=-h\dfrac{d^{2}u}{d\theta^{2}}\dot{\theta}=-h^{2}u^{2}\frac{d^{2}u}{d\theta^{2}}
+\end{equation}
+
+\begin{equation}
+\dfrac{d^{2}u}{d\theta^{2}}+u=\frac{\mu}{h^{2}}
+\end{equation}
+
+\begin{equation}
+u=\frac{\mu}{h^{2}}\left[1+e\cos(\theta-\phi)\right]
+\end{equation}
+
+\end{frame}
+
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: $1^{th}$ Kepler law}
+\begin{columns}
+\column{0.5\textwidth}
+\begin{equation}
+r=\dfrac{p}{1+e\cos(\theta-\phi)}
+\end{equation}
+
+\begin{center}
+$e$ is \textcolor{red}{eccentricity}
+\end{center}
+
+\begin{itemize}
+\item circle:  $e=0$ \quad $p=a$
+\item ellipse: $0<e<1$ \quad $p=b$
+\item parabola: $e=1$ \quad $p=2q$
+\item hyperbola: $e>1$ \quad $p=a(e^{2}-1)$
+\end{itemize}
+\begin{center}
+$b$ is the  \textcolor{red}{semi-major axis} of the conic
+\end{center}
+\column{0.4\textwidth}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{Pic/Conics.png}
+\caption{\cite{murray1999solar}}
+\label{Area_dynamics}
+\end{center}
+\end{figure}
+
+\end{columns}
+\end{frame}
+
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: $3^{th}$ Kepler law}
+\begin{equation}
+b^{2}=a^{2}(1-e^{2})
+\end{equation}
+
+\begin{equation}
+r=\frac{a(1-e^{2})}{1+e\cdot cos(\theta-\phi)}
+\label{eq-mot}
+\end{equation}
+
+\begin{center}
+Area swept in one \textcolor{red}{orbital period} T $ \implies A=\pi ab$
+We know that: $hT/2 \quad h^{2}=\mu a(1-e^{2})$ 
+\end{center}
+
+Therefore
+
+\begin{equation}
+T^{2}=\dfrac{4\pi^{2}}{\mu}a^{3}
+\end{equation}
+
+\end{frame}
+
+\begin{frame}{$3^{th}$ Kepler law}
+
+\begin{equation}
+\frac{m_{c}+m}{m_{c}+m'}=\left(\frac{a}{a'}\right)\left(\frac{T'}{T}\right)^{2}
+\end{equation}
+
+But since $m$,$m'<<m_{c}$
+
+\begin{equation}
+(a\a')^{3}\approx (T\T')^{2}
+\end{equation}
+
+And therefore
+
+\begin{equation}
+T'\approx a'^{3/2}
+\end{equation}
+
+\end{frame}
+
+\begin{frame}{Orbital parameters}
+\begin{center}
+\textcolor{red}{Mean motion}
+\end{center}
+\begin{equation}
+n=\frac{2\pi}{T}
+\end{equation}
+
+\begin{equation}
+v_{perihelion}=na\sqrt{\dfrac{1+e}{1-e}}
+\end{equation}
+
+\begin{equation}
+v_{aphelion}=na\sqrt{\dfrac{1-e}{1+e}}
+\end{equation}
+
+\end{frame}
+
+\begin{frame}{Orbital parameters}
+\begin{center}
+\textcolor{red}{Mean anomaly}
+\end{center}
+\begin{equation}
+M=n(t-\tau)
+\end{equation}
+
+\begin{center}
+\begin{itemize}
+\item $M=f=0$\quad$t=\tau$\quad Perihelion
+\item $M=f=\pi$\quad$t=\tau+T/2$ \quad Aphelion
+\end{itemize}
+\end{center}
+
+\begin{equation}
+M=E-e\sin E
+\end{equation}
+
+\begin{center}
+\textcolor{red}{Jupiter Tisserard invariant }
+\end{center}
+
+\begin{equation}
+T_{P}=\frac{a_{p}}{a}+2\cos I\sqrt{\dfrac{a}{a_{p}}(1-e^{2})}
+\end{equation}
+
+\end{frame}
+
+\begin{frame}[t,allowframebreaks]
+\frametitle{References}
+\printbibliography
+\end{frame}
+
 \end{document}

### 2021-12-08 Presentation/Presentation.tex
PUB

@@
 \item circle:  $e=0$ \quad $p=a$
-\item ellipse: $0<e<1$ \quad $p=b$
+\item ellipse: $0<e<1$ \quad $p=a(1-e^{2})$
 \item parabola: $e=1$ \quad $p=2q$
@@
 \begin{center}
-$b$ is the  \textcolor{red}{semi-major axis} of the conic
+$a$ is the  \textcolor{red}{semi-major axis} of the conic
 \end{center}
@@
 \begin{frame}{Celestial mechanics \cite{murray1999solar}: $3^{th}$ Kepler law}
+\begin{columns}
+\column{0.4\textwidth}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth ]{Pic/Elliptical_orbit.png}
+\caption{\cite{murray1999solar}}
+\label{Area_dynamics}
+\end{center}
+\end{figure}
+
+\column{0.6\textwidth}
 \begin{equation}
@@
 \begin{center}
-Area swept in one \textcolor{red}{orbital period} T $ \implies A=\pi ab$
+Area swept in one \textcolor{red}{orbital period} T 
+\end{center} 
+\begin{center}
+$A=\pi ab$
+\end{center} 
+\begin{center}
 We know that: $hT/2 \quad h^{2}=\mu a(1-e^{2})$ 
 \end{center}
-
+\begin{center}
 Therefore 
-
+\end{center}
 \begin{equation}
@@
 
+\end{columns}
 \end{frame}
@@
 \end{equation}
+
+\begin{center}
+The mass of the asteroid is not involve
+\end{center}d
 
@@
 \begin{frame}{Orbital parameters}
+\begin{columns}
+\column{0.4\textwidth}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth ]{Pic/Mean_anomaly.png}
+\caption{\cite{murray1999solar}}
+\label{Area_dynamics}
+\end{center}
+\end{figure}
+\column{0.6\textwidth}
 \begin{center}
@@
 \end{equation}
-
 \begin{center}
@@
 \end{center}
-
 \begin{equation}
@@
 \end{equation}
+\end{columns}
+\end{frame}
 
+\begin{frame}{Orbital parameters}
+\begin{columns}
+\column{0.4\textwidth}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth ]{Pic/Inclination.png}
+\caption{\cite{murray1999solar}}
+\label{Area_dynamics}
+\end{center}
+\end{figure}
+\column{0.6\textwidth}
+\begin{center}
+I: \textcolor{red}{inclination} of the orbit
+\end{center}
+\begin{center}
+$\Omega$: \textcolor{red}{longitude of the ascending node}
+\end{center}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Magnitude}
+\begin{equation}
+\Phi=\frac{L}{4\pi r^{2}}
+\end{equation}
+\begin{equation}
+m=-2.5\log_{10}\Phi+C
+\end{equation}
+\begin{equation}
+m_{1}-m_{2}=-2.5\log_{10}\frac{\Phi_{1}}{\Phi_{2}}
+\end{equation}
+\begin{equation}
+M-m=-2.5\log_{10}\frac{\Phi\cdot d^{2} }{\Phi\cdot 10^{2}}
+\end{equation}
+\begin{equation}
+M=m+5-5log_{10}d
+\end{equation}
+Where $\Phi$ is the flux for a sphere of radius r, m the relative magnitude and M the \textcolor{red}{Absolute magnitude}
+\end{frame}
+
+\begin{frame}{Magnitude}
+\begin{equation}
+\Phi=\frac{L}{4\pi r^{2}}
+\end{equation}
+\begin{equation}
+m=-2.5\log_{10}\Phi+C
+\end{equation}
+\begin{equation}
+m_{1}-m_{2}=-2.5\log_{10}\frac{\Phi_{1}}{\Phi_{2}}
+\end{equation}
+\begin{equation}
+M-m=-2.5\log_{10}\frac{\Phi\cdot d^{2} }{\Phi\cdot 10^{2}}
+\end{equation}
+\begin{equation}
+M=m+5-5log_{10}d
+\end{equation}
+Where $\Phi$ is the flux for a sphere of radius r, m the relative magnitude and M the \textcolor{red}{Absolute magnitude}
+\end{frame}
+
+\begin{frame}{Classification}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{Pic/neo_orbit_types.jpg}
+\caption{\cite{nasa_classification}}
+\label{Area_dynamics}
+\end{center}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Classification}
+\begin{itemize}
+\item \textbf{Potentially Hazardous Asteroids}: MOID $\leq 0.05$ au $M \leq22.0$ \textit{NEAs whose Minimum Orbit Intersection Distance (MOID) with the Earth is 0.05 au or less and whose absolute magnitude (M) is 22.0 or brighter}
+\end{itemize}
+\end{frame}
+
+\section{Dataset}
+
+\begin{frame}{Classification}
+\begin{itemize}
+
+\item The asteroid dataset was retrieved from Kaggle \cite{kaggle_dataset}, which reports into a more machine readable form the dataset of The Center for Near-Earth Object Studies (CNEOS) \cite{cneos+nasa}, a NASA research centre.
+
+\item 3552 Asteroids
+
+\item Among the 40 the features, the ones connected only to the other name of
+the asteroid, or connected only to the name of the orbit and
+the one connected with the orbiting planet ( since for all it was
+the Earth) were discarted
+
+\item The proportion hazardous/not hazardous was set 1:5
+
+\item The continuous measures were standardised and demeaned
+
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Features}
+\begin{table}[]
+\begin{center}
+\begin{tabular}{c|c}
+\hline
+\textbf{Features}             & \textbf{Type}        \\ \hline
+Neo Reference ID              & not used             \\ \hline
+Absolute Magnitude            & Continuous           \\ \hline
+Est Dia in KM (min)           & Continuous           \\ \hline
+Est Dia in KM (max)           & Continuous           \\ \hline
+Close Approach Date           & Continuous           \\ \hline
+Epoch Date Close Approach     & Continuous           \\ \hline
+Relative\_Velocity            & Continuous           \\ \hline
+Miss\_Dist                    & Continuous           \\ \hline
+Min\_Orbit\_Intersection      & Continuous           \\ \hline
+Jupiter\_Tisserand\_Invariant & Continuous           \\ \hline
+Epoch\_Osculation             & Continuous           \\ \hline
+Eccentricity                  & Continuous           \\ \hline
+
+\end{tabular}
+\end{center}
+\label{tab_features}
+\end{table}
+
+\end{frame}
+
+\begin{frame}{Features}
+\begin{table}[]
+\begin{center}
+\begin{tabular}{c|c}
+\hline
+\textbf{Features}             & \textbf{Type}        \\ \hline
+Semi Major Axis               & Continuous           \\ \hline
+Inclination                   & Continuous           \\ \hline
+Asc Node Longitude            & Continuous           \\ \hline
+Orbital Period                & Continuous           \\ \hline
+Perihelion Distance           & Continuous           \\ \hline
+Perihelion Arg                & Continuous           \\ \hline
+Perihelion Time               & Continuous           \\ \hline
+Mean\_Anomaly                 & Continuous           \\ \hline
+Mean\_Motion                  & Continuous           \\ \hline
+Hazardous                     & Categorical (Binary)
+\end{tabular}
+\end{center}
+\label{tab_features}
+\end{table}
+
+\end{frame}
+
+\section{Preliminary analysis}
+
+\begin{frame}
+\begin{columns}
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Pic/DENSITY_Perihelion_Distance.pdf}
+    \subcaption{ \begin{center}
+    a) Perihelion Distance
+    \end{center}}
+    \vspace{4ex}
+  \end{figure} 
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Pic/DENSITY_Eccentricity.pdf}
+    \subcaption{ \begin{center}
+    b) Eccentricity
+    \end{center}}
+    \vspace{4ex}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}
+\begin{columns}
+
+\column{0.5\textwidth}
+    \begin{figure}[b]{\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Pic/DENSITY_Min_orbit_intersection.pdf}
+    \subcaption{ \begin{center}
+    c) Min orbit intersection
+    \end{center}}
+    \vspace{4ex}
+  \end{figure}
+  \column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Pic/DENSITY_Absolute_magnitude.pdf}
+    \subcaption{ \begin{center}
+    d) Absolute magnitude
+    \end{center}}
+    \vspace{4ex}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{FAMD}
+\begin{columns}
+
+\column{0.5\textwidth}
+    \begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Pic/FAMD.pdf}
+\label{FADM}
+\end{center}
+\end{figure}
+  \column{0.5\textwidth}
+ \begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{Pic/FAMD_Quantitative variables.pdf}
+\label{FAMD_Quantitative variables}
+\end{center}
+\end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Mutual information}
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.9\textwidth]{Pic/Mutual_information.pdf}
+\label{Mutual_information}
+\end{center}
+\end{figure}
+\end{frame}
+
+\section{Probabilistic models}
+
+\begin{frame}{GLASSO}
+\begin{columns}
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Pic/GLASSO_0.1.pdf}
+    \subcaption{$\rho$=0.2 \begin{center}
+    \end{center}}
+  \end{figure} 
+  \column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Pic/GLASSO_0.2.pdf}
+    \subcaption{$\rho$=0.2 \begin{center}
+    \end{center}}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{GLASSO}
+\begin{columns}
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Pic/GLASSO_0.3.pdf}
+    \subcaption{$\rho$=0.3 \begin{center}
+    \end{center}}
+  \end{figure} 
+  \column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \centering
+    \includegraphics[width=\textwidth]{Pic/GLASSO_0.4.pdf}
+    \subcaption{$\rho$=0.4 \begin{center}
+    \end{center}}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Mixed interactions: mgm}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{Pic/mgm.pdf}
+\label{mgm}
+\end{center}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Mixed interactions: minforest}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{Pic/minforest.pdf}
+\label{mgm}
+\end{center}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Mixed interactions: mmod}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{Pic/mmod.pdf}
+\label{mgm}
+\end{center}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Mixed interactions}
+The mgm model is the one that has the list of connection more coherent with the celestial mechanics laws. 
+\begin{columns}
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/mgm_confusion.pdf}
+  \end{figure} 
+  \column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/ROC_mgm.pdf}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\section{Other ML algorithms}
+\begin{frame}{Random Forest}
+\begin{columns}
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/RF_confusion.pdf}
+  \end{figure} 
+  \column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/ROC_RF.pdf}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Random Forest}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=0.7\textwidth]{Pic/RF_Importance.pdf}
+  \end{figure} 
+\end{frame}
+
+\begin{frame}{Support Vector Machines}
+\begin{columns}
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/SVM_confusion.pdf}
+  \end{figure} 
+  \column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/ROC_SVM.pdf}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Quadratic Discriminant Analysis (QDA)}
+\begin{columns}
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/QDA_confusion.pdf}
+  \end{figure} 
+  \column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/ROC_QDA.pdf}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Logistic regression}
+\begin{columns}
+\column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/Logisic_confusion.pdf}
+  \end{figure} 
+  \column{0.5\textwidth}
+  \begin{figure}[b]{\textwidth}
+    \includegraphics[width=\textwidth]{Pic/ROC_Logistic.pdf}
+  \end{figure}
+\end{columns}
+\end{frame}
+
+\begin{frame}{$\phi$ coefficient}
+\begin{table}[]
+\caption{$\phi$ coefficient (also known as Matthews correlation coefficient )}
+\begin{center}
+\begin{tabular}{c|c}
+Algorithm & $\phi$ \\ \hline
+RF        & 0.9876 \\ \hline
+SVM       & 0.7111 \\ \hline
+logistic  & 0.6173 \\ \hline
+mgm       & 0.5997 \\ \hline
+QDA       & 0.5562 
+\end{tabular}
+\end{center}
+\label{phi_values}
+\end{table}
+\end{frame}
+
+\section{Conclusions}
+
+\begin{frame}{forecast performances vs intepretability}
+\begin{itemize}
+\item The mgm algorithm is not the best one in term of performances, but it provides the connections between the features.
+On the other side, except for the variable importance in RF, the other are black box one
+\item The mgm model, as the other graphical model is open to a true scientific validation, the other not. 
+\item The probabilistic models lack in the forecast is definitely compensated by their interetability 
+\item This is meaningful since this two features are in conflict 
+\item The probabilistic models provide a good trade-off between intepretability and forecast performances, as long as one is interest to produce a really scientific result (e.g if the only aim is the forecast the RF is definitely better.
+However how long one can trust to the RF result ?)
+\end{itemize}
 \end{frame}

### 2021-12-08 Presentation/Presentation.tex
PUB

@@
 
-\section{Introduction}
+\section{Intro}
 
@@
 
-\section{Preliminary analysis}
+\section{Prelim. analysis}
 
@@
 
-\section{Probabilistic models}
+\section{Prob. models}
 
@@
 
-\section{Other ML algorithms}
+\section{Other ML alg.}
 \begin{frame}{Random Forest}
@@
 
-\begin{frame}{forecast performances vs intepretability}
+\begin{frame}{Conclusions: forecast performances vs intepretability}
 \begin{itemize}
@@
 
-\begin{frame}[t,allowframebreaks]
-\frametitle{References}
+\section{SI}
+
+\begin{frame}
+\begin{center}
+\frametitle{$\phi$ coefficent}
+\begin{table}[]
+\begin{tabular}{c|c|c}
+\multicolumn{1}{l|}{\textbf{}} & \textbf{Actual - N} & \textbf{Actual - P} \\ \hline
+\textbf{Predicted - N}         & \#TP                & \#FN                \\ \hline
+\textbf{Predicted - P}         & \#FP                & \#TP               
+\end{tabular}
+\end{table}
+\begin{equation*}
+\phi=\dfrac{TP\times TN - FP\times FN}{\sqrt{\left(TP+FP\right)\left(TP+FN\right)\left(TN+FP\right)\left(TN+FN\right)}}
+\end{equation*}
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Receiver operating characteristic}
+\begin{columns}
+\column{0.5\textwidth}
+\begin{center}
+\includegraphics[width=1\linewidth]{Pic/ROC/ROC_A.png}
+\includegraphics[width=1\linewidth]{Pic/ROC/ROC_B.png}\\
+\end{center}
+\column{0.5\textwidth}
+\begin{center}
+\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_A_R.png}
+\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_B_R.png}\\
+\end{center}
+\end{columns}
+\begin{center}
+Images taken from \cite{ROC_fig}
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Receiver operating characteristic}
+\begin{columns}
+\column{0.5\textwidth}
+\begin{center}
+\includegraphics[width=1\linewidth]{Pic/ROC/ROC_C.png}
+\includegraphics[width=1\linewidth]{Pic/ROC/ROC_D.png}\\
+\end{center}
+\column{0.5\textwidth}
+\begin{center}
+\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_C_R.png}
+\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_D_R.png}\\
+\end{center}
+\end{columns}
+\begin{center}
+Images taken from \cite{ROC_fig}
+\end{center}
+\end{frame}
+
+\begin{frame}{Factor analysis of mixed data - FAMD}
+\begin{center}
+r(z,k) correlation coefficient (z and k quantitative) \\
+\end{center}
+\begin{center}
+$\eta^{2}$(z,q) correlation ratio (z quantitative and q qualitative)
+\end{center}
+\begin{center}
+PCA $\rightarrow$ $max\sum_{k}r^{2}(z,k)$
+\end{center}
+\begin{center}
+MCA $\rightarrow$ $max\sum_{q}\eta^{2}(z,q)$
+\end{center}
+\begin{center}
+FAMD $\rightarrow$ $max\sum_{k}r^{2}(z,k)+max\sum_{q}\eta^{2}(z,q)$
+\end{center}
+\end{frame}
+
+\begin{frame}{mgm algorithm}
+\begin{equation}
+P(X_{s}|X_{\backslash s})=\exp\left\lbrace E_{s}(X_{\backslash s})\phi_{s}\left(X_{s}\right)+B_{s}(X_{s})-\Phi\left(X_{\backslash s}\right) \right\rbrace
+\end{equation}
+\begin{center}
+$\phi_{s}$ function of sufficient statistics $B_{s}$ base measure
+\end{center}
+
+\begin{equation}
+\begin{split}
+P(X)= & \exp \left\lbrace \sum_{s\in V}\theta_{s}\phi_{s}(X_{s})+\sum_{s\in V}\sum_{r\in N(s)}\theta_{s,r}\phi_{s}(X_{s})\phi_{r}(X_{r})+...+\\&\sum_{r_{1},...,r_{k}\in C}\theta_{r1,...,rk}\prod_{j=1}^{k}\phi_{r}_{j}\left(X_{r}_{j}\right)+\sum_{s\in V}B_{s}(X_{S})-\Phi(\theta) \right\rbrace
+\end{split}
+\end{equation}
+
+\begin{equation}
+\hat{\theta}=arg \min_{theta}\left\lbrace -\mathcal{L}(\theta,X)+\lambda ||\theta ||_{1}  \right\rbrace\quad ||\theta ||_{1}=\sum_{j=1}^{J}|\theta_{j}|
+\end{equation}
+
+\end{frame}
+
+\begin{frame}{GLASSO}
+\begin{equation}
+L_{pen}(K,\hat{\mu})=log\det(K)-tr(K\;S)-\rho||K|| 
+\end{equation}
+\begin{center}
+$K=\Sigma^{-1}$
+\end{center}
+\begin{center}
+S: empirical covariance matrix
+\end{center}
+\end{frame}
+
+\begin{frame}{Parameters tuning}
+\begin{equation}
+p\left(\textbf{y}|\boldsymbol{\theta}\right) = \dfrac{1}{Z(\theta)}\exp \left( \sum_{c}\boldsymbol{\theta}^{T}_{c}\phi_{c}\left(\textbf{y}\right)\right)
+\end{equation}
+\begin{equation}
+\mathcal{L}\left(\boldsymbol{\theta}\right):= \frac{1}{N}\sum_{i}\log p\left(\textbf{y}_{i}|\boldsymbol{\theta}\right)=\frac{1}{N}\sum_{i}\left[\sum_{c} \boldsymbol{\theta}^{T}_{c}\phi_{c}(y_{i})-\log Z\left(\boldsymbol{\theta}\right)\right]
+\end{equation}
+\end{frame}
+
+\begin{frame}{Parameters tuning}
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\frac{1}{N}\sum_{i}\left[\phi_{c}(y_{i})-\frac{\partial}{\partial\boldsymbol{\theta}_{c}}\log Z(\boldsymbol{\theta})\right]
+\end{equation}
+
+\begin{equation}
+\begin{split}
+\frac{\partial \log Z(\boldsymbol{\theta})}{\partial\boldsymbol{\theta}}=\mathbb{E}\left[\phi_{c}(\textbf{y})\right|\theta]= \sum_{\textbf{y}}\phi_{c}(\textbf{y})p(\textbf{y}|\boldsymbol{\theta})
+\end{split}
+\end{equation}
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\left[\frac{1}{N}\sum_{i}\phi_{c}(y_{i})\right]-\mathbb{E}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]-\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+\begin{equation}
+\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]=\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+\end{frame}
+
+\begin{frame}{mmod}
+\begin{equation}
+\begin{split}
+f(i,y)=& p(i)(2\pi)^{-q/2}det(\Sigma)^{-1/2} \\
+& exp\left[-\dfrac{1}{2}\left(y-\mu(i)\right)^{T}\Sigma^{-1}\left(y-\mu(i)\right)\right]
+\end{split}
+\label{gaussMix}
+\end{equation}
+\begin{equation}
+\begin{split}
+f(i,y) & = \exp\left\lbrace g(i)+\sum_{u}h^{u}(i)y_{u}-\dfrac{1}{2}\sum_{uv} y_{u}y_{v}k_{uv}\right\rbrace \\
+&= \exp\left\lbrace g(i)+h(i)^{T}y-\dfrac{1}{2}y^{T}Ky \right\rbrace
+\end{split}
+\end{equation}
+
+\begin{center}
+where $g(i)$, $h(i)$ and $K$ are the canonical parameters
+\end{center}
+
+\end{frame}
+\begin{frame}{mmod}
+\begin{equation}
+\begin{split}
+K=&\Sigma^{-1} \\
+h(i)=&\Sigma^{-1}\mu(i) \\
+g(i)=&\log p(i) -\frac{1}{2}\log det (\Sigma) \\
+&-\dfrac{1}{2}\mu(i)^{T}\Sigma^{-1}\mu(i)-\dfrac{q}{2}\log 2\pi
+\end{split}
+\end{equation}
+\end{frame}
+
+\begin{frame}{Graphical models}
+\begin{equation}
+p(x_{1},x_{2},...,x_{n})
+\end{equation}
+\begin{equation}
+p(x_{1:V})=p(x_{1})p(x_{2}|x_{1})p(x_{3}|x_{2},x_{1})...p(x_{V}|x_{1:V-1})
+\end{equation}
+\begin{equation}
+X  \perp Y| Z \iff  p(X,Y|Z) = p(X|Z)p(Y|Z)
+\end{equation}
+\begin{equation}
+p(\textbf{x}_{1:V})=p(x_{1})\prod^{V}_{t=1}p(x_{t}|x_{t-1})
+\end{equation}
+\end{frame}
+
+\begin{frame}{Graphical models}
+\begin{theorem}[Hammersley-Clifford]
+A positive distribution p(\textbf{y})>0 satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
+\begin{equation}
+p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
+\end{equation}
+where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
+\begin{equation}
+Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
+\end{equation}
+Note that this partition function is what ensures the overall distribution sums to 1
+\end{theorem}
+\end{frame}
+\begin{frame}{Graphical models}
+\begin{equation}
+p(y|\theta)=\dfrac{1}{Z(\theta)} exp\left(-\sum_{c}E(y_{c}|\theta_{c})\right)
+\end{equation}
+\begin{equation}
+\psi_{c}(y_{c}|\theta_{c})=exp\left(-E(y_{c}|\theta_{c})\right)
+\end{equation}
+\end{frame}
+
+\begin{frame}{Information theory}
+\begin{equation}
+H(X)=-\sum_{x \in X} p(x)log p(x)
+\end{equation}
+\begin{equation}
+H(X,Y)=-\sum_{x\in \chi}\sum_{y\in \mathcal{Y}}p(x,y)\log p(x,y)
+\end{equation}
+\begin{equation}
+\begin{split}
+H(X|Y)=& \sum_{x\in \mathcal{X} }p(x)H(Y|X=x)\\
+=& -\sum_{x\in \mathcal{X}}p(x)\sum_{y\in \mathcal{Y}}p(x,y)\log p(y|x) \\
+=& -E\log p(Y|X)
+\end{split}
+\end{equation}
+\begin{equation}
+H(X,Y)=H(X)+H(Y|X)
+\end{equation}
+\begin{equation}
+D(p||q)=\sum p(x)\log\frac{p(x)}{q(x)} \quad D(p||q)\geq 0
+\end{equation}
+
+\end{frame}
+
+\begin{frame}{Information theory}
+\begin{equation}
+\begin{split}
+I(X;Y)&=\sum\sump(x,y)\log\dfrac{p(x,y)}{p(x)p(y)}=D(p(x,y)||p(x)p(y)) \\
+&=H(X)-H(X|Y)=H(Y)-H(Y|X)
+\end{split}
+\end{equation}
+\begin{equation}
+I(X_{1},X_{2},...,X_{n};Y)=\sum^{n}_{i=1}I(X_{i};Y|X_{i-1},X_{i-2},...,X_{1})
+\end{equation}
+\begin{equation}
+g(\alpha,\textbf{C},\textbf{S},f_{i})=MI(f_{i};\textbf{C})-\sum_{f_{s}\in S}\alpha(f_{i},f_{s},\textbf{C},\textbf{S})MI(f_{i};f_{s})
+\end{equation}
+\end{frame}
+
+\begin{frame}{Ref}
 \printbibliography

### 2021-12-09 Presentation/Presentation.tex
PUB

@@
 \addbibresource{bibliography.bib}
-
+\newtheorem*{remark}{Remark}
 \usepackage[english]{babel}
@@
 \item \textbf{Final Goal} Assessment of forecasts and interpretability for different machine learning algorithms, including the probabilistic models
-\item \textbf{Method} Use a dataset for which the laws that interconnects the different features are known from general principles
+\item \textbf{Method} Use a dataset for which the laws that interconnect the different features are known from general principles
 \item \textbf{Dataset} CNEOS asteroids dataset for more than 3500 asteroids
@@
 
-\begin{frame}{Celestial mechanics}
-\begin{equation}
-\textbf{F}_{1}=\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{1}
-\end{equation}
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: equations of motion}
+
+\begin{center}
+Lets Consider the interaction between a planet of mass $m_{1}$ at the  position $r_{1}$ (inertial frame) and an asteroid of mass $m_{2}$  at the position $r_{2}$
+\end{center}
 
 \begin{equation}
-\textbf{F}_{2}=-\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{2}
+\textbf{F}_{1}=\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{1}\quad \textbf{F}_{2}=-\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{2}
 \end{equation}
@@
 
+\begin{center}
 $\textbf{r} \times \ddot{ \textbf{r}}=0  \implies $ $\textbf{r}$  and $\dot{\textbf{r}}$ lies in the same plane
 
+\end{center}
 \end{frame}
 
-\begin{frame}{Celestial mechanics}
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: equations of motion}
+\begin{center}
 With polar coordinates  $\hat{\textbf{r}}$ and $\hat{\boldsymbol{\theta}}$
-
+\end{center}
 \begin{equation}
@@
 
-\begin{frame}{$3^{th}$ Kepler law}
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: $3^{th}$ Kepler law}
 
 \begin{equation}
-\frac{m_{c}+m}{m_{c}+m'}=\left(\frac{a}{a'}\right)\left(\frac{T'}{T}\right)^{2}
+\frac{m_{c}+m}{m_{c}+m'}=\left(\frac{a}{a'}\right)^{3}\left(\frac{T'}{T}\right)^{2}
 \end{equation}
@@
 \begin{equation}
-(a\a')^{3}\approx (T\T')^{2}
+\left(\frac{a}{a'}\right)^{3}\approx \left(\frac{T}{T'}'\right)^{2}
 \end{equation}
@@
 \begin{center}
-The mass of the asteroid is not involve
-\end{center}d
+\textbf{Remark:} The mass of the asteroid is \textbf{not} involved
+\end{center}
 
@@
 
-\begin{frame}{Orbital parameters}
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: Orbital parameters}
 \begin{center}
-\textcolor{red}{Mean motion}
+\textcolor{red}{Mean motion} $n=\frac{2\pi}{T}$
 \end{center}
-\begin{equation}
-n=\frac{2\pi}{T}
-\end{equation}
 
@@
 
+\begin{center}
+\textbf{Remark:} The mean motion of an asteroid is different with respect to the the asteroid relative velocity (measured from Earth), since the latter is different at the perihelion an at the aphelion
+\end{center}
+
 \end{frame}
 
-\begin{frame}{Orbital parameters}
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: Orbital parameters}
 \begin{columns}
@@
 
-\begin{frame}{Orbital parameters}
-\begin{columns}
-\column{0.4\textwidth}
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: Orbital parameters}
+
 \begin{figure}[h]
 \begin{center}
-\includegraphics[width=\textwidth ]{Pic/Inclination.png}
+\includegraphics[width=0.6\textwidth ]{Pic/Inclination.png}
 \caption{\cite{murray1999solar}}
@@
 \end{figure}
-\column{0.6\textwidth}
+
 \begin{center}
@@
 \end{center}
-\end{columns}
 \end{frame}
 
-\begin{frame}{Magnitude}
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: Magnitude}
 \begin{equation}
@@
 
-\begin{frame}{Magnitude}
+\begin{frame}{Celestial mechanics \cite{murray1999solar}: Magnitude}
 \begin{equation}
@@
 
-\begin{frame}{Classification}
+\begin{frame}{Celestial mechanics \cite{nasa_classification}: Classification}
 \begin{figure}[h]
@@
 
-\begin{frame}{Classification}
+\begin{frame}{Celestial mechanics \cite{nasa_classification}: Classification}
 \begin{itemize}
@@
 
-\begin{frame}{Classification}
+\begin{frame}{Dataset}
 \begin{itemize}
@@
 
-\begin{frame}
+\begin{frame}{Density Plot}
 \begin{columns}
@@
 
-\begin{frame}{FAMD}
+\begin{frame}{FAMD }
 \begin{columns}
@@
 \end{columns}
+\begin{center}
+Performed with the FactoMineR package \cite{le2008factominer} 
+\end{center}
 \end{frame}
 
-\begin{frame}{Mutual information}
+\begin{frame}{Mutual information analysis}
 \begin{figure}
 \begin{center}
-\includegraphics[width=0.9\textwidth]{Pic/Mutual_information.pdf}
+\includegraphics[width=0.8\textwidth]{Pic/Mutual_information.pdf}
 \label{Mutual_information}
@@
 \end{figure}
+\begin{center}
+Performed with the varrank package \cite{kratzer2018varrank} 
+\end{center}
 \end{frame}
 
-\section{Prob. models}
+\section{Prob. mod.}
 
@@
 \end{columns}
+\begin{center}
+Performed with the GLASSO package \cite{glasso}
+\end{center}
 \end{frame}
@@
 \end{figure}
+\begin{center}
+Performed with the mgm package \cite{haslbeck2015mgm}
+\end{center}
 \end{frame}
@@
 \end{figure}
+\begin{center}
+Performed with the gRapHD package \cite{de2009high}
+\end{center}
 \end{frame}
@@
 \end{figure}
+\begin{center}
+Performed with the gRim package \cite{grim}
+\end{center}
 \end{frame}
@@
 \end{columns}
+\begin{center}
+Performed with the rfor package \cite{rfor}
+\end{center}
 \end{frame}
@@
 \end{columns}
+\begin{center}
+Performed with the e1071 package \cite{dimitriadou2008misc}
+\end{center}
 \end{frame}
@@
 \end{columns}
+\begin{center}
+Performed with the MASS package \cite{MASS}
+\end{center}
 \end{frame}
@@
 \end{columns}
+\begin{center}
+Performed with the stats package \cite{stats}
+\end{center}
 \end{frame}
@@
 \section{Conclusions}
+
+\begin{frame}{Interpretability and scientific validation}
+
+\begin{remark}[Interpretability - Tarski definition]
+The formal theory T can be translated into S if and only if S can prove the theorem of T in its language \cite{tarski1953undecidable}
+\end{remark}
+
+\end{frame}
+
+\begin{frame}{Interpretability and scientific validation}
+
+\begin{remark}[Scientific method - Einstein definition]
+Science uses the totality of the primary concepts, i.e., concepts directly connected with sense experiences, and propositions connecting them.
+Such a state of affairs cannot, however, satisfy a spirit which is really scientifically minded; because the totality of concepts and relations obtained in this manner is utterly lacking in logical unity.
+In order to supplement this deficiency, one invents a system poorer in concepts and relations, a system retaining the primary concepts and relations of the \textit{first layer} as logically derived concepts and relations.
+This new \textit{secondary system} pays for its higher logical unity by having elementary concepts (concepts of the second layer), which are no longer directly connected with complexes of sense experiences \cite{physics-reality}
+\end{remark}
+
+\end{frame}
+
+\begin{frame}{Interpretability and scientific validation}
+
+\begin{remark}[Scientific method - Einstein definition ]
+The essential thing is the aim to represent the multitude of concepts and propositions, close to experience, as propositions, logically deduced from a basis, as narrow as possible, of fundamental concepts and fundamental relations which themselves can be chosen freely (axioms) \cite{physics-reality}
+\end{remark}
+
+\end{frame}
 
@@
 \begin{split}
-P(X)= & \exp \left\lbrace \sum_{s\in V}\theta_{s}\phi_{s}(X_{s})+\sum_{s\in V}\sum_{r\in N(s)}\theta_{s,r}\phi_{s}(X_{s})\phi_{r}(X_{r})+...+\\&\sum_{r_{1},...,r_{k}\in C}\theta_{r1,...,rk}\prod_{j=1}^{k}\phi_{r}_{j}\left(X_{r}_{j}\right)+\sum_{s\in V}B_{s}(X_{S})-\Phi(\theta) \right\rbrace
+P(X)= & \exp \left( \sum_{s\in V}\theta_{s}\phi_{s}(X_{s})+\sum_{s\in V}\sum_{r\in N(s)}\theta_{s,r}\phi_{s}(X_{s})\phi_{r}(X_{r})\\ +&...+\sum_{r_{1},...,r_{k}\in C}\theta_{r1,...,rk}\prod_{j=1}^{k}\phi_{r}_{j}\left(X_{r}_{j}\right)+\sum_{s\in V}B_{s}(X_{S})-\Phi(\theta) \right)
 \end{split}
@@
 
-\begin{frame}{Ref}
+\begin{frame}[t,allowframebreaks]
+\frametitle{Bibliography}
 \printbibliography

### 2021-12-11 Presentation/Presentation.tex
PUB

@@
 {
-  \usetheme{Darmstadt}      % or try Darmstadt, Madrid, Warsaw, ...
+  \usetheme{Madrid}      % or try Darmstadt, Madrid, Warsaw, ...
   \usecolortheme{beaver} % or try albatross, beaver, crane, ...
@@
 
+\usepackage[bookmarks]{hyperref}
 \usepackage[backend=bibtex]{biblatex}
@@
 \usebackgroundtemplate{\includegraphics[width=\paperwidth]{Pic/Intro.jpg}}
-\title[Hazardous asteroids forecast via Markov random fields]{Hazardous asteroids forecast via Markov random fields}
+\title[Probabilistic modeling project]{Hazardous asteroids forecast via Markov random fields}
 \subtitle{Project for the exam: Probabilistic Modelling (DSE)}
@@
 \begin{frame}
-\vspace{+4 cm}  \titlepage
+\vspace{+4.2 cm}  \titlepage
 \end{frame}
@@
 % Uncomment these lines for an automatically generated outline.
-%\begin{frame}{Outline}
-%  \tableofcontents
-%\end{frame}
+\begin{frame}{Outline}
+\setcounter{tocdepth}{1}
+\begin{center}
+  \tableofcontents
+\end{center}
+\end{frame}
 
@@
 
+\begin{frame}{}
+\begin{center}
+{\Huge Celestial mechanics}
+\end{center}
+\end{frame}
+
 \begin{frame}{Celestial mechanics \cite{murray1999solar}: equations of motion}
@@
 \section{Dataset}
+
+\begin{frame}{}
+\begin{center}
+{\Huge Dataset}
+\end{center}
+\end{frame}
 
@@
 
+\begin{frame}{}
+\begin{center}
+{\Huge Prelim. analysis}
+\end{center}
+\end{frame}
+
 \begin{frame}{Density Plot}
@@
 
-\section{Prob. mod.}
+\section{Probabilistic modelling}
+
+\begin{frame}{}
+\begin{center}
+{\Huge Probabilistic modelling}
+\end{center}
+\end{frame}
 
@@
     \includegraphics[width=\textwidth]{Pic/GLASSO_0.1.pdf}
-    \subcaption{$\rho$=0.2 \begin{center}
+    \subcaption{$\rho$=0.1 \begin{center}
     \end{center}}
@@
 
-\section{Other ML alg.}
+\section{Other ML algorithms}
+
+\begin{frame}{}
+\begin{center}
+{\Huge Other ML algorithms}
+\end{center}
+\end{frame}
+
 \begin{frame}{Random Forest}
@@
 
+\begin{frame}{}
+\begin{center}
+{\Huge Conclusions}
+\end{center}
+\end{frame}
+
 \begin{frame}{Interpretability and scientific validation}
@@
 \item The mgm model, as the other graphical model is open to a true scientific validation, the other not. 
-\item The probabilistic models lack in the forecast is definitely compensated by their interetability 
+\item The probabilistic models lack in the forecast is definitely compensated by their interpretability 
 \item This is meaningful since this two features are in conflict 
@@
 
-\section{SI}
+\begin{frame}
+
+{\Large \textit{In its efforts to learn as much as possible about nature, modern physics has found that certain things can never be “known” with certainty.
+Much of our knowledge must always remain uncertain.
+The most we can know is in terms of probabilities.} Richard P.
+Feynman (1918-1988)}
+\end{frame}
+
+\begin{frame}[t,allowframebreaks]
+\frametitle{Bibliography}
+\printbibliography
+\end{frame}
+
+\section{Supporting Info}
+
+\subsection{ROC and $\phi$ factor}
+\begin{frame}{}
+\begin{center}
+{\Huge ROC and $\phi$ factor}
+\end{center}
+\end{frame}
 
@@
 
+\subsection{FAMD}
+\begin{frame}{}
+\begin{center}
+{\Huge FAMD}
+\end{center}
+\end{frame}
+
 \begin{frame}{Factor analysis of mixed data - FAMD}
@@
 FAMD $\rightarrow$ $max\sum_{k}r^{2}(z,k)+max\sum_{q}\eta^{2}(z,q)$
+\end{center}
+\end{frame}
+
+\subsection{Graphical models}
+\begin{frame}{}
+\begin{center}
+{\Huge Graphical models}
+\end{center}
+\end{frame}
+
+\begin{frame}{Graphical models}
+\begin{equation}
+p(x_{1},x_{2},...,x_{n})
+\end{equation}
+\begin{equation}
+p(x_{1:V})=p(x_{1})p(x_{2}|x_{1})p(x_{3}|x_{2},x_{1})...p(x_{V}|x_{1:V-1})
+\end{equation}
+\begin{equation}
+X  \perp Y| Z \iff  p(X,Y|Z) = p(X|Z)p(Y|Z)
+\end{equation}
+\begin{equation}
+p(\textbf{x}_{1:V})=p(x_{1})\prod^{V}_{t=1}p(x_{t}|x_{t-1})
+\end{equation}
+\end{frame}
+
+\begin{frame}{Graphical models}
+\begin{theorem}[Hammersley-Clifford]
+A positive distribution p(\textbf{y})>0 satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
+\begin{equation}
+p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
+\end{equation}
+where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
+\begin{equation}
+Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
+\end{equation}
+Note that this partition function is what ensures the overall distribution sums to 1
+\end{theorem}
+\end{frame}
+
+\begin{frame}{Graphical models}
+\begin{equation}
+p(y|\theta)=\dfrac{1}{Z(\theta)} exp\left(-\sum_{c}E(y_{c}|\theta_{c})\right)
+\end{equation}
+\begin{equation}
+\psi_{c}(y_{c}|\theta_{c})=exp\left(-E(y_{c}|\theta_{c})\right)
+\end{equation}
+\end{frame}
+
+\begin{frame}{Parameters tuning}
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\frac{1}{N}\sum_{i}\left[\phi_{c}(y_{i})-\frac{\partial}{\partial\boldsymbol{\theta}_{c}}\log Z(\boldsymbol{\theta})\right]
+\end{equation}
+
+\begin{equation}
+\begin{split}
+\frac{\partial \log Z(\boldsymbol{\theta})}{\partial\boldsymbol{\theta}}=\mathbb{E}\left[\phi_{c}(\textbf{y})\right|\theta]= \sum_{\textbf{y}}\phi_{c}(\textbf{y})p(\textbf{y}|\boldsymbol{\theta})
+\end{split}
+\end{equation}
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\left[\frac{1}{N}\sum_{i}\phi_{c}(y_{i})\right]-\mathbb{E}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+\begin{equation}
+\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]-\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+\begin{equation}
+\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]=\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
+\end{equation}
+\end{frame}
+
+\begin{frame}{Parameters tuning}
+\begin{equation}
+p\left(\textbf{y}|\boldsymbol{\theta}\right) = \dfrac{1}{Z(\theta)}\exp \left( \sum_{c}\boldsymbol{\theta}^{T}_{c}\phi_{c}\left(\textbf{y}\right)\right)
+\end{equation}
+\begin{equation}
+\mathcal{L}\left(\boldsymbol{\theta}\right):= \frac{1}{N}\sum_{i}\log p\left(\textbf{y}_{i}|\boldsymbol{\theta}\right)=\frac{1}{N}\sum_{i}\left[\sum_{c} \boldsymbol{\theta}^{T}_{c}\phi_{c}(y_{i})-\log Z\left(\boldsymbol{\theta}\right)\right]
+\end{equation}
+\end{frame}
+
+\begin{frame}{GLASSO}
+\begin{equation}
+L_{pen}(K,\hat{\mu})=log\det(K)-tr(K\;S)-\rho||K|| 
+\end{equation}
+\begin{center}
+$K=\Sigma^{-1}$
+\end{center}
+\begin{center}
+S: empirical covariance matrix
 \end{center}
@@
 
-\begin{frame}{GLASSO}
-\begin{equation}
-L_{pen}(K,\hat{\mu})=log\det(K)-tr(K\;S)-\rho||K|| 
-\end{equation}
-\begin{center}
-$K=\Sigma^{-1}$
-\end{center}
-\begin{center}
-S: empirical covariance matrix
-\end{center}
-\end{frame}
-
-\begin{frame}{Parameters tuning}
-\begin{equation}
-p\left(\textbf{y}|\boldsymbol{\theta}\right) = \dfrac{1}{Z(\theta)}\exp \left( \sum_{c}\boldsymbol{\theta}^{T}_{c}\phi_{c}\left(\textbf{y}\right)\right)
-\end{equation}
-\begin{equation}
-\mathcal{L}\left(\boldsymbol{\theta}\right):= \frac{1}{N}\sum_{i}\log p\left(\textbf{y}_{i}|\boldsymbol{\theta}\right)=\frac{1}{N}\sum_{i}\left[\sum_{c} \boldsymbol{\theta}^{T}_{c}\phi_{c}(y_{i})-\log Z\left(\boldsymbol{\theta}\right)\right]
-\end{equation}
-\end{frame}
-
-\begin{frame}{Parameters tuning}
-\begin{equation}
-\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\frac{1}{N}\sum_{i}\left[\phi_{c}(y_{i})-\frac{\partial}{\partial\boldsymbol{\theta}_{c}}\log Z(\boldsymbol{\theta})\right]
-\end{equation}
-
-\begin{equation}
-\begin{split}
-\frac{\partial \log Z(\boldsymbol{\theta})}{\partial\boldsymbol{\theta}}=\mathbb{E}\left[\phi_{c}(\textbf{y})\right|\theta]= \sum_{\textbf{y}}\phi_{c}(\textbf{y})p(\textbf{y}|\boldsymbol{\theta})
-\end{split}
-\end{equation}
-\begin{equation}
-\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\left[\frac{1}{N}\sum_{i}\phi_{c}(y_{i})\right]-\mathbb{E}\left[\phi_{c}(\textbf{y})\right]
-\end{equation}
-\begin{equation}
-\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]-\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
-\end{equation}
-\begin{equation}
-\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]=\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
-\end{equation}
-\end{frame}
-
+\begin{frame}{mmod - Homogeneus Mixed Interaction (HMI) models}
+\begin{equation}
+f(\textbf{i},\textbf{y})= p(\textbf{i})(2\pi)^{-q/2}det(\Sigma)^{-1/2}\exp\left[-\dfrac{1}{2}\left(\textbf{y}-\mu(\textbf{i})\right)^{T}\Sigma^{-1}\left(\textbf{y}-\mu(\textbf{i})\right)\right]
+\label{gaussMix}
+\end{equation}
+
+\begin{equation}
+f(\textbf{i},\textbf{y})=\exp\left[g(\textbf{i}) +h(\textbf{i})^{T}\textbf{y}-\frac{1}{2}y^{T}K\textbf{y}\right]
+\end{equation}
+
+\begin{center}
+where $K$ is the concentration matrix, $g(i)$ and $h(i)$ are the log-linear expansion of the probability $p(\textbf{i})$ (canonical parameters)
+\end{center}
+
+\end{frame}
 \begin{frame}{mmod}
-\begin{equation}
-\begin{split}
-f(i,y)=& p(i)(2\pi)^{-q/2}det(\Sigma)^{-1/2} \\
-& exp\left[-\dfrac{1}{2}\left(y-\mu(i)\right)^{T}\Sigma^{-1}\left(y-\mu(i)\right)\right]
-\end{split}
-\label{gaussMix}
-\end{equation}
-\begin{equation}
-\begin{split}
-f(i,y) & = \exp\left\lbrace g(i)+\sum_{u}h^{u}(i)y_{u}-\dfrac{1}{2}\sum_{uv} y_{u}y_{v}k_{uv}\right\rbrace \\
-&= \exp\left\lbrace g(i)+h(i)^{T}y-\dfrac{1}{2}y^{T}Ky \right\rbrace
-\end{split}
-\end{equation}
-
-\begin{center}
-where $g(i)$, $h(i)$ and $K$ are the canonical parameters
-\end{center}
-
-\end{frame}
-\begin{frame}{mmod}
-\begin{equation}
-\begin{split}
-K=&\Sigma^{-1} \\
-h(i)=&\Sigma^{-1}\mu(i) \\
-g(i)=&\log p(i) -\frac{1}{2}\log det (\Sigma) \\
-&-\dfrac{1}{2}\mu(i)^{T}\Sigma^{-1}\mu(i)-\dfrac{q}{2}\log 2\pi
-\end{split}
-\end{equation}
-\end{frame}
-
-\begin{frame}{Graphical models}
-\begin{equation}
-p(x_{1},x_{2},...,x_{n})
-\end{equation}
-\begin{equation}
-p(x_{1:V})=p(x_{1})p(x_{2}|x_{1})p(x_{3}|x_{2},x_{1})...p(x_{V}|x_{1:V-1})
-\end{equation}
-\begin{equation}
-X  \perp Y| Z \iff  p(X,Y|Z) = p(X|Z)p(Y|Z)
-\end{equation}
-\begin{equation}
-p(\textbf{x}_{1:V})=p(x_{1})\prod^{V}_{t=1}p(x_{t}|x_{t-1})
-\end{equation}
-\end{frame}
-
-\begin{frame}{Graphical models}
-\begin{theorem}[Hammersley-Clifford]
-A positive distribution p(\textbf{y})>0 satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
-\begin{equation}
-p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
-\end{equation}
-where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
-\begin{equation}
-Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
-\end{equation}
-Note that this partition function is what ensures the overall distribution sums to 1
-\end{theorem}
-\end{frame}
-\begin{frame}{Graphical models}
-\begin{equation}
-p(y|\theta)=\dfrac{1}{Z(\theta)} exp\left(-\sum_{c}E(y_{c}|\theta_{c})\right)
-\end{equation}
-\begin{equation}
-\psi_{c}(y_{c}|\theta_{c})=exp\left(-E(y_{c}|\theta_{c})\right)
-\end{equation}
+\begin{center}
+Decomposable model
+\end{center}
+\begin{equation}
+\hat{f}(x)=\prod_{j=1}^{k}\dfrac{\hat{f}_{C_{j}}(x_{C}_{j})}{\hat{f}_{S_{j}}(x_{S}_{j})} 
+\end{equation}
+Likehood 
+\begin{equation}
+\hat{\mathcal{L}_{s}}=\sum_{i}n(i)log\left(\dfrac{n(i)}{N}\right)-Nq\dfrac{log(2pi)}{2}-N\dfrac{log(det(S))}{2}-N\frac{q}{2}
+\end{equation}
+
+\end{frame}
+
+\subsection{Information theory}
+\begin{frame}{}
+\begin{center}
+{\Huge Information theory}
+\end{center}
 \end{frame}
@@
 
-\begin{frame}[t,allowframebreaks]
-\frametitle{Bibliography}
-\printbibliography
-\end{frame}
-
 \end{document}

### 2021-12-12 Presentation/Presentation.tex
PUB

@@
 \begin{center}
-Lets Consider the interaction between a planet of mass $m_{1}$ at the  position $r_{1}$ (inertial frame) and an asteroid of mass $m_{2}$  at the position $r_{2}$
-\end{center}
-
-\begin{equation}
-\textbf{F}_{1}=\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{1}\quad \textbf{F}_{2}=-\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{2}
-\end{equation}
-
-If we consider the motion of the second item with respect to the first one
+The interaction between a planet of mass $m_{1}$ at the  position $r_{1}$ (inertial frame) and an asteroid of mass $m_{2}$  at the position $r_{2}$ is given by: 
+\end{center}
+
+\begin{equation}
+\textbf{F}_{1}=\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{1} \ddot{\textbf{r}}_{1}\quad \textbf{F}_{2}=-\mathcal{G} \cdot \frac{m_{1}m_{2}}{r^{3}}\textbf{r}=m_{2} \ddot{\textbf{r}}_{2}
+\end{equation}
+
+\begin{center}
+Where $\mathcal{G}$ is the universal gravitational constant. If we consider the motion of the second item with respect to the first one 
+\end{center}
 
@@
 
+\begin{center}
+Consider two asteroids of mass $m$ and $m'$ orbiting the Earth $m_{c}$ , with semi-major axes $a$ and $a'$ and orbital periods $T$ and $T'$
+\end{center}
+
 \begin{equation}
@@
 \begin{equation}
-\left(\frac{a}{a'}\right)^{3}\approx \left(\frac{T}{T'}'\right)^{2}
-\end{equation}
-
-And therefore
-
-\begin{equation}
-T'\approx a'^{3/2}
-\end{equation}
-
-\begin{center}
-\textbf{Remark:} The mass of the asteroid is \textbf{not} involved
+\left(\frac{a}{a'}\right)^{3}\approx \left(\frac{T}{T'}\right)^{2}
+\end{equation}
+
+\begin{center}
+\textbf{Remark:} The mass of the asteroids is \textbf{not} involved
 \end{center}

### 2021-12-12 Presentation/Presentation.tex
PUB

@@
 \usepackage[backend=bibtex]{biblatex}
+\usepackage{braket}
 \addbibresource{bibliography.bib}
@@
 \begin{center}
-With polar coordinates  $\hat{\textbf{r}}$ and $\hat{\boldsymbol{\theta}}$
-\end{center}
+Integrating $\textbf{r} \times \ddot{ \textbf{r}}=0$
+\end{center}
+
+\begin{equation}
+\textbf{r} \times \dot{ \textbf{r}}=\textbf{h} 
+\end{equation}
+
+\begin{center}
+Where \textbf{h}  is a constant of Integration. Using the polar coordinates  $\hat{\textbf{r}}$ and $\hat{\boldsymbol{\theta}}$
+\end{center}
+
 \begin{equation}
@@
 \begin{equation}
-\Phi=\frac{L}{4\pi r^{2}}
+\Phi=\frac{L}{4\pi d^{2}}
 \end{equation}
@@
 Science uses the totality of the primary concepts, i.e., concepts directly connected with sense experiences, and propositions connecting them.
-Such a state of affairs cannot, however, satisfy a spirit which is really scientifically minded; because the totality of concepts and relations obtained in this manner is utterly lacking in logical unity.
+Such a state of affairs cannot, however, satisfy a spirit which is really scientifically minded;\textbf{ because the totality of concepts and relations obtained in this manner is utterly lacking in logical unity}.
 In order to supplement this deficiency, one invents a system poorer in concepts and relations, a system retaining the primary concepts and relations of the \textit{first layer} as logically derived concepts and relations.
-This new \textit{secondary system} pays for its higher logical unity by having elementary concepts (concepts of the second layer), which are no longer directly connected with complexes of sense experiences \cite{physics-reality}
-\end{remark}
-
-\end{frame}
-
-\begin{frame}{Interpretability and scientific validation}
-
-\begin{remark}[Scientific method - Einstein definition ]
-The essential thing is the aim to represent the multitude of concepts and propositions, close to experience, as propositions, logically deduced from a basis, as narrow as possible, of fundamental concepts and fundamental relations which themselves can be chosen freely (axioms) \cite{physics-reality}
+\textbf{This new \textit{secondary system} pays for its higher logical unity by having elementary concepts (concepts of the second layer), which are no longer directly connected with complexes of sense experiences} \cite{physics-reality}
 \end{remark}
@@
 \begin{theorem}[Hammersley-Clifford]
-A positive distribution p(\textbf{y})>0 satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
+A positive distribution $p(\textbf{y})>0$ satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
 \begin{equation}
@@
 
+\begin{frame}{Gaussian graphical model}
+\begin{equation}
+f_{Y}(y)=\dfrac{det(K)^{1/2}}{(2\pi)^{d/2}}\exp\left(-\dfrac{1}{2}\textbf{y}^{T}K\textbf{y}\right)
+\end{equation}
+
+\begin{equation}
+K=\Sigma^{-1}
+\end{equation}
+
+\end{frame}
+
 \begin{frame}{GLASSO}
@@
 \begin{frame}{mmod - Homogeneus Mixed Interaction (HMI) models}
+\begin{center}
+N observation, d discrete variables and q continuous variable.
+The observation has the form $x=(i,y)=(i_{1},...,i_{d},y_{1}...y_{q})$.
+The probability of discrete variables falling in the cell i is denoted as $p(i)$.
+The conditional distribution of continous variables to fall in the cell $i$ is given by the multivariate gaussian $\mathcal{N}(\mu(i),\Sigma)$
+\end{center}
+
 \begin{equation}
@@
 \end{equation}
-Likehood 
-\begin{equation}
-\hat{\mathcal{L}_{s}}=\sum_{i}n(i)log\left(\dfrac{n(i)}{N}\right)-Nq\dfrac{log(2pi)}{2}-N\dfrac{log(det(S))}{2}-N\frac{q}{2}
+Maximized likehood
+\begin{equation}
+\hat{\mathcal{L}_{s}}=\sum_{i}n(i)log\left(\dfrac{n(i)}{N}\right)-Nq\dfrac{log(2\pi)}{2}-N\dfrac{log(det(S))}{2}-N\frac{q}{2}
+\end{equation}
+
+\begin{equation}
+\hat{p}(\textbf{i})=n(\textbf{i})/N \quad \hat{\mu}(\textbf{i})=\bar{y}(\textbf{i})\quad\hat{\Sigma}=S=\sum_{i}n(i)S_{i}/N
 \end{equation}

### 2021-12-12 Presentation/Presentation.tex
PUB

@@
 } 
-
+\usepackage{pifont}
 \usepackage[bookmarks]{hyperref}
@@
 \begin{frame}{Graphical models}
-\begin{equation}
-p(x_{1},x_{2},...,x_{n})
-\end{equation}
+\begin{center}
+\textbf{Problem:} How one can represent in a compact and elegant way the joint distribution $p(\textbf{x}|\boldsymbol{\theta})$ (where $\theta$ are the parameters)
+\end{center}
+\begin{center}
+First one can consider to use the chain rule
+\end{center}
 \begin{equation}
@@
 \end{equation}
+\begin{center}
+One can represent $p(x_{2},x_{1})$ as a $\mathcal{O}(K^{2})$ where the elements are $p(x_{2}=j|x_{1}=i)=T_{ij}$ and so on with high order tensors. T is called stochastic matrix
+\end{center}
+\end{frame}
+
+\begin{frame}{Graphical models}
+
+\begin{center}
+In order to reduce the computational cost $\mathcal{O}(K^{2}V^{2})$ one can consider to take in to account the concept of conditional independence
+\end{center}
 \begin{equation}
@@
 \end{equation}
+\begin{center}
+Therefore the distribution can be factorized as:
+\end{center}
 \begin{equation}
@@
 \end{equation}
-\end{frame}
-
-\begin{frame}{Graphical models}
+\begin{center}
+Which is a first order Markov chain
+\end{center}
+\end{frame}
+
+\subsection{Markov random field}
+
+\begin{frame}{}
+\begin{center}
+{\Huge Markov random field}
+\end{center}
+\end{frame}
+
+\begin{frame}{Markov random field}
+\begin{center}
+\textbf{Problem:} differently with respect to Bayesian Network for an undirected graph there is no topological ordering associated to it. Therefore the chain rule cannot be used 
+\end{center}
+\begin{center}
+\textbf{Solution:} One can use the potential function associated with maximal clique $\psi_{c}(\textbf{y}_{c}|\boldsymbol{\theta}_{c})$
+\end{center}
+
+\end{frame}
+
+\begin{frame}{Markov random field}
 \begin{theorem}[Hammersley-Clifford]
@@
 \begin{equation}
-p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
-\end{equation}
-where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
-\begin{equation}
-Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
+p(\textbf{y}|\boldsymbol{\theta})= \dfrac{1}{Z(\boldsymbol{\theta})}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\boldsymbol{\theta}_{c})
+\end{equation}
+where C is the set of all the (maximal) cliques of G,  and Z($\boldsymbol{\theta}$) is the partition function given by 
+\begin{equation}
+Z(\boldsymbol{\theta}):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\boldsymbol{\theta}_{c})
 \end{equation}
@@
 
-\begin{frame}{Graphical models}
-\begin{equation}
-p(y|\theta)=\dfrac{1}{Z(\theta)} exp\left(-\sum_{c}E(y_{c}|\theta_{c})\right)
-\end{equation}
-\begin{equation}
-\psi_{c}(y_{c}|\theta_{c})=exp\left(-E(y_{c}|\theta_{c})\right)
-\end{equation}
-\end{frame}
-
-\begin{frame}{Parameters tuning}
+\begin{frame}{Markov random field}
+\begin{center}
+Gibbs distribution 
+\end{center}
+\begin{equation}
+p(\textbf{y}|\boldsymbol{\theta})=\dfrac{1}{Z(\boldsymbol{\theta})}\exp\left(-\sum_{c}E(\textbf{y}|\boldsymbol{\theta})\right)
+\end{equation}
+\begin{center}
+\begin{center}
+Where $E(\textbf{y})$ is the energy associated with the variables in clique c. Therefore
+\end{center}
+\end{center}
+\begin{equation}
+\psi(\textbf{y}_{c}|\boldsymbol{\theta}_{c})=\exp\left(-E(\textbf{y}_{c}|\boldsymbol{\theta}_{c})\right)
+\end{equation}
+
+\end{frame}
+
+\begin{frame}{Markov random field}
 \begin{equation}
@@
 
-\begin{frame}{Parameters tuning}
+\begin{frame}{Markov random field}
 \begin{equation}
@@
 
+\subsection{Graphical model algorithms}
+
+\begin{frame}{}
+\begin{center}
+{\Huge Graphical model algorithms}
+\end{center}
+\end{frame}
+
 \begin{frame}{Gaussian graphical model}
@@
 \end{equation}
-
 \begin{equation}
 K=\Sigma^{-1}
+\end{equation}
+\begin{equation}
+\log L(K,\boldsymbol{\mu})=\frac{n}{2}\log\det(K)-\dfrac{n}{2}tr(KS)-\dfrac{n}{2}\left(\bar{\textbf{y}}-\boldsymbol{\mu}\right)^{T}\left(\bar{\textbf{y}}-\boldsymbol{\mu}\right)
 \end{equation}
@@
 \end{equation}
+\begin{center}
 Maximized likehood
+\end{center}
 \begin{equation}

### 2021-12-16 OLD/Manuscript/manuscript.tex
PUB

--- 2021-12-03 Manuscript/manuscript.tex
@@
 \usepackage[T1]{fontenc}		% font coding
+\usepackage{pifont}
 
@@
 that underlie the probabilistic methods here used: these will be
-exposed following the approaches of Murphy \cite{murphy2012machine} , Koller
-et al.
-\citep{koller2009probabilistic}, Højsgaardand et al.
-\citep{hojsgaard2012graphical} and Russel et al.
-\citep{russell2010artificial}.
-Furthermore, a rapid overview of the main concepts of information theory will also be provided following the Cover \citep{cover2006elements}, and MacKay \citep{mackay2003information} approaches.
+exposed following the approaches of Murphy \cite{murphy2012machine} , Kolleret al.
+\cite{koller2009probabilistic}, Højsgaardand et al.
+\cite{hojsgaard2012graphical} and Russel et al.
+\cite{russell2010artificial}.
+Furthermore, a rapid overview of the main concepts of information theory will also be provided following the Cover \cite{cover2006elements}, and MacKay \cite{mackay2003information} approaches.
 This is because some of its quantities in the preliminary analysis of the dataset were used.
@@
 
-Where $w_{st}=J$ represent the coupling strength between two neighbour site.  The collective state is described by
+Where $w_{st}=J$ represent the coupling strength between two neighbour site.  The collective state is described by \cite{murphy2012machine}:
 
@@
   \end{minipage}
-\caption{Comparison between the density distributions of hazardous (red) and non-hazardous (light blue) asteroids for a selected set of features that, according to the theory, are interesting.}
+\caption{Comparison between the density distributions of hazardous (red) and non-hazardous (light blue) asteroids for a selected set of features that, according to the theory, are interesting.
+Plot obtained with the ggplot2 package \cite{ggplot2}}
 \label{Density_relevant}
@@
 The other side of the coin of this connection is the positive relationships within the semimajor axis and the eccentricity.
-Furthermore, the positive relationship between the mean motion and the semimajor axis is correct: this is the result of the 3$^{th}$ Kepler law.
 Finally the connections with the Jupiter Tisserand Invariant comes directly from the definition of this parameter.
-A full explanation of these connections is provided in the Appendix A.
-Thus it can be said that the model produced is consistent with the astronomic laws.
+Such arguments can be much more clear with the equations and plots provided in Appendix A.
+Thus it can be said that the model produced is almost consistent with the astronomic laws.
 On this basis, the performances of the model in terms of confusion matrix, ROC (Receiver operating characteristic) curve and $\phi$ coefficient (also known as Matthews correlation coefficient ) were evaluated (for this purpose the Caret R package \cite{kuhn2008building,caret} was used).
@@
 However, a direct link from the volume parameters (and thus from the mass) with the orbital ones is not acceptable.
-In principle, one can set these links as forbidden in the algorithm (blacklist), but in the author view, a model where the result is obtained without boundaries should be preferred to those obtained with a large number of boundaries.
-Therefore we reject the mmod and minforest ones favouring the mgm, which was obtained without boundaries.
+In principle, one can set these links as forbidden in the algorithm (blacklist), but in the author view, a model where a better\footnote{In terms of consistency with astronomic laws} result is obtained without boundaries should be preferred to those obtained with a large number of constraints.
+Therefore we reject the mmod and minforest ones favouring the mgm, obtained without constraints (blacklist and/or whitelist) .
 
@@
 \includegraphics[width=1.5\textheight]{Figures/minforest.pdf}
-\caption{The graphical mixed interaction model obtained with the gRapHD package \cite{de2009high} with a stepwise algorithm. The plot was obtained with the \cite{qgraph} package }
+\caption{The graphical mixed interaction model obtained with the gRapHD package \cite{de2009high} with a stepwise algorithm. The plot was obtained with the qgraph \cite{qgraph} package }
 \label{minforest}
@@
 \section{Machine learning algorithms} The performances of mgm model will be now compared with the one obtained from other ML algorithms.
-These include the random-forest (RF, as implemented in the RandomForest package \cite{rfor}), the support vector machines (SVM, as implemented in the e1071 package \cite{dimitriadou2008misc}), the quadratic discriminant anchiedersialysis (QDA, as implemented in the MASS package \cite{MASin reverseS} and the logistic regression (as implemented in stats package \cite{stats}).
+These include the random-forest (RF, as implemented in the RandomForest package \cite{rfor}), the support vector machines (SVM, as implemented in the e1071 package \cite{dimitriadou2008misc}), the quadratic discriminant analysis (QDA, as implemented in the MASS package \cite{MASS} and the logistic regression (as implemented in stats package \cite{stats}).
 Their performances are reported in Fig.
@@
 
-The GLASSO (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that correctly reproduce the physical connections of the dataset feature.
+The GLASSO (with a penalizing parameter $\rho=0.3$) and the mgm algorithms provided graphical models that almost correctly reproduce the physical connections of the dataset feature.
 In particular, for both models, it can be seen that no one of the orbital parameters is conneted to the diameter of the asteroid, as stated by theory.
@@
 \chapter{Appendix B: R code} Here I provide the code which I run to obtain the results previously showed.
-For its writing I took in consideration the examples provided in the  packages documentation, the Prof.Nicolussi lectures and he examples provided in the Hojsgaard textbook  \cite{hojsgaard2012graphical}.
+For its writing I took in consideration the examples provided in the  packages documentation, the Prof.Nicolussi lectures and the examples provided in the Hojsgaard textbook  \cite{hojsgaard2012graphical}.
 

### 2021-12-16 OLD/Manuscript/manuscript_SAVE.tex
PUB

% !TEX TS-program = pdflatex
% !TEX root = Tes.tex
% !TEX spellcheck = en-EN

\documentclass[12pt,%                      % corpo del font principale
               a4paper,%                   % A4 papers
               %twoside,openright,%         % twoside with free right side
               oneside,openany,%           % one side
               titlepage,%                 % use a titlepage
               headinclude,footinclude,%  % header and foot header
               BCOR5mm,%                   % rilegatura di 5 mm
               cleardoublepage=empty,%     % empty pages with no header and foot
               tablecaptionabove,%         % table caption above tables
               floatperchapter,
               ]{scrreprt}                 % KOMA-Script report class;

\usepackage{braket}
\usepackage{changepage}
\usepackage[english]{babel}	% latest language is predefined
\usepackage[T1]{fontenc}		% font coding

\usepackage{indentfirst}		% indent first paragraph of each section
\usepackage{mparhack,fixltx2e,relsize}	% fancy typographies stuff

\usepackage[eulerchapternumbers,%	% chapter font Euler
            subfig,%			% in subfig objects
            beramono,%			% Bera Mono as fixed spacing font
            eulermath,%			% AMS Euler as math font
            pdfspacing,%		% improves line filling
            listings,%			% code output
 %          parts,%			% uncomment for a document divided in parts
            listsseparated,
            ]{classicthesis}		% style ClassicThesis
%\setlength{\cftbeforeloftitleskip}{100pt}
%\renewcommand{\cftbeforeloftitleskip}{1000pt}

\usepackage{arsclassica}		% modifies some aspects of ClassicThesis package
\let\marginpar\oldmarginpar		% for margin notes with \todonotes (overwise conflict with the new definition of \marginpar in classic thesis)
\usepackage[shadow]{todonotes}			% for margin notes and comments

\usepackage{bookmark}			% bookmarks

%*********************************************************************************
% Bibliography
%*********************************************************************************
%%\usepackage[style=authoryear,hyperref,backref,natbib, ,maxcitenames=1, mincitenames = 1, citestyle=authoryear-comp, backend=biber,sortcites,sorting=ynt]{biblatex}
%\usepackage[style=numeric,hyperref,backref,natbib]{biblatex}

%*********************************************************************************
% Graphics
%*********************************************************************************
\usepackage{graphicx}			% images
\usepackage{subfigure}
\usepackage{wrapfig}
\usepackage{tikz}
\usetikzlibrary{mindmap,trees}
\usetikzlibrary{backgrounds}
\usepackage{verbatim}
\usepackage[dvipsnames]{xcolor}
% \usepackage{morefloats}
%\usepackage{chngcntr}
\usepackage{pdfpages}

%*********************************************************************************
% Tables
%*********************************************************************************
\usepackage{tabularx}			% table of predefined length
\usepackage{siunitx}
\usepackage{pbox}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{rotating}
\usepackage{hhline}
\setlength\tabcolsep{3pt}
\usepackage{changepage}
% \usepackage[showframe=true]{geometry}

%*********************************************************************************
% Mathematics and symbols
%*********************************************************************************
\usepackage{amsmath, amssymb, amsthm}	% mathematics stuff
\usepackage{mathrsfs}
\usepackage{calc}
\usepackage{algorithmic}
\usepackage[ruled]{algorithm}
\usepackage{latexsym}
\usepackage[geometry]{ifsym}
\usepackage{mathabx}
\usepackage{pifont}

%*********************************************************************************
% Personal
%*********************************************************************************
\usepackage[font=itshape]{quoting}			% fancy quotation packages. [font=small] old option
\usepackage[english]{varioref}		% complete reference package
\usepackage{hyperref}
\usepackage{url}
\usepackage[intoc, english, noprefix]{nomencl}	%for list of symbols
\usepackage[normalem]{ulem}
\usepackage{chemfig} %for chemical formulas
\usepackage{eurosym}			% euro symbol
\usepackage{epigraph}
\usepackage{calligra}
\usepackage{soul}
\usepackage[utf8]{inputenc}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\setlength{\epigraphwidth}{\textwidth}

%*********************************************************************************
% Calling personal settings and making nomenclature
%*********************************************************************************
\input{custom-commands}
\input{general-settings}  % general custom settings (margins etc)

% \makenomenclature
% \renewcommand{\nomname}{List of Symbols and Abbreviations}

\bibliographystyle{plain}

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE AND AUTHOR(S)
%----------------------------------------------------------------------------------------

\title{\normalfont\spacedallcaps{Association analysis and tweets}} % The article title

\subtitle{Project for the course Algorithms for massive datasets (DSE)} % Uncomment to display a subtitle

\author{
  Martina Corsini \\
  Marzio De Corato
}

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS & LISTS OF FIGURES AND TABLES
%----------------------------------------------------------------------------------------

\maketitle % Print the title/author/date block

%\setcounter{tocdepth}{2} % Set the depth of the table of contents to show sections and subsections only

%\listoffigures % Print the list of figures%

%\listoftables % Print the list of tables

\newpage

\epigraph{
\textit{Facts are stubborn things, but statistics are pliable}\\ Mark Twain
}

\newpage

\epigraph{
\textit{“We declare that this material, which we now submit for assessment, is entirely our own work and has not been taken from the work of others, save and to the extent that such work has been cited and acknowledged within the text of our work. We understand that plagiarism, collusion, and copying are grave and serious offences in the university and accept the penalties that would be imposed should we engage in plagiarism, collusion or copying. This assignment, or any part of it, has not been previously submitted by us or any other person for assessment on this or any other course of study.“}
}

\newpage

\section*{\begin{center}
Abstract
\end{center}}
The aim of this research is to perform an association analysis on several textual datasets, each of which is made up of tweets that contain a specific hashtag, so to understand which words or groups of words are linked with the hashtag itself and the relevance of these connections. The datasets are extracted from Twitter exploiting the Twitter APIs, that make it possible to access public Tweets by searching for specific hashtags.  After a pre-processing phase of data clean-up, in which all characters are converted to lowercase and stop words and punctuation marks are removed, and tokenization,  two algorithm, implemented in Spark \cite{zaharia2016apache}  and MLxtend \cite{raschka2018mlxtend} , are used in order to reach the goal, the FP-growth and the A-priori algorithms. Depending on the technicality and specificity levels of the hashtags used to retrieve the dataset, the words and the groups of words associated to a dataset can belong to only one or several fields. Such behaviour is well-captured by the shape of the associated support distribution: therefore, it could be possible to use this characteristic to gain a first assessment of the quality of information retrieved with a particular hashtag.

\newpage

\tableofcontents % Print the table of contents

\newpage % Start the article content on the second page, remove this if you have a longer abstract that goes onto the second page

\newpage
%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\chapter{Introduction}
This paper is organized according to the following scheme. A first section will provide a rapid theoretical overview, based on cited textbooks, of the concepts and the techniques used in this work. In particular, we will take into account the many steps involved in the process of preparing text data for use, such as the cleaning and the tokenization; subsequently, we will consider the theoretical framework of the association analysis and the algorithms involved, with a particular focus on its connection with the information theory and on the possibility to improve memory management.  A second section will describe how our datasets were generated, describing the process of extraction of tweets with common hashtags from Twitter through the Twitter APIs and the ratio behind the choice of the hashtags, namely to have several levels of specificity and technicality and to have datasets concerning different topics. The last paragraph will contain a discussion of the results obtained for each dataset. Finally, we will propose a possible way to obtain a first approximation about the quality of the information obtained from a hashtag.

\chapter{Theoretical Framework}
\section{Textual documents mining}

The kind of data one gets from social media is usually unstructured. It contains, for example, a vast number of stop words and symbols, but also typos and abbreviations that need to be cleaned up so that a model can grasp the truly interesting associations between the words in the texts. The reliability of the model is highly dependent upon the quality of the data. The process of preparing text data for the analysis is called text pre-processing. Many steps can be involved in this task.
First of all, it may be helpful to get rid of unhelpful parts of the data, or noise, by converting all characters to lowercase, removing stop words and punctuation marks. In particular, removing stop words (the most common words like articles, prepositions, conjunctions etc.) allows the model to consider only key features because these words typically do not contain much information \cite{etaiwi2017impact}. Many variations of those words do not bring any new information and create redundancy, ultimately bringing ambiguity when analyzing the critical information. Moreover, it is crucial to remove stop words since this process reduces the dataset size. Concerning the Python framework, NLTK (Natural Language Toolkit) \cite{bird2009natural}  has a list of stop words stored in 16 different languages. It is also possible to add other words in the list according to the task we perform and the goal we want to achieve.
Finally, text must be tokenized. Tokenization is the process of splitting a text document into smaller units, such as words, numbers or punctuation marks, that are called tokens.  Finally, in order to make our analysis, we must vectorize the text, for example using the bag of words model, in which each tweet is represented as a word-count vector, with a size equal to the number of elements in your vocabulary. It is important to note that the resulting matrix will be sparse since most of its elements will be zero.

\section{Association Analysis}
In this section, we are going to recap the main concepts and algorithms involved with the market basket analysis. After a brief introduction about the main indices involved, in which we will follow the Leskovec et al. \cite{leskovec2020mining} as well the Tan et al. approaches \cite{tan2019introduction}, we will report how such indices can be connected with the information theory. For this purpose, we will follow the Yao paper \cite{yao2003information}; furthermore, concerning the rapid exposition about the main concepts of the information theory, we will consider the approaches of the two standard textbooks about this topic, written by D.MacKay \cite{mackay2003information}, and T.Cover \cite{cover2006elements}. Finally, following the Tan et al. approaches \cite{tan2019introduction}, we will analyze and compare the two algorithms used in this research for the association analysis: the A-priori and the FP-growth.

\subsection{Basic concepts}

Given a large dataset, a first-order approach to mine useful information from it is to perform an association analysis: the underlying idea of this approach is that itemsets that appear together may\footnote{But not necessary} reveal an interesting relationship. Such analysis is commonly referred to as basket analysis, since one of the first and common uses of this technique is to retrieve itemsets that are bought together; however, such technique can be used in a number of field, such as for the analysis of the earth science data or even for astronomical purposes \cite{borah2019rare}. It is worth noting that the patterns discovered with this technique can contain spurious correlations. Therefore, the conclusions obtained can be seriously affected by the Simpson paradox \cite{simpson1951interpretation,blyth1972simpson}. In essence, the Simpson paradox states that the existence and the order of causality can depend by the variables considered: the removal or the inclusion of them can seriously alter the inference obtained. Thus, the conclusions obtained from an association analysis should be validated by a model and not taken recklessly for granted. With this important premise in mind, let us move to consider how this analysis works. A market basket dataset can be represented with a binary table in which each row \textit{i} represent a case to be analyzed (transaction) and each column \textit{j} an element that may be present in the transaction. The matrix element $ m_{ij} $ is set to 1 if the object j is present in the case (transaction) i, otherwise it is equal to 0 (perhaps a boolean TRUE/FALSE can also be considered). Once defined how the transaction can be described into a compact form, we can move to the description of indices that highlight the important associations. For this purpose, we label with $I=\left\lbrace i_{1},i_{2},...,i_{d} \right\rbrace$ the set of all itemset and with $T=\left\lbrace t_{1},t_{2},...,t_{d} \right\rbrace$ the set of all transactions: in this way, the support count $\sigma(X)$ can be defined as follow \cite{tan2019introduction}:

\begin{equation}
\sigma(X)= \left| \left\lbrace t_{i} | X \subseteq t_{i}, t_{i} \in T \right\rbrace \right|
\end{equation}

Such quantity can be normalized with the overall number of transaction N; the resulting value \cite{tan2019introduction}:

\begin{equation}
s(X)= \frac{\sigma(X)}{N}
\end{equation}

is called support index. If this value for an item X is greater than a certain threshold arbitrarily chosen called \textit{minsup}, item X is said to be frequent. By now, we focused on single items, but our propose is to find a connection between them: in particular, we are interested in finding associations, or implications, $X \rightarrow Y $ with the condition that $X \cap Y = \emptyset $. These can be measured with the \textit{support} index  \cite{tan2019introduction}:

\begin{equation}
s(X \rightarrow Y) = \dfrac{\sigma(X \cup Y)}{N}
\end{equation}

as the formula shows, such an index measures how often a rule can be applied for the chosen dataset. In addition, one can replace the denominator of the following expression with the support index of X, thus giving \cite{tan2019introduction}:

\begin{equation}
c(X \rightarrow Y) = \dfrac{\sigma(X \cup Y)}{\simgm(X)}
\end{equation}

this index is called \textit{confidence} and measures the frequency of item Y in the transactions that contain item X. The meaning of these two indices will become more evident within the information theory framework provided in the following paragraph: here we would anticipate that, as pointed out by Tan et al.\cite{tan2019introduction} a low support signals an association that is likely to happen by chance, while the second one provides an estimate of the conditional probability of Y given X. Finally, another helpful index for the assessment of an association rule is the \textit{lift} defined as \cite{tuffery2011data}:

\begin{equation}
lift = \frac{P\left( A \cap B \right)}{P(A) \cdot P(B)}
\end{equation}

If the lift has an absolute value vastly different from 1, we have an association rule that largely prevails on the product between the ratios of the two separate events; on the contrary, if the lift value is close to one, the value of the association is poor. As we will see in the paragraph dedicated to the information theory, the lift index is connected to the point mutual information concept. On this basis, following the Tan et al. approach, we can state the Association Rule discovery problem as \cite{tan2019introduction}:

\begin{quote}
\textbf{Association Rule Discovery}: Given a set of transactions T, find all the rules having support $\ge$ \textit{minsup} and confidence $\ge$ minconf where \textit{minsup} and \textit{minconf} are the corresponding support and confidence thresholds
\end{quote}

What is the computational effort required for this task ? It can be proven that with a dataset that contain d items the possible association rules are \cite{tan2019introduction}:

\begin{equation}
R=3^{d}-2^{d+1}+1
\end{equation}

As one can point out, this number grows fast with $d$; thus it is crucial to prune the rules that will never have sufficient support and confidence values in order the make the association rule discovery task affordable. This aim can be achieved with the algorithms that will be reviewed at the end of this section.

\paragraph{Memory management}

In order to store the pair counters, one could build a matrix. However, in this way there is a waste of memory that amount to half the space, because we are not interested in the order of the association.
A possible solution can be the linearisation of the matrix, so a one-dimensional representation of the association with no wasted space. Furthermore, the efficiency in finding pairs is maintained using the formula k = (i-1) (n-1/2) + j- i, which makes us able to know the correct position of the counter.
If the matrix is sparse, another possible and more space-efficient alternative is to represent counts of pairs as triples (i, j, c). An index structure such as a hash table allows us to efficiently find the triple for (i, j). It is important to note that we can have collisions, so using the linearised matrix is preferable if we have no space problems.

\subsection{Information theory and its connections with association analysis}

As stated by Cover textbook, \cite{cover2006elements}, the information theory comes out from the issue of how much and at which rate data can be compressed. Indeed, its initial purpose was to quantify how much a signal can be compressed and send via a finite communication channel. However, as its founder, Claude Shannon, discovered, such problem that was initially conceived within the communication theory involves many fields of science such as statistical physics, computer science, statistical inference, and economics: a bird's-eye view about the formidable ability of this theory to combine different disciplines is given in Fig. \ref{Information_theory_connections}. The fundamental concept of information theory is entropy. In this context, such quantity, which was originally conceived in physics to explain the lack of symmetry for processes \footnote{like the expansion of gas} concerning the time-reversed transformation, measures the degree of uncertainty of a random variable. Formally, given a random variable $X$ with the accessible states (alphabet) $\mathcal{X}$ and probability distribution $p(x)= Pr\left\lbrace X=x\right\rbrace, x \in \mathcal{X} $ \footnote{It is worth noting that this is the standard definition of a stochastic variable \cite{van1992stochastic}}, the entropy of this variable is defined as \cite{cover2006elements}:

\begin{equation}
H(X)=-\sum_{\mathcal{X} \in \chi}p(x)\log p(x)
\end{equation}

where the logarithm is in base 2. With this definition, the entropy, which is usually referred to as Shannon entropy, is measured in bits. Once defined this quantity for a single variable, we can move to two variables. The joint entropy $H(X,Y)$ of a discrete random variables with joint distribution $p(x,y)$ is given by the following expression \cite{cover2006elements}:

\begin{equation}
H(X,Y)=-\sum_{x \in \mathcal{X}}\sum_{y \in \mathcal{Y}}p(x,y)\log p(x,y)
\end{equation}

Furthermore we can define the conditional entropy as \cite{cover2006elements}:

\begin{equation}
H(Y|X)=-\sum_{x \in \mathcal{X}}\sum_{y \in \mathcal{Y}}p(x,y)\log p(x|y)
\end{equation}

The joint entropy and the conditional entropy are mutually bounded by the following relation \cite{cover2006elements}:

\begin{equation}
H(X,Y)=H(X)+H(Y|X)
\end{equation}

Finally we can measure the mutual information of two random variables as

\begin{equation}
I(X,Y)=\sum_{x \in \mathcal{X}}\sum_{y \in \mathcal{Y}} p(x,y)\log \frac{p(x,y)}{p(x)p(y)}
\end{equation}

Note that the entropy of the single variable are bounded to the mutual information according to the following expression \cite{cover2006elements}:

\begin{equation}
I(X,Y)=H(x)+H(y)-H(X,Y)
\end{equation}

 As can be pointed out such quantity is an average between the two alphabet, in our case where we are interested in the single associations, the point-wise mutual information should be considered \cite{cover2006elements}:

\begin{equation}
i(X,Y)=\log \frac{p(x,y)}{p(x)p(y)}
\end{equation}

It is worth mentioning that the mutual information can be recast in a more elegant  way with the Kullback-Leibler distance. Lets focus for while on this crucial concept. Given two probability distribution p(x) and q(x) the Kullback-Leibler distance or the relative entropy is defined as \cite{cover2006elements}:

\begin{equation}
D(p||q)=\sum_{x \in \mathcal{X}} p(x)\log \frac{p(x)}{q(x)}
\end{equation}

The meaning of this quantity, not by chance called distance, is that given a distribution probability p, it would describe it with another distribution q we would have an excess of entropy equal to $D(p||q)$ concerning the entropy $H(p)$. In this way, the KL distance is an inefficiency measure connected with the distribution $p$. Moreover, as stated by the Gibbs inequality \cite{mackay2003information}:

\begin{equation}
D(p||q)\geq 0
\end{equation}

the inefficiency can always be positive or at least null\footnote{This is the equivalent, in the information theory framework, of the second principle of thermodynamics. This is particularly clear looking to the M.Plank statement: \textit{Every process occurring in nature proceeds in the sense in which the sum of the entropies of all bodies taking part in the process is increased. In the limit, i.e. for reversible processes, the sum of the entropies remains unchanged} \cite{planck1900theory}}. This is the formalization, and more importantly, the quantification, of the argument that given a random variable (the weather of tomorrow) we cannot reach with a forecast an entropy lower concerning the phenomenon itself. We would remark that these measures, which were initially used for the inference of a transmitted signal, can be applied basically for every inference problem, including the object of this work: the inference of the association pattern within a dataset. Once defined and clarified the fundamental concept of DL distance, we could recast the mutual information expression into the elegant form \cite{mackay2003information}:

\begin{equation}
I(X,Y)=D(p(x,y)||p(x)p(y))
\end{equation}

Based on this theoretical framework, following the Yao paper \cite{yao2003information}, we can recognize in the logarithm (with base 2) of lift index, described before, the point-wise mutual information described before. This allows, as one expects, to use the formalism and the results achieved by the information theory in data mining. We say as one expects because we can think of the dataset as a communication channel in which nature encodes the association rules: the data mining process can be seen as the decoding of these association rules given a dataset (a remarkable analysis on how the information theory concepts can be used in the data mining context can be found in \cite{zaki2014data}). A remarkable description of the connections between learning algorithms and the information theory is provided by D.MacKay \cite{mackay2003information} and by E.T.Jaynes \cite{jaynes2003probability} .

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/Information_theory_connections.jpg}
\caption{Connections of information theory with different science fields. Image taken from \cite{cover2006elements}}
\label{Information_theory_connections}
\end{center}
\end{figure}

\subsection{Algorithms}

Before reviewing the algorithms used in this work for the frequent item generation, we would introduce the itemset lattice: in each n-level (line) the possible combination between n elements is given (starting from 0, the void ensemble). Furthermore, each cell is connected with the cell that contains a subset of itself. The advantage of such representation will become evident during the A-priori description. There are three way by which the computational costs can be reduced: the first possibility is to reduce the number of candidates (this is the underlying idea of A-priori principle); the second option is to reduce the number of comparisons by exploiting an intelligent data structure (this is the idea of the FP growth algorithm); finally, one can reduce the number of transactions. In this work, we will consider the algorithms that use the first two approaches; thus, we would limit our theoretical analysis to these two only.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/itemset_lattice.jpg}
\caption{An example of itemset lattice for $I=\left\lbrace a,b,c,d,e\right\rbrace$. Image taken from \cite{tan2019introduction}}
\label{itemset_lattice}
\end{center}
\end{figure}

\paragraph{A priori}

The A-priori algorithm is based on the following principle, as stated by Tan et. al \cite{tan2019introduction}:

\begin{quote}
\textbf{Apriori principle}: If an itemset is frequent, then all of its subsets must also be frequent
\end{quote}

The support benefits of the Anti-monotone property \cite{tan2019introduction}:

\begin{quote}
\textbf{Anti-monotone property}: A measure f possesses the antimonotonone property if for every itemset X tha is a proper subset of itemset Y, i.e $X \subset Y $ we have $f(Y) \leq f(X) $
\end{quote}

Therefore, we can prune the candidates by fixing a support threshold and removing those whose subsets are not frequent enough. The crucial point, as shown in Fig. \ref{A-priori-ex} is that the algorithm starts from the smallest subset, and then, if an infrequent item is found, all the larger subset that include it are removed from the candidates. In this way the overall number of candidates is vastly reduced, thus lowering the computational cost required for the association analysis.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/A-priori-ex.png}
\caption{The mechanism of A-priori algorithm: in the upper panel the fact that the $\left\lbrace cde \right\rbrace$ is frequent implies that the subset by wich it is composed are also frequent. On this basis one can implement an algortihm that prunes the larger itemset by investing its subset: if a subset is infrequent all the set that include it are also infrequent (lower panel). This approach considerably reduce the candidates and thus the computational cost of the association analysis. Images taken from \cite{tan2019introduction} }
\label{A-priori-ex}
\end{center}
\end{figure}

\paragraph{FP growth}

The Frequent Pattern (FP) algorithm represents an alternative approach to the A-priori: here, the key concept for reducing computational cost is played by how the data are stored, instead of the candidates' pruning. For this purpose, the dataset is stored in to the FP-tree: this is a diagram, whose construction is explained in Fig. \ref{FP_tree} ,  in which each node represent a label of an item accompanied by a counter that reports the number of transactions mapped on the given path. The key point is that different transactions may have elements in common: thus, since some nodes in the diagram may overlap, the overall number of nodes is reduced (more precisely, we have an overlap between two paths if they have a common prefix). As we will see, the computation cost for the association analysis, given a FP-tree, depends on the number of distinct nodes: if this number is reduced, the computational effort is reduced.  Now that we described the core concept for the data compression, we move on to how the frequent itemset are generated. This procedure is performed through a bottom-up approach in which the algorithm first evaluates the support for a single suffix, then evaluates the support of two element suffix, then for three element suffix, and so on. The evaluation of the support can be quickly obtained by evaluating the paths that contain a chosen node (this is why also the dashed lines were included in the diagram reported in Fig. \ref{FP_tree}). As illustrated in Fig \ref{FP_gen_1}, once a particular suffix is fixed, the algorithm generates the conditional trees that are nothing more than the part of the FP tree that verifies a particular condition i.e., to end with a particular suffix. The key point that represents one of the significant advantages of the FP algorithm is that once a particular suffix is chosen, the conditional tree that is generated provides a new subset of problems (see Fig. \ref{FP_gen_1}) in this way, the complex problem of the association analysis is attacked with a divide and conquer strategy. At this point is clear how the run-time of the FP algorithm is strictly dependent on the number of distinct nodes: if the overlap is low, the performance of this approach degrades. On the contrary, if the compression of the transaction dataset into the FP-tree is high, this method can outperform the A-priori algorithm by a different order of magnitude \cite{tan2019introduction}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/FP_tree.jpg}
\caption{The implementation of an FP-tree given a transaction. Note the dashed lines which provide the paths that connect identical elements in the diagram, but are not in the same path due to the fact that there is no common prefix.  Image taken from \cite{tan2019introduction} }
\label{FP_tree}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/FP_gen_1.jpg}
\caption{The problem of basket analysis is decomposed into different sub-problem for each suffix. Image taken from \cite{tan2019introduction} }
\label{FP_gen_1}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{Figures/FP_gen_2.jpg}
\caption{The conditional FP-trees generated as a condition is imposed. Image taken from \cite{tan2019introduction} }
\label{FP_gen_2}
\end{center}
\end{figure}

\subsection{PCY variant}
The a priori algorithm is characterized by a waste of memory in the first pass of the process: in fact, a considerable portion of RAM is not used, namely the amount of RAM used in the second pass to count the pairs.  The PCY variant of the a priori algorithm uses the remaining part of the memory in the first pass to reduce the candidates. In fact, it constructs a hash table on the first pass, using all main-memory space that is not needed to count the items. Pairs of items are hashed to be mapped into buckets; the buckets are used as counters of the number of times a pair hashed to that bucket. On the second pass, we only have to count pairs of frequent items that hashed to a frequent bucket (one whose count is at least the support threshold) on the first pass.
}

\begin{figure}[ht]
  \fbox{\begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Apriori_memory.png}
    \subcaption{A-priori}
    \vspace{4ex}
  \end{minipage} }
  \fbox{\begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/PCY_memory.png}
    \subcaption{PCY}
    \vspace{4ex}
  \end{minipage}}%%
\caption{Use of memory for a-priori algorithm and for PCY variant. Image taken from \cite{leskovec2020mining}}
\label{Support distribution_ALL}
\end{figure}

\chapter{Dataset generation}

\section{Tweet retrieval}

Twitter and other social media platforms represent an important and updated source of information with regard to the development of our society, since social media platforms generate a vast amount of data daily on a variety of topics. Therefore, it is important to be able to retrieve tweets in order to be able to analyze them. In order to reach this goal, Twitter offers companies, developers, and users programmatic access to Twitter data through API (Application Programming Interfaces). Using the Twitter APIs, developers can access public Tweets by searching for specific keywords \cite{twitter}. To retrieve all tweets with a specific hashtag, it is possible to use the search/tweets GET endpoint. By default, the results return tweets sent in the previous seven days. Tweets are delivered in reverse-chronological order, starting with the most recent.  Therefore, the sample extracted can not be considered a random one.  However, under the condition that there is no particular event from the first to the last tweet extracted that alters the association,\footnote{We verified that this condition was true from different on-line press agencies } we can consider the sample as randomly extracted.  Furthermore, it is also possible to select the language of the tweets. The resulting csv file is available in the following format: hashtag|timestamp|tweet\_text|user\_name|language .

\section{Choice of the dataset}

We chose to work on five datasets composed of 10000 tweets that contain respectively the following hashtag: \textit{freedom, Afghanistan, rare earth, Covid-19 and Pfizer}.  The ratio behind the choice of the hashtag has been to have tweeted with several levels of specificity and technicality, and concerning different topical questions, in particular the Afghanistan crisis and the pandemic situation. The hashtag \textit{freedom} is a generic hashtag that can be linked both with the rise to power of the Taliban regime and with the possibility to make vaccination compulsory.  \textit{delta}, \textit{Pfizer} and \textit{vaccines} are related to the pandemic,  but they show a different level of technicality, since vaccines is a common word while the others two refer to a more technical field, even if they have become very popular in the public debate.  \textit{Afghanistan} and \textit{rare earth } refer to the Afghanistan situation: in this case, the latter shows a high level of specificity since it is about an issue that is not widely known, namely the interest of some countries to exploit Afghanistan’s rare earth metals, some crucial elements in modern technology.

\chapter{Results and discussion}

For each term, we conducted a preliminary analysis on its cleaned document-term matrix: as it is possible to point out from Fig. \ref{Matrix_ALL}, a sparse matrix was obtained, as expected,  for each hashtag.  We then applied the FP-growth algorithm,  as implemented in Spark suite of codes \cite{zaharia2016apache} ,  and the A-priori algorithm,  as implemented in MLxtend suite of codes \cite{raschka2018mlxtend}. With the first one, we retrieved the frequency for each itemset,  the association rule for each basket, and the predictions of the items that are most likely to appear given a basket.  A support threshold equal to 0.03 was chosen. A preview of the results is given in the Tab. \ref{POP_RareEarth}, \ref{POP_Afghanistan}, \ref{POP_Delta}, \ref{POP_Pfizer},\ref{POP_Vaccines} and \ref{POP_Freedom}  for the 5 most popular items, while the associations (the five with the higher lifts) are given in \ref{ASS_RareEarth}, \ref{ASS_Afghanistan}, \ref{ASS_Delta}, \ref{ASS_Pfizer} and \ref{ASS_Vaccine} (note that the table for hashtag freedom is not present since there is no association with a support equal to 0.03). It is worth noting that in the pre-processing phase the hashtag was removed, thus the singleton represents, in this perspective, a binary combination between the word of the hashtag and the singleton itself. For this reason, in our analysis we will consider as multiple association a combination of two or more words, since the hashtag is implicitly included in the association. The full results (in the form of an Excel table) obtained for each hashtag are provided here \cite{Folder},  and a good visual representation of them is provided in Fig. \ref{Support distribution_ALL} and Fig. \ref{WordCloud_ALL}.  Let us start from the Rare earth hashtag, that present a large number of high frequency peaks given by the words \textit{minerals, china, Afghanistan, metals} and \textit{taliban} (see Tab. \ref{POP_RareEarth}): the presence of the words China, Afghanistan and talibans are associated with this hashtag due to the interest of some countries, such as China, to exploit Afghanistan’s rare earth metals. Furthermore, also the multiple association table Tab. \ref{POP_RareEarth}, in which we see a high lift value for the association \textit{lithium} and \textit{copper}, provide intersting information: in fact, this association is due to the fact that these two metals are largely present in Afghanistan together with rare earth \footnote{The reader must be aware that lithium and copper are not rare-earth. A full discussion about the chemistry and the industrial uses of these materials can be found into a general chemistry textbook such as \cite{petrucci2010general}}. In the case of the Afghanistan hashtag, we have a support distribution with a very sharp peak connected to \textit{Taliban} and \textit{Kabul}: while the first one is connected because it is the Afghanistan capital (see Tab. \ref{POP_Afghanistan}), the second one is an effect of Taliban conquer of Afghanistan happened at the end of August 2021, after US troops withdraw. The importance of the US in the political scenario of the Afghanistan is also captured by the words \textit{us}, \textit{biden}, while the association between Afghanistan and \textit{kabulairport} is given both by the importance of this place for the evacuation of the US troops and by the terroristic attack that took place during the evacuation. These words and their associations provide the first peak and the second broader one. Looking to the multiple associations table (Tab. \ref{ASS_Afghanistan}), we can see that the association with larger lift is given by \textit{nowazd} and \textit{usarmy}; this connection is due to the Nowzad Dogs charity, whose supporters provided financial support for the evacuation of the association staff and their animals \cite{Nowzad}. The following association are basically connected with the capture of Kabul by Taliban and the civilian evacuation. Concerning the hashtag delta, we can see that we have an overall sharp peak made firstly by the words \textit{covid19},\textit{covid}, \textit{variant}, \textit{amp} and \textit{vaccine} (see Tab. \ref{POP_Delta}). This findings clearly signal the extensive use of word delta within the pandemic framework, while other fields in which this word appears, such as, for instance, the delta airlines, does not contribute to the spread of this hashtag in the social media. An inspection of the multiple associations between the word delta and other words, reported in Tab \ref{ASS_Delta}, highlights the words \textit{vaccine}, \textit{covid19} and \textit{variant}: however, in contrast to the previous case (hashtag Afghanistan), we can note that the maximum lift is very close to 1, thus the interest of these associations is poor. Moving to the hashtag pfizer, we see that we have a very sharp peak connected with the term \textit{vaccine}, followed by \textit{covid19}, \textit{fda} (food and drug administration) and by the association \textit{covid19,vaccine}. In this case, as we can note looking to Tab \ref{ASS_Pfizer}, we have an interesting lift value for \textit{full,vaccine}: this is the effect of the public debate about the vaccination campaign, also connected with the discussion on the possibility of a third dose, as one can see from the lift value of the association \textit{full, approval}. Concerning the hashtag vaccines, we point out the presence of large peak at the origin connected with the word \textit{covid19}, followed by \textit{covid} and \textit{amp}. With regards to the multiple association, we can see that these has a lift value close to 1. Finally, the distribution connected with freedom is almost horizontal, highlighting that this hashtag is associated with several fields, and therefore a low information content can be retrieved with it. The lack of interesting associations is also captured by the fact that no multiple association with a support equal or larger with respect to 0.03 is present, thus in this case not table is provided. In order to quantify this contrast, we propose to evaluate the area under the curve (remembering that the curves here reported referring only to the itemsets whose support is larger to the chosen threshold): the idea is that if this area is lower, the specificity of the hashtag searched is very low; on the contrary, if the area is high we have a fine specificity (i.e. only a few words are usually combined with a hashtag with the high area). From an information theory point of view, the first case corresponds to a large entropy (every state is equally probable), the second one to low entropy (few probable states). The results of this evaluation for the hashtag here considered is provided in Tab. \ref{TAB_area}: the RareEarth is associated with the higher value, followed by Afghanistan and Pfizer; on the other side, freedom has low specificity since it can be associated with a cumbersome amount of words. In the author's view, the area can be considered as an assessment of the information content obtained by mining tweets with this word: if a word with a high area (above the support threshold) is used, it is associated almost with specific terms. Thus the tweets retrieved used with this hashtag were will more likely to provide useful information. On the other side, a hashtag connected with a low area will provide many spurious tweets, and thus, in the end, no useful information will be retrieved.

\begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Matrix_afghanistan.jpg}
    \subcaption{Afghanistan}
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Matrix_delta.jpg}
    \subcaption{Delta}
    \vspace{4ex}
  \end{minipage}
    \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Matrix_vaccine.jpg}
    \subcaption{Vaccine}
    \vspace{4ex}
  \end{minipage}
      \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Matrix_pfizer.jpg}
    \subcaption{Pfizer}
    \vspace{4ex}
  \end{minipage}
        \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Matrix_rareearth.jpg}
    \subcaption{rareearth}
    \vspace{4ex}
  \end{minipage}
          \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Matrix_freedom.jpg}
    \subcaption{freedom}
    \vspace{4ex}
  \end{minipage}

\caption{Document-term matrix for the different hashtag. The high quality ones are available at \cite{Folder}}
\label{Matrix_ALL}
\end{figure}

\begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Support_distrubution_Afghanistan.png}
    \subcaption{Afghanistan}
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Support_distrubution_delta.png}
    \subcaption{Delta}
    \vspace{4ex}
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Support_distrubution_Vaccine.png}
    \subcaption{Vaccines}
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Support_distrubution_pfizer.png}
    \subcaption{Pfizer}
    \vspace{4ex}
  \end{minipage}
   \begin{minipage}[b]{0.5\linewidth}
   \centering
    \includegraphics[width=.9\linewidth]{Figures/Support_distrubution_Rareearth.png}
    \subcaption{Rareearth}
    \vspace{4ex}
  \end{minipage}
     \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/Support_distrubution_Freedom.png}
    \subcaption{Freedom}
    \vspace{4ex}
  \end{minipage}

\caption{Support distribution for the different hashtag}
\label{Support distribution_ALL}
\end{figure}

\begin{figure}[ht]
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/WordCloud_Afghanistan.png}
    \subcaption{Afghanistan}
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/WordCloud_delta.png}
    \subcaption{Delta}
    \vspace{4ex}
  \end{minipage}
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/WordCloud_Vaccine.png}
    \subcaption{Vaccines}
    \vspace{4ex}
  \end{minipage}%%
  \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/WordCloud_pfizer.png}
    \subcaption{Pfizer}
    \vspace{4ex}
  \end{minipage}
   \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/WordCloud_Rareearth.png}
    \subcaption{Rareearth}
    \vspace{4ex}
  \end{minipage}
     \begin{minipage}[b]{0.5\linewidth}
    \centering
    \includegraphics[width=.9\linewidth]{Figures/WordCloud_Freedom.png}
    \subcaption{Freedom}
    \vspace{4ex}
  \end{minipage}
\caption{WordCloud for the different hashtag}
\label{WordCloud_ALL}
\end{figure}

\begin{table}[]
\caption{Most popular items for hashtag \textit{Afghanistan}}
\begin{center}
\begin{tabular}{l|l}
        & \textbf{Support} \\ \hline
Taliban & 1803 \\ \hline
Kabul   & 1675 \\ \hline
Us      & 1598 \\ \hline
Biden   & 1135 \\ \hline
Amp     & 1016 \\ \hline
Afghan  & 914
\label{POP_Afghanistan}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Most popular items for hashtag \textit{Delta}}
\begin{center}
\begin{tabular}{l|l}
        & \textbf{Support} \\ \hline
Covid19 & 2728    \\ \hline
Covid   & 1774    \\ \hline
Variant & 1721    \\ \hline
Amp     & 1189    \\ \hline
Vaccine & 968     \\ \hline
Cases   & 836
\label{POP_Delta}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Most popular items for hashtag \textit{Freedom}}
\begin{center}
\begin{tabular}{l|l}
       &\textbf{ Support} \\ \hline
Amp    & 872     \\ \hline
Get    & 802     \\ \hline
People & 709     \\ \hline
Free   & 636     \\ \hline
Us     & 532     \\ \hline
Dont   & 506
\label{POP_Freedom}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Most popular items for hashtag \textit{Pfizer}}
\begin{center}
\begin{tabular}{l|l}
         & \textbf{Support} \\ \hline
Vaccine  & 3821    \\ \hline
Covid19  & 2607    \\ \hline
Fda      & 1847    \\ \hline
Covid19  & 1467    \\ \hline
Approval & 1120    \\ \hline
Moderna  & 1086
\label{POP_Pfizer}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Most popular items for hashtag \textit{Vaccines}}
\begin{center}
\begin{tabular}{l|l}
           & \textbf{Support} \\ \hline
Covid19    & 4007    \\ \hline
Covid      & 1596    \\ \hline
Amp        & 1080    \\ \hline
Get        & 965     \\ \hline
People     & 755     \\ \hline
Vaccinated & 740
\label{POP_Vaccines}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Most popular items for hashtag \textit{RareEarth}}
\begin{center}
\begin{tabular}{l|l}
            & \textbf{Support} \\ \hline
Minerals    & 3215    \\ \hline
China       & 2948    \\ \hline
Afghanistan & 2400    \\ \hline
Metals      & 1804    \\ \hline
Taliban     & 1551    \\ \hline
Us          & 1217
\label{POP_RareEarth}
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Association rule for the hashtag \textit{Afghanistan}}
\begin{tabular}{l|l|l|l|l}
\textbf{Antecedent} & \textbf{Consequent}  & \textbf{Lift} & \textbf{Confidence} & \textbf{Support}\\ \hline
nowzad       & usarmy       & 21.16 & 0.68       & 0.03    \\ \hline
nowzad       & penfarthing  & 16.80 & 0.72       & 0.03    \\ \hline
kabulattack  & kabulairport & 9.01  & 0.70       & 0.03    \\ \hline
airport      & kabul        & 4.60  & 0.77       & 0.05    \\ \hline
kabulairport & kabul        & 3.49  & 0.27       & 0.05
\label{ASS_Afghanistan}
\end{tabular}
\end{table}

\begin{table}[]
\caption{Association rule for the hashtag \textit{Delta}}
\begin{tabular}{l|l|l|l|l}
\textbf{Antecedent} & \textbf{Consequent}  & \textbf{Lift} & \textbf{Confidence} & \textbf{Support }\\ \hline
vaccine    & covid19    & 1.65 & 0.45       & 0.04    \\ \hline
covid19    & cases      & 1.40 & 0.12       & 0.03    \\ \hline
covid19    & variant    & 1.38 & 0.24       & 0.06    \\ \hline
covid19    & covid      & 1.15 & 0.20       & 0.06    \\ \hline
amp        & covid19    & 0.95 & 0.26       & 0.03
\label{ASS_Delta}
\end{tabular}
\end{table}

\begin{table}[]
\caption{Association rule for the hashtag \textit{Pfizer}}
\begin{tabular}{l|l|l|l|l}
\textbf{Antecedent} & \textbf{Consequent}  & \textbf{Lift} & \textbf{Confidence} & \textbf{Support}\\ \hline
full,vaccine     & approval   & 8.19 & 0.92       & 0.04    \\ \hline
full             & approval   & 7.08 & 0.79       & 0.05    \\ \hline
approved,vaccine & fda        & 3.97 & 0.73       & 0.04    \\ \hline
approval,covid19 & vaccine    & 2.18 & 0.83       & 0.03    \\ \hline
full, approval   & vaccine    & 2.04 & 0.78       & 0.04

\label{ASS_Pfizer}
\end{tabular}
\end{table}

\begin{table}[]
\caption{Association rule for the hashtag \textit{Vaccine}}
\begin{tabular}{l|l|l|l|l}
\textbf{Antecedent} & \textbf{Consequent}  & \textbf{Lift} & \textbf{Confidence} & \textbf{Support }\\ \hline
covid19    & please     & 1.96 & 0.08       & 0.03    \\ \hline
news       & covid19    & 1.91 & 0.76       & 0.03    \\ \hline
covid19    & well       & 1.84 & 0.08       & 0.03    \\ \hline
covid19    & pandemic   & 1.64 & 0.09       & 0.03    \\ \hline
covid19    & need       & 1.56 & 0.10       & 0.04
\label{ASS_Vaccine}
\end{tabular}
\end{table}

\begin{table}[]
\caption{Association rule for the hashtag \textit{RareEarth}}
\begin{tabular}{l|l|l|l|l}
\textbf{Antecedent} & \textbf{Consequent}  & \textbf{Lift} & \textbf{Confidence} & \textbf{Support }\\ \hline
lithium     & copper      & 6.07 & 0.27       & 0.03    \\ \hline
deposits    & lithium     & 4.56 & 0.53       & 0.03    \\ \hline
russia      & china       & 2.99 & 0.88       & 0.04    \\ \hline
china       & afghanistan & 2.62 & 0.13       & 0.04    \\ \hline
afghanistan & biden       & 2.44 & 0.14       & 0.03
\label{ASS_RareEarth}
\end{tabular}
\end{table}

\begin{table}[]
\begin{center}
\caption{Area under the support distribution curve normalized by the lowest value (Freedom). Note that the items with support lower with respect to the threshold 0.03 were discarded}
\begin{tabular}{l|l}
            & Area  \\ \hline
RareEarth   & 6.14 \\ \hline
Afghanistan & 4.60  \\ \hline
Pfizer      & 2.86  \\ \hline
Vaccines    & 2.85  \\ \hline
Delta       & 2.63  \\ \hline
Freedom     & 1.00     \\
\label{TAB_area}
\end{tabular}
\end{center}
\end{table}

\chapter{Conclusion}
From the basket analysis that we conducted on the dataset generated from a set of hashtags, we saw that these involve words that can span from very specific to many fields. The first case is what happens with the  \textit{RareEarth} hashtag, the second one with the \textit{Freedom} one. The shape of the support distribution captures such behaviour: a horizontal one (high entropy) highlight a word that is used in an extensive set of context, while a distribution with a high peak at the origin (low entropy) highlight a word that is used into a very reduced context. Since the quality of information is linked to the specificity of words used with it, we propose to use the overall area under the support distribution (once the area below the support is removed) to assets numerically the information quality connected with a particular hashtag: in this framework, the behaviour of \textit{RareEarth} hashtag is emblematic since it is more specific and technical with respect to \textit{freedom} that can be used almost in every context.

\bibliographystyle{unsrt}

\bibliography{sample.bib} % The file containing the bibliography

\newpage

%----------------------------------------------------------------------------------------

\end{document}

### 2021-12-16 Presentation/Presentation.tex
PUB

--- 2021-12-12 Presentation/Presentation.tex
@@
 % Uncomment these lines for an automatically generated outline.
-\begin{frame}{Outline}
-\setcounter{tocdepth}{1}
-\begin{center}
-  \tableofcontents
-\end{center}
-\end{frame}
+%\begin{frame}{Outline}
+%\setcounter{tocdepth}{1}
+%\begin{center}
+%  \tableofcontents
+%\end{center}
+%\end{frame}
 
@@
 \Phi=\frac{L}{4\pi d^{2}}
-\end{equation}
-\begin{equation}
-m=-2.5\log_{10}\Phi+C
-\end{equation}
-\begin{equation}
-m_{1}-m_{2}=-2.5\log_{10}\frac{\Phi_{1}}{\Phi_{2}}
-\end{equation}
-\begin{equation}
-M-m=-2.5\log_{10}\frac{\Phi\cdot d^{2} }{\Phi\cdot 10^{2}}
-\end{equation}
-\begin{equation}
-M=m+5-5log_{10}d
-\end{equation}
-Where $\Phi$ is the flux for a sphere of radius r, m the relative magnitude and M the \textcolor{red}{Absolute magnitude}
-\end{frame}
-
-\begin{frame}{Celestial mechanics \cite{murray1999solar}: Magnitude}
-\begin{equation}
-\Phi=\frac{L}{4\pi r^{2}}
 \end{equation}
@@
 \begin{itemize}
-\item The mgm algorithm is not the best one in term of performances, but it provides the connections between the features.
-On the other side, except for the variable importance in RF, the other are black box one
-\item The mgm model, as the other graphical model is open to a true scientific validation, the other not. 
-\item The probabilistic models lack in the forecast is definitely compensated by their interpretability 
-\item This is meaningful since this two features are in conflict 
-\item The probabilistic models provide a good trade-off between intepretability and forecast performances, as long as one is interest to produce a really scientific result (e.g if the only aim is the forecast the RF is definitely better.
-However how long one can trust to the RF result ?)
+\item The mgm algorithm is not the best one in terms of performance, but it provides the connections between the features.
+On the other side, except for the variable importance in RF, the SVM and RF are a black box one
+\item The mgm model, as the different graphical model is open to a proper scientific validation, the SVM and the RF not. 
+\item The probabilistic models lack in the forecast is compensated by their interpretability 
+\item This is meaningful since interpretability and completeness/accuracy are in conflict 
+\item It is worth noting that the logistic model, which is highly interpretable, has performances similar to the mgm
+\item The probabilistic models provide a good trade-off between interpretability and forecast performances.
+As long as one is interested in producing a really scientific result, such algorithms as the logistic regression should be considered (e.g if the only aim is the forecast, the RF is better.
+However, how long one can trust the RF result ?)
 \end{itemize}
@@
 
-\section{Supporting Info}
-
-\subsection{ROC and $\phi$ factor}
-\begin{frame}{}
-\begin{center}
-{\Huge ROC and $\phi$ factor}
-\end{center}
-\end{frame}
-
-\begin{frame}
-\begin{center}
-\frametitle{$\phi$ coefficent}
-\begin{table}[]
-\begin{tabular}{c|c|c}
-\multicolumn{1}{l|}{\textbf{}} & \textbf{Actual - N} & \textbf{Actual - P} \\ \hline
-\textbf{Predicted - N}         & \#TP                & \#FN                \\ \hline
-\textbf{Predicted - P}         & \#FP                & \#TP               
-\end{tabular}
-\end{table}
-\begin{equation*}
-\phi=\dfrac{TP\times TN - FP\times FN}{\sqrt{\left(TP+FP\right)\left(TP+FN\right)\left(TN+FP\right)\left(TN+FN\right)}}
-\end{equation*}
-\end{center}
-\end{frame}
-
-\begin{frame}
-\frametitle{Receiver operating characteristic}
-\begin{columns}
-\column{0.5\textwidth}
-\begin{center}
-\includegraphics[width=1\linewidth]{Pic/ROC/ROC_A.png}
-\includegraphics[width=1\linewidth]{Pic/ROC/ROC_B.png}\\
-\end{center}
-\column{0.5\textwidth}
-\begin{center}
-\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_A_R.png}
-\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_B_R.png}\\
-\end{center}
-\end{columns}
-\begin{center}
-Images taken from \cite{ROC_fig}
-\end{center}
-\end{frame}
-
-\begin{frame}
-\frametitle{Receiver operating characteristic}
-\begin{columns}
-\column{0.5\textwidth}
-\begin{center}
-\includegraphics[width=1\linewidth]{Pic/ROC/ROC_C.png}
-\includegraphics[width=1\linewidth]{Pic/ROC/ROC_D.png}\\
-\end{center}
-\column{0.5\textwidth}
-\begin{center}
-\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_C_R.png}
-\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_D_R.png}\\
-\end{center}
-\end{columns}
-\begin{center}
-Images taken from \cite{ROC_fig}
-\end{center}
-\end{frame}
-
-\subsection{FAMD}
-\begin{frame}{}
-\begin{center}
-{\Huge FAMD}
-\end{center}
-\end{frame}
-
-\begin{frame}{Factor analysis of mixed data - FAMD}
-\begin{center}
-r(z,k) correlation coefficient (z and k quantitative) \\
-\end{center}
-\begin{center}
-$\eta^{2}$(z,q) correlation ratio (z quantitative and q qualitative)
-\end{center}
-\begin{center}
-PCA $\rightarrow$ $max\sum_{k}r^{2}(z,k)$
-\end{center}
-\begin{center}
-MCA $\rightarrow$ $max\sum_{q}\eta^{2}(z,q)$
-\end{center}
-\begin{center}
-FAMD $\rightarrow$ $max\sum_{k}r^{2}(z,k)+max\sum_{q}\eta^{2}(z,q)$
-\end{center}
-\end{frame}
-
-\subsection{Graphical models}
-\begin{frame}{}
-\begin{center}
-{\Huge Graphical models}
-\end{center}
-\end{frame}
-
-\begin{frame}{Graphical models}
-\begin{equation}
-p(x_{1},x_{2},...,x_{n})
-\end{equation}
-\begin{equation}
-p(x_{1:V})=p(x_{1})p(x_{2}|x_{1})p(x_{3}|x_{2},x_{1})...p(x_{V}|x_{1:V-1})
-\end{equation}
-\begin{equation}
-X  \perp Y| Z \iff  p(X,Y|Z) = p(X|Z)p(Y|Z)
-\end{equation}
-\begin{equation}
-p(\textbf{x}_{1:V})=p(x_{1})\prod^{V}_{t=1}p(x_{t}|x_{t-1})
-\end{equation}
-\end{frame}
-
-\begin{frame}{Graphical models}
-\begin{theorem}[Hammersley-Clifford]
-A positive distribution $p(\textbf{y})>0$ satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
-\begin{equation}
-p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
-\end{equation}
-where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
-\begin{equation}
-Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
-\end{equation}
-Note that this partition function is what ensures the overall distribution sums to 1
-\end{theorem}
-\end{frame}
-
-\begin{frame}{Graphical models}
-\begin{equation}
-p(y|\theta)=\dfrac{1}{Z(\theta)} exp\left(-\sum_{c}E(y_{c}|\theta_{c})\right)
-\end{equation}
-\begin{equation}
-\psi_{c}(y_{c}|\theta_{c})=exp\left(-E(y_{c}|\theta_{c})\right)
-\end{equation}
-\end{frame}
-
-\begin{frame}{Parameters tuning}
-\begin{equation}
-\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\frac{1}{N}\sum_{i}\left[\phi_{c}(y_{i})-\frac{\partial}{\partial\boldsymbol{\theta}_{c}}\log Z(\boldsymbol{\theta})\right]
-\end{equation}
-
-\begin{equation}
-\begin{split}
-\frac{\partial \log Z(\boldsymbol{\theta})}{\partial\boldsymbol{\theta}}=\mathbb{E}\left[\phi_{c}(\textbf{y})\right|\theta]= \sum_{\textbf{y}}\phi_{c}(\textbf{y})p(\textbf{y}|\boldsymbol{\theta})
-\end{split}
-\end{equation}
-\begin{equation}
-\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\left[\frac{1}{N}\sum_{i}\phi_{c}(y_{i})\right]-\mathbb{E}\left[\phi_{c}(\textbf{y})\right]
-\end{equation}
-\begin{equation}
-\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]-\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
-\end{equation}
-\begin{equation}
-\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]=\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
-\end{equation}
-\end{frame}
-
-\begin{frame}{Parameters tuning}
-\begin{equation}
-p\left(\textbf{y}|\boldsymbol{\theta}\right) = \dfrac{1}{Z(\theta)}\exp \left( \sum_{c}\boldsymbol{\theta}^{T}_{c}\phi_{c}\left(\textbf{y}\right)\right)
-\end{equation}
-\begin{equation}
-\mathcal{L}\left(\boldsymbol{\theta}\right):= \frac{1}{N}\sum_{i}\log p\left(\textbf{y}_{i}|\boldsymbol{\theta}\right)=\frac{1}{N}\sum_{i}\left[\sum_{c} \boldsymbol{\theta}^{T}_{c}\phi_{c}(y_{i})-\log Z\left(\boldsymbol{\theta}\right)\right]
-\end{equation}
-\end{frame}
-
-\begin{frame}{Gaussian graphical model}
-\begin{equation}
-f_{Y}(y)=\dfrac{det(K)^{1/2}}{(2\pi)^{d/2}}\exp\left(-\dfrac{1}{2}\textbf{y}^{T}K\textbf{y}\right)
-\end{equation}
-
-\begin{equation}
-K=\Sigma^{-1}
-\end{equation}
-
-\end{frame}
-
-\begin{frame}{GLASSO}
-\begin{equation}
-L_{pen}(K,\hat{\mu})=log\det(K)-tr(K\;S)-\rho||K|| 
-\end{equation}
-\begin{center}
-$K=\Sigma^{-1}$
-\end{center}
-\begin{center}
-S: empirical covariance matrix
-\end{center}
-\end{frame}
-
-\begin{frame}{mgm algorithm}
-\begin{equation}
-P(X_{s}|X_{\backslash s})=\exp\left\lbrace E_{s}(X_{\backslash s})\phi_{s}\left(X_{s}\right)+B_{s}(X_{s})-\Phi\left(X_{\backslash s}\right) \right\rbrace
-\end{equation}
-\begin{center}
-$\phi_{s}$ function of sufficient statistics $B_{s}$ base measure
-\end{center}
-
-\begin{equation}
-\begin{split}
-P(X)= & \exp \left( \sum_{s\in V}\theta_{s}\phi_{s}(X_{s})+\sum_{s\in V}\sum_{r\in N(s)}\theta_{s,r}\phi_{s}(X_{s})\phi_{r}(X_{r})\\ +&...+\sum_{r_{1},...,r_{k}\in C}\theta_{r1,...,rk}\prod_{j=1}^{k}\phi_{r}_{j}\left(X_{r}_{j}\right)+\sum_{s\in V}B_{s}(X_{S})-\Phi(\theta) \right)
-\end{split}
-\end{equation}
-
-\begin{equation}
-\hat{\theta}=arg \min_{theta}\left\lbrace -\mathcal{L}(\theta,X)+\lambda ||\theta ||_{1}  \right\rbrace\quad ||\theta ||_{1}=\sum_{j=1}^{J}|\theta_{j}|
-\end{equation}
-
-\end{frame}
-
-\begin{frame}{mmod - Homogeneus Mixed Interaction (HMI) models}
-\begin{center}
-N observation, d discrete variables and q continuous variable.
-The observation has the form $x=(i,y)=(i_{1},...,i_{d},y_{1}...y_{q})$.
-The probability of discrete variables falling in the cell i is denoted as $p(i)$.
-The conditional distribution of continous variables to fall in the cell $i$ is given by the multivariate gaussian $\mathcal{N}(\mu(i),\Sigma)$
-\end{center}
-
-\begin{equation}
-f(\textbf{i},\textbf{y})= p(\textbf{i})(2\pi)^{-q/2}det(\Sigma)^{-1/2}\exp\left[-\dfrac{1}{2}\left(\textbf{y}-\mu(\textbf{i})\right)^{T}\Sigma^{-1}\left(\textbf{y}-\mu(\textbf{i})\right)\right]
-\label{gaussMix}
-\end{equation}
-
-\begin{equation}
-f(\textbf{i},\textbf{y})=\exp\left[g(\textbf{i}) +h(\textbf{i})^{T}\textbf{y}-\frac{1}{2}y^{T}K\textbf{y}\right]
-\end{equation}
-
-\begin{center}
-where $K$ is the concentration matrix, $g(i)$ and $h(i)$ are the log-linear expansion of the probability $p(\textbf{i})$ (canonical parameters)
-\end{center}
-
-\end{frame}
-\begin{frame}{mmod}
-\begin{center}
-Decomposable model
-\end{center}
-\begin{equation}
-\hat{f}(x)=\prod_{j=1}^{k}\dfrac{\hat{f}_{C_{j}}(x_{C}_{j})}{\hat{f}_{S_{j}}(x_{S}_{j})} 
-\end{equation}
-Maximized likehood
-\begin{equation}
-\hat{\mathcal{L}_{s}}=\sum_{i}n(i)log\left(\dfrac{n(i)}{N}\right)-Nq\dfrac{log(2\pi)}{2}-N\dfrac{log(det(S))}{2}-N\frac{q}{2}
-\end{equation}
-
-\begin{equation}
-\hat{p}(\textbf{i})=n(\textbf{i})/N \quad \hat{\mu}(\textbf{i})=\bar{y}(\textbf{i})\quad\hat{\Sigma}=S=\sum_{i}n(i)S_{i}/N
-\end{equation}
-
-\end{frame}
-
-\subsection{Information theory}
-\begin{frame}{}
-\begin{center}
-{\Huge Information theory}
-\end{center}
-\end{frame}
-
-\begin{frame}{Information theory}
-\begin{equation}
-H(X)=-\sum_{x \in X} p(x)log p(x)
-\end{equation}
-\begin{equation}
-H(X,Y)=-\sum_{x\in \chi}\sum_{y\in \mathcal{Y}}p(x,y)\log p(x,y)
-\end{equation}
-\begin{equation}
-\begin{split}
-H(X|Y)=& \sum_{x\in \mathcal{X} }p(x)H(Y|X=x)\\
-=& -\sum_{x\in \mathcal{X}}p(x)\sum_{y\in \mathcal{Y}}p(x,y)\log p(y|x) \\
-=& -E\log p(Y|X)
-\end{split}
-\end{equation}
-\begin{equation}
-H(X,Y)=H(X)+H(Y|X)
-\end{equation}
-\begin{equation}
-D(p||q)=\sum p(x)\log\frac{p(x)}{q(x)} \quad D(p||q)\geq 0
-\end{equation}
-
-\end{frame}
-
-\begin{frame}{Information theory}
-\begin{equation}
-\begin{split}
-I(X;Y)&=\sum\sump(x,y)\log\dfrac{p(x,y)}{p(x)p(y)}=D(p(x,y)||p(x)p(y)) \\
-&=H(X)-H(X|Y)=H(Y)-H(Y|X)
-\end{split}
-\end{equation}
-\begin{equation}
-I(X_{1},X_{2},...,X_{n};Y)=\sum^{n}_{i=1}I(X_{i};Y|X_{i-1},X_{i-2},...,X_{1})
-\end{equation}
-\begin{equation}
-g(\alpha,\textbf{C},\textbf{S},f_{i})=MI(f_{i};\textbf{C})-\sum_{f_{s}\in S}\alpha(f_{i},f_{s},\textbf{C},\textbf{S})MI(f_{i};f_{s})
-\end{equation}
-\end{frame}
-
 \end{document}

### 2021-12-16 OLD/Manuscript/Script_FINAL.R

@@ -1,0 +1,379 @@
+library(readxl)
+library(PerformanceAnalytics)
+library(corrplot)
+library(ggcorrplot)
+library(mgm)
+library(gRim)
+library(bnlearn)
+library(gRapHD)
+library(qgraph)
+library(Rgraphviz)
+library(igraph)
+library(gRbase)
+library(RBGL)
+library(caret)
+library(pROC)
+library(ROCR)
+library(ggplot2)
+library(FactoMineR)
+library(factoextra)
+library(ggplot2)
+library(ggpubr)
+library(gRapHD)
+library(glasso)
+library(SIN)
+library(gRain)
+library(randomForest)
+library(randomForestExplainer)
+library(e1071)
+library(MASS)
+library(infotheo)
+library(varrank)
+library(minet)
+
+#Data Loading: this portion of code loads the original asteroid dataset and randomly extracts a set of not Hazardous asteroids. It is not necessary to 
+#execute since the operation was previously performed and its output was put on to an .rda file. 
+
+#Asteroids2 <- read_excel("Dataset/Asteroids2.xlsx")
+#Asteroids2<-na.omit(Asteroids2)
+
+#Asteroids2_h<-subset(Asteroids2,Hazardous==TRUE)
+#Asteroids2_n<-subset(Asteroids2,Hazardous==FALSE)
+
+#Asteroids_NH<- sample(3878,2800)
+
+#Asteroids2_n<-Asteroids2[Asteroids_NH,]
+
+#Asteroids_NH_frame<-Asteroids2[Asteroids_NH,]
+
+#Asteroids_FINAL<-rbind(Asteroids2_h,Asteroids2_n)
+
+#save(Asteroids_FINAL,file="Asteroids_FINAL.rda")
+
+load("Asteroids_FINAL.rda")
+
+Asteroids_FINAL[,2:21]<-scale(Asteroids_FINAL[,2:21])
+
+Asteroids_FINAL_double<-as.data.frame(lapply(Asteroids_FINAL[,2:22], as.double))
+Asteroids_FINAL_double["Hazardous"] <- as.factor(Asteroids_FINAL_double[,21])
+
+#Preliminary analysis
+
+SS <- CGstats(Asteroids_FINAL_double)
+SS
+cc<-apply(SS$center,1,sd) /apply(SS$center,1,mean)
+cc
+
+M_cov <- cov(Asteroids_FINAL_double[,-21])
[…]

### 2021-12-16 OLD/Manuscript/latex.py

@@ -1,0 +1,1 @@
+jobname="manuscript"

### 2021-12-16 OLD/Manuscript/manuscript.py

@@ -1,0 +1,17 @@
+model_cov = keras.models.Sequential([
+      keras.layers.Conv2D(64,7,activation="relu",padding="same",input_shape=[64,64,3]),
+      keras.layers.MaxPooling2D(2),
+      keras.layers.Conv2D(128,3,activation="relu",padding="same"),
+      keras.layers.Conv2D(128,3,activation="relu",padding="same"),
+      keras.layers.MaxPooling2D(2),
+      keras.layers.Conv2D(256,3,activation="relu",padding="same"),
+      keras.layers.Conv2D(256,3,activation="relu",padding="same"),
+      keras.layers.MaxPooling2D(2),
+      keras.layers.Flatten(),
+      layers.Dense(128, kernel_initializer="he_normal", name="layer1",kernel_regularizer=tf.keras.regularizers.L2(0.01)),
+      keras.layers.LeakyReLU(alpha=0.2),
+      layers.Dense(64, kernel_initializer="he_normal", name="layer2",kernel_regularizer=tf.keras.regularizers.L2(0.01)),
+      keras.layers.LeakyReLU(alpha=0.2),
+      keras.layers.Dense(10,activation="softmax")
+])
+
